---
title: "Measuring the growth and *lac* induction lags in switching environments"
author: Thomas Julou
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

## Estimation

First, let's select the cells that undergo a switch (i.e. the first frame occured before a switch to lactose, and the last frame after (or at) the switch). We "prolonge" each cell by summing the variables (length and total GFP) of its two daughters, and discard cells for which the acquisition stopped less than 30' after the switch (10 points; it is difficult to distinguish growth from lag on shorter traces).

Based on the log of the length traces, the growth lag is computed by fitting a flat lag followed by a linear increase to the subset of each trace where the length is smaller than 3 times the length at the switch (note that the cells do more than doubling because daughters' lengths are summed).

Estimating the induction lag is more cumbersome because GFP traces do not follow a general shape. We compute two lag estimates:

- the GFP lag is computed by fitting a flat lag followed by a linear increase to the subset of each GFP trace where it is lower than the GFP at the switch plus 1000 GFP molecules (NB: the typical value at the switch is 100 molecules while fully induced cells have ca. 6000 molecules in average).
- the lag\_200 is computed by finding the delay until the cell increases its GFP by 200 molecules after the switch. <!-- If the GFP becomes lower than this value after this delay, then lag\_200 is undefined. -->

```{r extract frames for lag estimation}
myframes_switching <- myframes %>% 
  ungroup %>% 
  filter(str_detect(condition, "switch_")) %>% 
   # discard pos with low acq rate (phototoxicity control)
   filter(!(date==20181008 & pos %in% 5:9), 
          !(date %in% c(20181009, 20181024) & pos %in% c(1,3,5,7,9))) %>% 
  filter(!discard_start, !discard_top) %>%
  select(condition, date, pos, gl, id, genealogy, start_time, end_time, end_type, dt,
         time_sec, medium, m_start, m_end, gfp_nb, length_um, m_cycle, cell_num_in_lane, total_cell_in_lane) %>% 
  # prolonge traces with daughters' data
  group_by(date, pos, gl) %>%
  partition(mycluster) %>%
  do((function(.dfgl) {
    # browser()
    .dfgl %>% 
      group_by(id) %>% 
      mutate(is_lac=str_detect(medium, 'lactose') | str_detect(medium, 'lactulose') | 
               str_detect(medium, fixed('+TMG')) | str_detect(medium, 'lacCM'), 
             t_lac_switch=min(m_start[is_lac]), t_lac_end=min(m_end[is_lac])) %>% 
      filter(first(time_sec)<=t_lac_switch & last(time_sec)>=unique(t_lac_switch)) %>% 
      do((function(.dfc) bind_rows(
        mutate(.dfc, type='cell'),
        .dfc %>% 
          ungroup %>% slice(1L) %>% # head(1) %>% 
          (function(.df)
            # find the genealogy of the daughters and extract the corresponding rows from myframes
           bind_rows(mutate(.df, join_gen=paste0(genealogy, 'B')),
                     mutate(.df, join_gen=paste0(genealogy, 'T')) ) %>% 
            semi_join(.dfgl %>% mutate(join_gen=genealogy),
                      ., by='join_gen') ) %>%
         ungroup %>% mutate(genealogy=get_parent_cid(genealogy)) %>% 
         select(condition, date, pos, gl, id, genealogy, dt,
                time_sec, medium, m_cycle, m_start, m_end, gfp_nb, length_um) %>% 
         # group by time, compute the stat of both daughters' and keep only the first row
         # compared to summarise, this keeps all other variables
         group_by(genealogy, time_sec) %>% 
         mutate(gfp_nb=ifelse(n()==2, sum(gfp_nb), 2*gfp_nb), 
                length_um=ifelse(n()==2, sum(length_um), 2*length_um), 
                # medium=unique(medium), m_cycle=unique(m_cycle),
                type=paste0(n(), "dg")) %>%
         slice(1L) 
     ) )(.))
   })(.) ) %>% 
  collect() %>% 
# myframes_switching <- myframes_switching %>% 
  # fill in missing values
  group_by(date, pos, gl, genealogy) %>%
  fill(start_time, end_time, t_lac_switch, t_lac_end) %>% 
  mutate(ugen=paste(date, pos, gl, genealogy, sep='.'),
         pre_switch=time_sec>t_lac_switch+60 & time_sec<t_lac_switch+10*60) %>% 
  # select data used for lag estimation
  group_by(condition, ugen) %>%
  mutate(switch_idx=first(m_cycle),
         fit_lag=ifelse(
    type!='1dg' & # discard traces after one daughter is lost (because of discontinuity)
      tail(time_sec[type!='1dg']-t_lac_switch[type!='1dg'], 1) > first(4 * dt) & # require at least 4 frames after the switch (one point after the gfp_ini window)
      m_start==t_lac_switch # focus on the medium after the switch
    , TRUE, FALSE) )

```

Let's visualise the data used to estimate the lags:

```{r fig.height=21}
myframes_switching %>% 
  filter(fit_lag) %>% 
  ggplot(aes(time_sec - 2*3600, gfp_nb, col=ugen, alpha=type)) +
  facet_wrap(~condition+date, dir="v", ncol=2, labeller=labeller(.cols=label_value, .multi_line=FALSE)) +
  # geom_rect(aes(xmin=t_start - 2*3600, xmax=t_end - 2*3600, ymin=-Inf, ymax=Inf, x=NaN, y=NaN, group=1), fill=rgb(1, 0, 0, .1), col='red', data=filter(condition_ts, condition=='switch', medium=='lactose')) +
  geom_path() + # geom_point() +
  scale_colour_periodic(guide='none') +
  scale_alpha_discrete(range=c(0.1, 0.6)) +
  scale_x_hours() + #, limits=c(19*3600, 23*3600)) +
  ylim(-100, 1e4) +   expand_limits(x=0, y=0) +
  labs(y='total fluorescence (GFP molecules)') +
  theme(legend.position="top")

# myframes_switching %>% 
#   filter(condition=='switch_long_lac_hiExpr') %>% 
#   filter(fit_lag) %>% 
#   # group_by(date, pos, gl, id) %>% summarise(n=n()) %>% qplot(n, data=., binwidth=10)
# ggplot(aes(time_sec - 2*3600, length_um, col=ugen)) +
#   facet_wrap(~condition, ncol=1) +
#   # geom_rect(aes(xmin=t_start - 2*3600, xmax=t_end - 2*3600, ymin=-Inf, ymax=Inf, x=NaN, y=NaN, group=1), fill=rgb(1, 0, 0, .1), col='red', data=filter(condition_ts, condition=='switch', medium=='lactose')) +
#   # geom_path() + # geom_point() +
#   stat_smooth(se=FALSE, geom='line', alpha=.2) +
#   scale_colour_periodic(guide='none') +
#   # scale_alpha_discrete(range=c(0.1, 0.6)) +
#   scale_x_hours(limits=c(3*3600, 12*3600)) +
#   scale_y_continuous(trans='log2')

```


```{r lag estimation}
mycells_switching <- myframes_switching %>%
  filter(fit_lag) %>% 
  group_by(condition, switch_idx, date, pos, gl, ugen) %>%
  partition(mycluster) %>%
  do((function(.df){
    # browser()
    # print(unique(.df$ugen))
    # if (unique(.df$ugen)=="20150703.0.11.0:BBBBTBB") browser()
    # if (unique(.df$ugen)=="20150703.3.11.0:BBTBBB") browser()
    
    .cond <- head(.df$condition)
    .preswitch_gfp <- mean(filter(.df, pre_switch)$gfp_nb)
    .preswitch_length <- mean(filter(.df, pre_switch)$length_um)
    # compute growth lag
    .growth_lag <- .df %>% filter(time_sec>=t_lac_switch, length_um<3*.preswitch_length) %>%
      pull(length_um) %>% log %>% compute_lag(.scaling_factor=1) #,.plot=TRUE)
    if (length(.growth_lag) == 0) { warning('growth lag estimation failed')
      .growth_lag <- list(tau_idx=NaN, x_lag=NaN, slope_after=NaN)
    }

    if (.df$condition %in% c('switch_long_lac_hiExpr')) {
      .gfp_lag <- list(tau_idx=NaN, x_lag=NaN, slope_after=NaN)
      .t_200 <- NaN
      .t_200_fail <- NaN
      .t_1000 <- NaN
      .t_1000_fail <- NaN
    } else {
    .gfp_lag <- .df %>% filter(time_sec>=t_lac_switch, gfp_nb<.preswitch_gfp+ifelse(.cond!='switch_iptg', 1000, 2000)) %>%
      pull(gfp_nb) %>% compute_lag #(.plot=TRUE)
    if (length(.gfp_lag) == 0)
      .gfp_lag <- list(tau_idx=NaN, x_lag=NaN, slope_after=NaN)

    # compute induction lag
    .tmin <- filter(.df, pre_switch)$time_sec %>% max
    .t_200 <- .df %>% filter(time_sec>.tmin, gfp_nb>.preswitch_gfp+200) %>% pull(time_sec) %>% first() # returns NA if empty
    .t_200_fail <- .df %>% filter(time_sec>.t_200, gfp_nb<.preswitch_gfp+200) %>% pull(time_sec) %>% first()
    .t_1000 <- .df %>% filter(time_sec>.tmin, gfp_nb>.preswitch_gfp+1000) %>% pull(time_sec) %>% first() # returns NA if empty
    .t_1000_fail <- .df %>% filter(time_sec>.t_1000, gfp_nb<.preswitch_gfp+1000) %>% pull(time_sec) %>% first()
    }
    
    .dt <- first(.df$dt) # don't use unique as some values are NA
    return(data.frame(dt=.dt, n_fit=nrow(.df), genealogy=gsub(":", "", unique(.df$genealogy)), 
                      t_lac_switch=first(.df$t_lac_switch), t_lac_end=first(.df$t_lac_end),
                      tmax_lag_fit=max(.df$time_sec), is_tmax_end=max(.df$time_sec)==first(.df$t_lac_end),
                      cell_num=filter(.df, time_sec==t_lac_switch)$cell_num_in_lane,
                      cell_num_from_top=filter(.df, time_sec==t_lac_switch)$cell_num_in_lane, 
                      cell_num_from_bottom=filter(.df, time_sec==t_lac_switch)$total_cell_in_lane-filter(.df, time_sec==t_lac_switch)$cell_num_in_lane,
                      # switch_idx=filter(.df, is_lac)$m_cycle %>% first, 
                      time_birth=unique(.df$start_time), time_div=unique(.df$end_time),
                      growth_lag=(.growth_lag$tau_idx-1)*.dt, growth_lag_length=exp(.growth_lag$x_lag), growth_lag_rate_after=.growth_lag$slope_after / .dt,
                      gfp_lag=(.gfp_lag$tau_idx-1)*.dt, gfp_lag_length=.gfp_lag$x_lag, gfp_lag_slope_after=.gfp_lag$slope_after / .dt,
                      gfp_ini=.preswitch_gfp, length_ini=.preswitch_length, 
                      lac_200=.t_200, lac_200_fail=.t_200_fail,
                      lac_1000=.t_1000, lac_1000_fail=.t_1000_fail))
    })(.)) %>%
  collect() %>% 
  arrange(condition, ugen) %>%  # sort data after `partition()`
  # mycells_switching <- mycells_switching %>% 
  mutate(lag_200=ifelse(is.na(lac_200) & !is.nan(lac_200), Inf, lac_200-t_lac_switch), # set lag_200 to Inf if no restart (is `& !is.nan(lac_200)` really needed?)
         lag_1000=ifelse(is.na(lac_1000) & !is.nan(lac_1000), Inf, lac_1000-t_lac_switch)) %>%  
  # # more complicated heuristic (dropped during MoMA_Methods revisions)
  # mutate(lac_200=ifelse(is.na(lac_200), Inf, lac_200), # set lag_200 to Inf if no restart
  #        lag_200 = ifelse(is.na(lac_200_fail) | lac_200_fail-lac_200 > 20*60, lac_200-t_lac_switch, NaN) ) %>%
# discard lag fit with slope too close to 0
  mutate(gfp_lag=ifelse(!is.na(gfp_lag_slope_after) & between(gfp_lag_slope_after, 0.12, Inf), gfp_lag, Inf),
         gfp_lag_slope_after=ifelse(!is.na(gfp_lag_slope_after) & between(gfp_lag_slope_after, 0.12, NaN), gfp_lag_slope_after, NaN),
         growth_lag=ifelse(!is.na(growth_lag_rate_after) & between(growth_lag_rate_after, 4e-5, 2.5e-4), growth_lag, Inf),
         growth_lag_rate_after=ifelse(!is.na(growth_lag_rate_after) & between(growth_lag_rate_after, 4e-5, 2.5e-4), growth_lag_rate_after, NaN)) %>% 
  # set to NA lags of uninduced cells with traces shorter than the lactose episode
  mutate(reach_lac_end = ifelse(!is_tmax_end, near(tmax_lag_fit, t_lac_end-dt, to=1e-3), 
                                near(tmax_lag_fit, t_lac_end, to=1e-3)) ) %>% 
  mutate(lag_200=ifelse(is.infinite(lag_200) & !reach_lac_end, NA, lag_200), 
  			 lag_1000=ifelse(is.infinite(lag_1000) & !reach_lac_end, NA, lag_1000), 
  			 gfp_lag=ifelse(is.infinite(gfp_lag) & !reach_lac_end, NA, gfp_lag),
         growth_lag=ifelse(is.infinite(growth_lag) & !reach_lac_end, NA, growth_lag) )

# ggplot(mycells_switching) +
#   facet_grid(condition~., margins=TRUE, scales='free') +
#   geom_vline(xintercept=0.12, col='red') +
#   geom_histogram(aes(gfp_lag_slope_after), binwidth=0.025)
# 
# ggplot(mycells_switching) +
#   facet_grid(condition~., margins=TRUE, scales='free') +
#   geom_vline(xintercept=4e-5, col='red') +
#   geom_histogram(aes(growth_lag_rate_after), binwidth=2e-6)

  
# add fit prediction and residual to myframes_switching
myframes_switching <- myframes_switching %>% 
  group_by(ugen) %>% 
  mutate(switch_idx=m_cycle[is_lac] %>% first) %>% 
  ungroup %>% 
  left_join(mycells_switching %>% ungroup %>% 
              select(ugen, growth_lag, growth_lag_length, growth_lag_rate_after, 
                     gfp_lag, gfp_lag_length, gfp_lag_slope_after,
                     gfp_ini, lag_200, lac_200, lag_1000, lac_1000)) %>%
  mutate(length_predict=ifelse(time_sec<t_lac_switch+growth_lag, growth_lag_length,
    growth_lag_length * exp(growth_lag_rate_after*(time_sec-t_lac_switch-growth_lag) )),
         length_predict=ifelse(time_sec>=t_lac_switch, length_predict, NA),
         llength_residual=log(length_um)-log(length_predict))

# # visual control
# for(i in 1:1000) {
# filter(myframes_switching, cell==nth(unique(ugen), i)) %>%
# filter(myframes_switching, ugen==unique(ugen)[i]) %>%
#   (function(.x) ggplot(.x)+
#   geom_vline(aes(xintercept = unique(t_lac_switch-start_time+growth_lag)), lty='dashed') +
#   geom_point(aes(time_sec-start_time, log(length_um), col=medium)) +
#   geom_line(aes(time_sec-start_time, log(length_predict))) ) %>%
#     print
# scan("")
# }

```

We use the following convention:

- NaN: lag estimation failed (or no lac-GFP data)
- NA: lag longer than the gfp (resp. length) time series (when time series is shorter than the lactose episode)
- Inf: lag longer than the lactose episode


```{r growth rate before the switch}
mycells_switching <- 
  # extract relevant frames
  # don't discard NA lags otherwise filter(!discard_arrested) also discard NA lags
  bind_rows(
    # switching cells
    mycells_switching %>% ungroup %>% 
      # filter(!is.na(lag_200)) %>% 
      mutate(type='pre', switch_ugen=ugen) %>% 
      select(type, ugen, switch_ugen, switch_idx, t_lac_switch) %>% 
      left_join(myframes) %>% 
      filter(time_sec <= t_lac_switch),
    # parent cells
    mycells_switching %>% ungroup %>% 
      # filter(!is.na(lag_200)) %>% 
      left_join(select(myframes, date, gl, pos, id, parent_id, ugen) %>% group_by(ugen) %>% slice(1)) %>% 
      mutate(type='parent', switch_ugen=ugen, 
             # cell=paste(date, pos, gl, parent_id, sep='.')) %>% 
             ugen=get_parent_cid(ugen) ) %>% 
      select(type, ugen, switch_ugen, switch_idx, t_lac_switch) %>% 
      left_join(myframes) %>% 
      filter(time_sec <= t_lac_switch)
  ) %>% 
  # compute elongation rate for these cells
  group_by(type, ugen, switch_ugen) %>% 
  do((function(.df) {
    # browser()
    if (dim(.df)[1] < 5) return(data.frame())
    
    .mod <- lm( log(length_um)~time_sec, .df)
    data.frame(logl_time_slope=.mod$coefficients[2],
               logl_time_r2=cor(.df$time_sec, log(.df$length_um))^2)
  })(.)) %>% 
  # reshape and join to mycells_switching
  ungroup %>% 
  mutate(ugen=switch_ugen, switch_ugen=NULL) %>% 
  gather(var, val, -type, -ugen) %>% 
  mutate(var=paste(var, type, sep='_'), type=NULL) %>% 
  spread(var, val) %>% 
  right_join(mycells_switching) %>% 
  mutate(n_preswitch=(t_lac_switch-time_birth)/dt,
         logl_time_slope_before=ifelse(n_preswitch<5, 
                                       logl_time_slope_parent, logl_time_slope_pre),
         discard_arrested=ifelse(condition=='switch_gly_lac', NA, logl_time_slope_before <= min_growth_rate),
         discard_arrested_=discard_arrested,
         discard_arrested=FALSE, # disable this filtering to keep analysis description simple
         )
  

# ggplot(mycells_switching) +
#   geom_abline() +
#   geom_point(aes(logl_time_slope_parent, logl_time_slope_pre, size=n_preswitch), alpha=0.2) +
#   coord_cartesian(ylim=c(0, 3e-4))

# ggplot(mycells_switching) +
#   geom_point(aes(n_preswitch, 1-logl_time_r2_pre), size=0.2, alpha=0.2) +
#   scale_y_log10() +
#   xlim(0, 50)

```

