---
title: "Controls"
author: Thomas Julou
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

## Growth between switches

First we perform quality control for all mother machine experiments by checking how cells are growing between switches. For each cell cycle, growth rate and $r^2$ are computed from a fit of exponential elongation; we compare their distributions between experiments.

```{r}
# extract cell cycle params for those not overlapping with a switch
mycells_btw <- myframes %>% 
  ungroup %>% 
  # filter(!(condition %in% c('glucose', 'lactose', 'mg1655'))) %>% 
  filter(!discard_top) %>%
  filter(start_time>0, start_time>=m_start, end_type=='div') %>% 
  # select(condition, date, pos, gl, id, genealogy, start_time, end_time, end_type, 
         # time_sec, medium, m_start, m_end, gfp_nb, length_um, m_cycle, cell_num_in_lane, total_cell_in_lane) %>% 
  group_by(condition, date, pos, gl, id) %>%
  partition(mycluster) %>%
  # filter(start_time>m_start, end_type=='div', # full cell cycles only
  #        length(unique(interaction(medium, m_cycle)))==1) %>%  # not overlaping a switch
  filter((unique(medium)!='glycerol' & n()>9) | # at least 10 time points (≈1/3 of cell cycle)
           (unique(medium)=='glycerol' & n()>19)) %>% # requires 20 time points in glycerol
  do((function(.df) {
    # browser()
    # .mod_ll_t <- fastLmPure( cbind(1, .df$time_sec), log(.df$length_um) )
    .mod_ll_t <- lm( log(length_um)~time_sec, .df)  # use fastLm() for predict
    .time_birth <- first(.df$time_sec)
    .time_div <- last(.df$time_sec)
    .logl <- predict(.mod_ll_t, se.fit=TRUE)
    data.frame(npoints=.mod_ll_t$df.residual+1,
               time_birth=.time_birth, time_div=.time_div, 
               cell_num_from_top=mean(.df$cell_num_in_lane),
               cell_num_from_bottom=mean(.df$total_cell_in_lane-.df$cell_num_in_lane), 
               # logl_time_slope=.mod_ll_t$coefficients[2], logl_time_slopesd=.mod_ll_t$stderr[2],
               loglength_start=first(.logl$fit), loglength_startse=first(.logl$se.fit), 
               loglength_end=last(.logl$fit), loglength_endse=last(.logl$se.fit), 
               logl_time_slope=.mod_ll_t$coefficients[2], logl_time_slopesd=summary(.mod_ll_t)$coefficients[2,2],
               logl_time_r2=cor(.df$time_sec, log(.df$length_um))^2 )
  })(.) ) %>% 
  collect() %>% 
  arrange(condition, date, pos, gl, id) %>% 
  left_join(myframes %>% group_by(date, pos, gl, id) %>% slice(1) %>% 
              select(date, pos, gl, id, medium, m_start, m_cycle, mstep, gl_id, cell, ugen, parent_id) )

```

This overview of growth rates over all experiments (only 1/10th of data shown, lines are loess fits computed on the full dataset) is useful to visualise changes of growth rate between conditions, as well as the transient when cells are exposed to a new condition. In particular, one can see that:

- at the start of all experiment, the growth rate decreases during the first two hours (due to fluorescence illumination, discarded in further analysis)
- after each switch, growth rates returns to its steady state value (specific to the new condition) after ca. 2h.

```{r fig.height=16}
mycells_btw %>% 
  filter(!is.na(medium)) %>% 
  group_by(condition) %>% 
  mutate(m_col=as.numeric(fct_inorder(factor(medium))) ) %>% # %>% interaction(m_cycle)
  ggplot(aes(time_birth, logl_time_slope/log(2) * 3600, col=factor(m_col))) +
  facet_wrap(~condition+date, ncol=2, dir="v",  strip.position="right",
             labeller=as_labeller(rename_conds)) +
  geom_point(alpha=1, size=.2, data=~sample_frac(., size=.1)) +
  geom_smooth(aes(group=interaction(m_col, m_cycle)), method='loess', show.legend=FALSE) +
  scale_x_hours() +
  labs(x="raw time (h)",  y="growth rate (doublings/h)", col="medium type") +
  guides(col = guide_legend(override.aes=list(alpha = 1, size=2))) +
  # ylim(0, NA) +
  theme(legend.position='top') +
  background_grid(major="y", minor = "none") +
  NULL

```


### Growth in M9 glucose before the first switch

Next, we focus on growth rate distributions in  initial conditions (M9 + 0.2% glucose or M9 + 0.4% glycerol) since we want to compare cells switching from similar physiological states on different days (and hence discard outlier experiments if any).

For each experiment, the distribution of initial growth rate is stratified by time. As noted previously, this demonstrates that steady state is reached in less than 2 hours. Note that grey lines show constant glucose as a reference.

```{r fig.height=10}
mycells_btw %>% 
  group_by(condition) %>% 
  mutate(m_col=as.numeric(fct_inorder(factor(medium))) %>% interaction(m_cycle)) %>% 
  filter(m_col=='1.1') %>% 
  ungroup %>% 
  filter(!is.na(medium)) %>% 
  mutate(time_bin=Hmisc::cut2(time_birth/3600, cuts=(0:15)*2)) %>% 
  ggplot(aes(logl_time_slope/log(2) * 3600)) +
  facet_wrap(~condition+date, labeller=as_labeller(rename_conds)) +
  stat_ecdf(aes(col=time_bin), alpha=.6) +
  stat_ecdf(lty='dotted', data = ~filter(., condition=='glucose') %>% select(-condition, -date)) +
  # ggplot2::scale_color_discrete() +
  scale_colour_viridis_d() +
  coord_cartesian(xlim=c(0, 1.5)) +
  labs(x='growth rate (doublings/h)', y='cdf', col='birth time (h)') +
  theme(legend.position='bottom') +
  NULL

```

In order to identify days where initial growth is slower than expected, distributions of growth rates are compared between 2h and 6h. Note that dotted lines show constant glucose (as a reference).

```{r fig.height=12}
mycells_btw %>% 
  filter(!is.na(medium)) %>% 
  # left_join(condition_ts %>% select(condition, medium, m_id) %>% distinct()) %>% 
  # mutate(m_col=interaction(m_id, m_cycle)) %>% 
  group_by(condition) %>% 
  mutate(m_col=as.numeric(fct_inorder(factor(medium))) %>% interaction(m_cycle)) %>% 
  filter(m_col=='1.1') %>% 
  ungroup %>% 
  filter(between(time_birth/3600, 2, 6),
         medium!='glycerol') %>% 
  ggplot(aes(logl_time_slope/log(2) * 3600)) +
  facet_wrap(~condition+date, labeller=as_labeller(rename_conds), ncol=6) +
  stat_ecdf(col='grey40', data=~filter(., condition=='glucose') %>% select(-condition, -date)) +
  stat_ecdf(col=qual_cols[1]) +
  # stat_ecdf(aes(group=gl_id), col=qual_cols[1], size=.2) +
  coord_cartesian(xlim=c(0, 1.5)) +
  labs(x='growth rate (doublings/h)', y='cdf', title='initial growth rates (between t=2h and t=6h)') +
  # theme(strip.text=element_text(size=6)) +
  NULL

```

```{r fig.height=2}
mycells_btw %>% 
  ungroup %>% 
  filter(!is.na(medium)) %>% 
  # left_join(condition_ts %>% select(condition, medium, m_id) %>% distinct()) %>% 
  # mutate(m_col=interaction(m_id, m_cycle)) %>% 
  group_by(condition) %>% 
  mutate(m_col=as.numeric(fct_inorder(factor(medium))) %>% interaction(m_cycle)) %>% 
  filter(m_col=='1.1') %>% 
  filter(medium=='glycerol') %>% 
  filter(time_birth> 2*3600) %>% 
  # pull(condition) %>% table
  ggplot(aes(logl_time_slope/log(2) * 3600)) +
  facet_wrap(~condition+date, labeller=as_labeller(rename_conds), ncol=6) +
  stat_ecdf(col='grey40', data = ~filter(., condition=='glucose') %>% select(-condition, -date)) +
  stat_ecdf(col=qual_cols[1]) +
  # stat_ecdf(aes(group=gl_id), col=qual_cols[1], size=.2) +
  coord_cartesian(xlim=c(0, 1.5)) +
  labs(x='growth rate (doublings/h)', y='cdf', title='initial growth rates (t≥2h)') +
  # theme(strip.text=element_text(size=6)) +
  NULL

```

```{r}
mycells_btw %>% 
  filter(!is.na(medium)) %>% 
  # left_join(condition_ts %>% select(condition, medium, m_id) %>% distinct()) %>% 
  # mutate(m_col=interaction(m_id, m_cycle)) %>% 
  group_by(condition) %>% 
  mutate(m_col=as.numeric(fct_inorder(factor(medium))) %>% interaction(m_cycle)) %>% 
  filter(m_col=='1.1') %>% 
  ungroup %>% 
  filter(! condition %in% c('mg1655', 'lactose'),
         between(time_birth/3600, 2, 6),
         medium!='glycerol') %>% 
  mutate(type=ifelse(date %in% discarded_dates, 'discarded', 'analysed')) %>% 
  (function(.df) {
    ggplot(.df, aes(logl_time_slope/log(2) * 3600)) +
      facet_grid(type~.) +
      # facet_wrap(~condition+date, labeller=as_labeller(rename_conds), ncol=6) +
      stat_ecdf(aes(col=factor(date))) +
      stat_ecdf(col='grey40', lty='dashed',
                data=.df %>% filter(condition=='glucose') %>% select(-type, -condition, -date)) +
      scale_colour_periodic(.n = 3, guide='none') +
      coord_cartesian(xlim=c(0, 1.2)) +
      labs(x='growth rate (doublings/h)', y='cumulative probability') +
      # theme(strip.text=element_text(size=6)) +
      NULL
  }) +
  labs(subtitle='initial growth rates (between t=2h and t=6h)')


(myplots[['gr_length_medians']] <- function(.article_ds=FALSE)
  mycells_btw %>% ungroup() %>% 
    filter(!is.na(medium)) %>% 
    filter(!.article_ds | ( !str_detect(condition, "_stdIllum") &
      !condition %in% c("switch_ramp15min", "switch_lactose_priming") ) ) %>%
    group_by(condition) %>% 
    mutate(m_col=as.numeric(fct_inorder(factor(medium))) %>% interaction(m_cycle)) %>% 
    filter(m_col=='1.1') %>% 
    ungroup %>% 
    filter(! condition %in% c('mg1655', 'lactose'),
           (time_birth>2*3600 & time_birth<6*3600)) %>%
    left_join(condition_ts %>% filter(t_start==0) %>% 
                select(condition, nutrient=medium) ) %>% 
    mutate(condition=fct_relabel(factor(condition), rename_conds),
           type=ifelse(date %in% discarded_dates, 'discarded', 'analysed')) %>% 
           # nutrient='glucose', 
           # nutrient=ifelse(medium=='glycerol', 'glycerol', nutrient),
           # nutrient=ifelse(condition=='lactose', 'lactose', nutrient)) %>% 
    (function(.df) {
      .dfs <- .df %>% group_by(condition, nutrient, type, date) %>% 
        do(median_hilow(.$loglength_start, conf.int=0.5) %>%
        # do(median_pi(.$loglength_start) %>% 
             setNames(c('lls_med', 'lls_min', 'lls_max')) ) %>% 
        left_join(
          .df %>% group_by(date) %>% 
            do(median_hilow(.$logl_time_slope, conf.int=0.5) %>% 
            # do(median_pi(.$logl_time_slope) %>% 
                 setNames(c('lltsl_med', 'lltsl_min', 'lltsl_max')) )
        )
      ggplot(.dfs) +
        geom_point(aes(exp(lls_med), lltsl_med/log(2) * 3600, col=nutrient, alpha=type, shape=type)) +
        geom_errorbar(aes(exp(lls_med), ymin=lltsl_min/log(2) * 3600, ymax=lltsl_max/log(2) * 3600, col=nutrient, alpha=type)) +
        geom_errorbarh(aes(exp(lls_med), lltsl_med/log(2) * 3600, xmin=exp(lls_min), xmax=exp(lls_max), col=nutrient, alpha=type))+
        scale_alpha_manual(values=c(1, 0.2)) +
        coord_cartesian(xlim=c(1.4, 2.), ylim=c(0, 0.85)) +
        labs(x='birth length (µm)', y='growth rate (doublings/h)') +
        NULL
    })
)() +
  labs(subtitle='initial condition (between t=2h and t=6h)')


```


```{r growth_rate_medians, eval=FALSE}
mycells_btw %>% 
  ungroup %>% 
  filter(time_birth>2*3600, time_birth<6*3600) %>% 
  filter(!is.na(medium)) %>% 
  group_by(condition) %>% 
  mutate(m_col=as.numeric(fct_inorder(factor(medium))) %>% interaction(m_cycle)) %>% 
  filter(m_col=='1.1') %>% 
  ungroup %>% 
  mutate(ds=interaction(date, condition, sep="\n")) %>% 
  ggplot() +
  geom_rect(aes(ymin=ymin, ymax=ymax), xmin=-Inf, xmax=Inf, alpha=.3,
            data=mycells_btw %>% ungroup %>% filter(condition=='glucose') %>%
              filter(time_birth>2*3600, time_birth<6*3600) %>% with(log(2)/logl_time_slope / 60) %>% median_pi) +
  stat_summary(aes(ds, log(2)/logl_time_slope / 60), fun.data=median_pi, size=.25) +
  expand_limits(y=0) +
  labs(y='median doubling time (min)', title='initial growth rates (between t=2h and t=6h)') +
  theme(axis.title.x=element_blank(), axis.text.x=element_text(angle=45, hjust=1, vjust=1)) +
  NULL

```

The systematic relationship between growth rate and cell size at birth indicates that cells growing slower are in a different physiological state. 

```{r}
mycells_btw %>% 
  filter(!is.na(medium)) %>% 
  group_by(condition) %>% 
  mutate(m_col=as.numeric(fct_inorder(factor(medium))) %>% interaction(m_cycle)) %>% 
  filter(m_col=='1.1') %>% 
  ungroup %>% 
  filter((time_birth>2*3600 & time_birth<6*3600)) %>%
  mutate(condition=fct_relabel(condition, rename_conds)) %>% 
  (function(.df) {
    .dfs <- .df %>% group_by(condition, date) %>% 
      do(median_pi(.$loglength_start) %>% setNames(c('lls_med', 'lls_pi_min', 'lls_pi_max')) ) %>% 
      left_join(
        .df %>% group_by(date) %>% 
          do(median_pi(.$logl_time_slope) %>% setNames(c('lltsl_med', 'lltsl_pi_min', 'lltsl_pi_max')) )
        )
         
    ggplot(.df) +
      geom_point(aes(exp(lls_med), lltsl_med/log(2) * 3600, col=condition), data=.dfs ) +
      geom_errorbar(aes(exp(lls_med), ymin=lltsl_pi_min/log(2) * 3600, ymax=lltsl_pi_max/log(2) * 3600, col=condition), 
                    alpha=.5, data=.dfs ) +
      geom_errorbarh(aes(exp(lls_med), lltsl_med/log(2) * 3600, xmin=exp(lls_pi_min), xmax=exp(lls_pi_max), col=condition),
                     alpha=.5, data=.dfs ) +
      ggrepel::geom_text_repel(aes(exp(lls_med), lltsl_med/log(2) * 3600, col=condition, label=date), alpha=0.3, data=.dfs) +
                               # force=4, point.padding=1) +
      scale_colour_periodic(.n = 3, guide='none') +
      coord_cartesian(xlim=c(1.25, 2.25), ylim=c(0, 1)) +
      labs(x='birth length (µm)', y='growth rate (doublings/h)') +
      # ggplot2::scale_color_discrete(guide='none') +
      theme(legend.position='bottom') +
      NULL
  }) +
  labs(subtitle='initial condition (between t=2h and t=6h)')


```


To the contrary, the distribution of correlation coefficients for the exponential elongation fits are difficult to interpret in order to identify outlier experiments.

```{r fig.height=10}
mycells_btw %>% 
  filter(!is.na(medium)) %>% 
  group_by(condition) %>% 
  mutate(m_col=as.numeric(fct_inorder(factor(medium))) %>% interaction(m_cycle)) %>% 
  filter(m_col=='1.1') %>% 
  ungroup %>% 
  filter((time_birth>2*3600 & time_birth<6*3600)) %>% 
  filter(npoints>19) %>% 
  (function(.df) {
    ggplot(.df, aes(logl_time_r2)) +
      facet_wrap(~condition+date, labeller=as_labeller(rename_conds), ncol=6) +
      stat_ecdf(col='grey40', data=.df %>% filter(condition=='glucose') %>% select(-condition, -date)) +
      stat_ecdf(col=qual_cols[1], alpha=.6) +
      coord_cartesian(xlim=c(0.95, 1)) +
      labs(x='Squared Pearson correl. coeff.', title='initial growth curves (between t=2h and t=6h)') +
      NULL
  })

```

Based on slow growth after inoculation in M9 + 0.2% glucose, we discard the following dates: 20161021, 20170901, 20180313, 20151218, 20170108, 20160526. Discarding 20160526 and 20170901 is also supported by poor fit quality as indicated by their $r^2$ distribution.

<!-- TO DO
?? check correl with inoc OD + loading duration + loading ratio
-->


