---
title: "Growth-coupled sensitivity to lactose during E. coli growth on glucose-lactose mixture"
author: Thomas Julou
date: "`r format(Sys.time(), '%d %B, %Y')`"
---


```{r}
# utility function to parse conditions
parse_sugarmix_conds <- function(data, colname) {
  .groups <- groups(data)
  data %>% 
    ungroup() %>% 
    extract({{colname}}, "glu", "glu(\\d+)uM", remove=FALSE, convert=TRUE) %>% 
    mutate(
      glu = ifelse({{colname}}=='lac002', 0, glu),
      glu = ifelse(str_detect({{colname}}, 'glu002'), 1009, glu),
      glu = ifelse(str_detect({{colname}}, 'glu02'), 10092, glu),
      treatment = NA,
      treatment = ifelse(str_detect({{colname}}, '^glu\\d+'), 'none', treatment),
      treatment = ifelse(str_detect({{colname}}, 'lac002'), 'lac002', treatment),
      treatment = ifelse(str_detect({{colname}}, 'IPTG'), 'iptg', treatment),
      treatment = fct_relevel(factor(treatment), 'none', 'lac002', 'iptg'),
    ) %>% 
    group_by(!!!.groups)
}

```

 
```{r eval=FALSE}
# visualizing the position of cell bounding box has helped fixing various issues with data 
# import (due to the unusual acquisition scheme with skipped frames for fluo)
myframes %>% 
  filter(str_detect(condition, "(?:glu\\d+)|(?:lac\\d+)")) %>% 
  # filter(date == 20210122) %>% 
  # remove frames after touching the exit
  group_by(id, .add=TRUE) %>%
  mutate(discard_top=which_touch_exit(vertical_top, 1)) %>% 
  mutate(discard_top=ifelse(discard_start, FALSE, discard_top)) %>% # not in the preexpt step (2h)
  mutate(end_type=ifelse(any(discard_top), 'touchtop', end_type)) %>% # update end_type to exit for cells which have touched the vertical cutoff
  # remove daughters of cells that touched the exit
  group_by(path) %>%
  mutate(discard_top=which_to_progeny(discard_top, cid)) %>%
  group_by(cell) %>% nest() %>% ungroup() %>% slice_sample(n=20) %>% unnest(data) %>% 
  ggplot(aes(time_sec, length_um, col=cell)) +
  facet_wrap(~cell, scales='free') +
  # geom_hline(yintercept = .8) +
  # geom_line(aes(col=cell, group=cell)) +
  # geom_point() +
  geom_segment(aes(x=time_sec, xend=time_sec, y=vertical_bottom, yend=vertical_top, alpha=!discard_top)) +
  scale_x_hours(.5) +
  scale_colour_periodic(guide='none') +
  theme(legend.position = 'none') +
  NULL

```


```{r check raw traces, eval=FALSE}
myframes %>% 
  filter(str_detect(condition, "(?:glu\\d+)|(?:lac\\d+)")) %>% 
  # filter(date == 20210122) %>% 
  mutate(condition=fct_inorder(condition)) %>% 
  # remove frames after touching the exit
  group_by(id, .add=TRUE) %>%
  mutate(discard_top=which_touch_exit(vertical_top, 1)) %>%
  mutate(discard_top=ifelse(discard_start, FALSE, discard_top)) %>% # not in the preexpt step (2h)
  mutate(end_type=ifelse(any(discard_top), 'touchtop', end_type)) %>% # update end_type to exit for cells which have touched the vertical cutoff
  # remove daughters of cells that touched the exit
  group_by(path) %>%
  mutate(discard_top=which_to_progeny(discard_top, cid)) %>%
  filter(!discard_top) %>% 
  group_by(condition, date, cell) %>% nest() %>% 
  group_by(condition, date) %>% sample_n(50L) %>% unnest(data) %>% 
  ggplot(aes(time_sec, length_um)) +
  facet_wrap(~date+condition, ) +
  # geom_hline(yintercept = .8) +
  geom_line(aes(col=cell, group=cell), alpha=.4) +
  scale_x_hours() +
  scale_y_continuous(trans='log2') +
  scale_colour_periodic(guide='none') +
  NULL

myframes %>% 
  filter(str_detect(condition, "(?:glu\\d+)|(?:lac\\d+)")) %>% 
  # filter(date == 20210122) %>% 
  mutate(condition=fct_inorder(condition)) %>% 
  # remove frames after touching the exit
  group_by(id, .add=TRUE) %>%
  mutate(discard_top=which_touch_exit(vertical_top, 1)) %>%
  mutate(discard_top=ifelse(discard_start, FALSE, discard_top)) %>% # not in the preexpt step (2h)
  mutate(end_type=ifelse(any(discard_top), 'touchtop', end_type)) %>% # update end_type to exit for cells which have touched the vertical cutoff
  # remove daughters of cells that touched the exit
  group_by(path) %>%
  mutate(discard_top=which_to_progeny(discard_top, cid)) %>%
  filter(!discard_top, !is.na(gfp_nb)) %>% 
  group_by(condition, date, cell) %>% nest() %>% 
  group_by(condition, date) %>% sample_n(50L) %>% unnest(data) %>% 
  ggplot(aes(time_sec, gfp_nb)) +
  facet_wrap(~date+condition) +
  # geom_hline(yintercept = .8) +
  geom_line(aes(col=cell, group=cell), alpha=.4) +
  scale_x_hours() +
  scale_colour_periodic(guide='none') +
  NULL

```


```{r}
myframes %>% 
  filter(str_detect(condition, "(?:glu\\d+)|(?:lac\\d+)")) %>% 
  # filter(date == 20210122) %>% 
  mutate(condition=fct_inorder(condition)) %>% 
  # remove frames after touching the exit
  group_by(id, .add=TRUE) %>%
  mutate(discard_top=which_touch_exit(vertical_top, 1)) %>%
  mutate(discard_top=ifelse(discard_start, FALSE, discard_top)) %>% # not in the preexpt step (2h)
  mutate(end_type=ifelse(any(discard_top), 'touchtop', end_type)) %>% # update end_type to exit for cells which have touched the vertical cutoff
  # remove daughters of cells that touched the exit
  group_by(path) %>%
  mutate(discard_top=which_to_progeny(discard_top, cid)) %>%
  filter(!discard_top, !is.na(gfp_nb)) %>% 
  mutate(induced=gfp_nb>250) %>% 
  group_by(condition, date, cell) %>% nest() %>% 
  group_by(condition, date) %>% sample_n(50L) %>% unnest(data) %>% 
  parse_sugarmix_conds(condition) %>% 
  ggplot(aes(time_sec, length_um, col=gfp_nb/length_um)) +
  # facet_wrap(~date+condition, labeller = as_labeller(label_value, multi_line=FALSE)) +
  facet_wrap(~ treatment + glu + date, labeller = as_labeller(label_value, multi_line=FALSE)) +
  # facet_wrap(~condition + induced) +
  geom_vline(xintercept = 4*3600, lty='dotted') +
  geom_line(aes(group=cell), alpha=.4) +
  scale_x_hours() +
  scale_y_continuous(limits=c(.8, 6), trans='log2') +
  scale_color_viridis_c(limits=c(NA, 1000), oob=scales::squish) +
  labs(y="cell length (µm)", col="[LacZ-GFP]") +
  NULL
# ggsave("plots/sugar_mixtures_traces.pdf", width=9, height=5)

```


## Analysing growth rates of induced and uninduced cells in each mixture

This analysis focuses on growth rates computed over full cell cycles.
We consider all traces with 4 observations or more, after discarding the first 4 hours (where [LacZ-GFP] obviously adjusts due to environmental changes compared to preculture conditions).

```{r cache=TRUE}
myrates_mix_all <- myframes %>% 
  filter(str_detect(condition, "(?:glu\\d+)|(?:lac\\d+)")) %>% 
  # filter(date == 20210122) %>% 
  mutate(condition=fct_inorder(condition)) %>%
  # filter(! (date==20210122 & condition %in% c('lac002_glu4uM', 'lac002_glu16uM'))) %>% 
  # remove frames after touching the exit
  group_by(id, .add=TRUE) %>%
  mutate(discard_top=which_touch_exit(vertical_top, 1)) %>%
  mutate(discard_top=ifelse(discard_start, FALSE, discard_top)) %>% # not in the preexpt step (2h)
  mutate(end_type=ifelse(any(discard_top), 'touchtop', end_type)) %>% # update end_type to exit for cells which have touched the vertical cutoff
  # remove daughters of cells that touched the exit
  group_by(path) %>%
  mutate(discard_top=which_to_progeny(discard_top, cid)) %>%
  filter(!discard_top) %>% 
  # call cell induced as soon as it crossed the 250 threshold
  mutate(gfp_nb_raw=gfp_nb) %>% 
  fill(gfp_nb, .direction="downup") %>% 
  mutate(induced=gfp_nb>250) %>% 
  # compute growth rate over cell cycles
  filter(time_sec > 4 * 3600) %>% 
  group_by(date, condition, cell) %>% 
  filter(n()>3) %>% 
  select(time_sec, length_um, gfp_nb, induced, b_rank) %>% 
  nest() %>% 
  group_by(date, condition) %>% 
  partition(mycluster) %>% 
  mutate(
    n=map_dbl(data, ~nrow(.x)),
    induced=map_dbl(data, ~mean(.x$induced)),
    mod = map(data, ~RcppArmadillo::fastLm( cbind(1, .x$time_sec), log(.x$length_um))),
    grate = map_dbl(mod, ~.$coef[2] ),
    r2 = map_dbl(data, ~cor(.x$time_sec, log(.x$length_um))^2),
    mod = NULL
  ) %>% 
  collect()

```


We first establish an induction criteria. Computing the proportion of observation where [LacZ-GFP] > 250 yields clearly bimodal distribution with cells either induced or not, so one can reasonably stratify by this variable.

Note that while it could look appealing to stratify by the expression state of the *lac* operon (i.e. whether a cell has produced Lac proteins during the current cell cycle), what actually matters for growing on lactose is the current concentration on Lac proteins, hence the criteria used here.

```{r}
(myplots[['sugarmix_indn_thr']] <- myframes %>% 
  ungroup() %>% 
  filter(str_detect(condition, "(?:glu\\d+)|(?:lac\\d+)")) %>% 
  parse_sugarmix_conds(condition) %>% 
  filter(treatment == 'lac002', glu %in% c(0, 16, 32, 64, 128)) %>%
  filter(!(date == 20210122 & glu==16) ) %>%
  # remove frames after touching the exit
  group_by(date, pos, gl, id) %>%
  mutate(discard_top=which_touch_exit(vertical_top, 1)) %>%
  mutate(discard_top=ifelse(discard_start, FALSE, discard_top)) %>% # not in the preexpt step (2h)
  mutate(end_type=ifelse(any(discard_top), 'touchtop', end_type)) %>% # update end_type to exit for cells which have touched the vertical cutoff
  # remove daughters of cells that touched the exit
  group_by(path) %>%
  mutate(discard_top=which_to_progeny(discard_top, cid)) %>%
  filter(!discard_top) %>% 
  # no need to fill up gfp_nb NAs
  filter(time_sec > 4 * 3600) %>% 
  # group_by(date, glu) %>% summarise(n=n(), nn=sum(!is.na(gfp_nb)))
  ggplot(aes(gfp_nb, col=glu, group=interaction(date, glu))) +
  stat_ecdf() +
  coord_cartesian(xlim=c(NA, 6000)) +
  geom_vline(xintercept = 250, lty='dotted') +
  scale_color_viridis_c(trans='log2') +
  labs(x='total LacZ-GFP per cell', y="cumul. proba.", col="[glucose] (µM)") +
  NULL)

myframes %>% 
  ungroup() %>% 
  filter(str_detect(condition, "(?:glu\\d+)|(?:lac\\d+)")) %>% 
  parse_sugarmix_conds(condition) %>% 
  filter(treatment == 'lac002', glu %in% c(0, 16, 32, 64, 128)) %>%
  filter(!(date == 20210122 & glu==16) ) %>%
  # remove frames after touching the exit
  group_by(date, pos, gl, id) %>%
  mutate(discard_top=which_touch_exit(vertical_top, 1)) %>%
  mutate(discard_top=ifelse(discard_start, FALSE, discard_top)) %>% # not in the preexpt step (2h)
  mutate(end_type=ifelse(any(discard_top), 'touchtop', end_type)) %>% # update end_type to exit for cells which have touched the vertical cutoff
  # remove daughters of cells that touched the exit
  group_by(path) %>%
  mutate(discard_top=which_to_progeny(discard_top, cid)) %>%
  filter(!discard_top) %>% 
  # no need to fill up gfp_nb NAs
  filter(time_sec > 4 * 3600) %>% 
  # group_by(date, glu) %>% summarise(n=n(), nn=sum(!is.na(gfp_nb)))
  ggplot(aes(gfp_nb, col=factor(date))) +
  facet_wrap(~glu) +
  geom_freqpoly(aes(y=pmax(after_stat(density), 1e-5)), binwidth=50, boundary=0) +
  geom_vline(xintercept = 250, lty='dotted') +
  # coord_cartesian(xlim=c(NA, 7000), ylim=c(NA, 5e-4)) +
  coord_cartesian(xlim=c(NA, 7000)) +
  scale_y_log10() +
  guides(col='none') +
  NULL

```


```{r}
myrates_mix_all %>% 
  parse_sugarmix_conds(condition) %>% 
  ggplot(aes(induced, col=factor(date))) +
  # facet_wrap(~date, scales='free') +
  facet_grid(glu~treatment) +
  # geom_freqpoly(aes(y=after_stat(density)))
  geom_freqpoly(aes(y=after_stat(density))) +
  xlim(0, 1)
```

```{r eval=FALSE}
# Let us check that our dataset is not dominated by very short traces.
# This actually reveals an unexpected change of traces length when adding IPTG (to be looked at in more details).
myrates_mix_all %>% 
  parse_sugarmix_conds(condition) %>% 
  ggplot(aes(n, col=factor(date), lty=factor(round(induced)))) +
  facet_grid(glu~treatment) +
  # geom_freqpoly(aes(y=after_stat(density)))
  stat_ecdf(aes(y=1-after_stat(y))) +
  coord_cartesian(xlim=c(0, 200)) + 
  NULL

# Let's check the distribution of r2... should maybe keep traces with 10 points or more
myrates_mix_all %>% 
  parse_sugarmix_conds(condition) %>% 
  ggplot(aes(r2, col=factor(date), lty=factor(round(induced)))) +
  facet_grid(glu~treatment) +
  # geom_freqpoly(aes(y=after_stat(density)))
  stat_ecdf() +
  # coord_cartesian(xlim=c(0, 200)) + 
  NULL

myrates_mix_all %>% 
  parse_sugarmix_conds(condition) %>% 
  ggplot(aes(n, r2, col=factor(date))) +
  facet_grid(glu~treatment) +
  geom_vline(xintercept=10, lty='dotted') +
  geom_point(stroke=0, alpha=.2, size=1) +
  coord_cartesian(xlim=c(0, 100)) +
  NULL
```


### Time windows for steady state of induction

We first examine how the proportion of induced cells changes during the experiment in order to study it after the transient.

```{r}
# plot prop_induced vs conc
myrates_mix_all %>% 
  # filter(n>10) %>% 
  mutate(time_birth=map_dbl(data, ~min(.$time_sec))) %>% 
  parse_sugarmix_conds(condition) %>% 
  ggplot(aes(time_birth, induced)) +
  facet_grid(glu~treatment) +
  # geom_hline(yintercept=c(0, 1), col='gray80') +
  geom_vline(xintercept = 5 * 3600, lty='dotted') +
  geom_point(stroke=0, alpha=.2, size=1, position = position_jitter(height = .1)) +
  scale_x_hours() +
  theme(panel.background = element_rect(fill='gray93')) +
  NULL

myrates_mix_all %>% 
  filter(n>10) %>% 
  filter(induced<.05 | induced>.95) %>% 
  mutate(induced=round(induced)) %>% 
  group_by(date, condition) %>% 
  mutate(time_birth=map_dbl(data, ~min(.$time_sec)),
         time_bin = cut_width(time_birth, width = 3600, boundary = 0, labels=FALSE)) %>% 
  group_by(date, condition, time_bin) %>% 
  summarise(n=n(), prop=sum(induced)/n) %>% 
  parse_sugarmix_conds(condition) %>% 
  ggplot(aes(time_bin, prop, col=factor(date))) +
  facet_grid(glu~treatment) +
  # geom_hline(yintercept=c(0, 1), col='gray80') +
  # geom_vline(xintercept = 5 * 3600, lty='dotted') +
  geom_point(aes(size=n)) +
  geom_line() +
  theme(panel.background = element_rect(fill='gray93')) +
  NULL
```

```{r eval=FALSE}
# Same analysis considering all observations (rather than one summary statistics per cell cycle)
myframes %>% 
  filter(str_detect(condition, "(?:glu\\d+)|(?:lac\\d+)")) %>% 
  # remove frames after touching the exit
  group_by(id, .add=TRUE) %>%
  mutate(discard_top=which_touch_exit(vertical_top, 1)) %>%
  mutate(discard_top=ifelse(discard_start, FALSE, discard_top)) %>% # not in the preexpt step (2h)
  mutate(end_type=ifelse(any(discard_top), 'touchtop', end_type)) %>% # update end_type to exit for cells which have touched the vertical cutoff
  # remove daughters of cells that touched the exit
  group_by(path) %>%
  mutate(discard_top=which_to_progeny(discard_top, cid)) %>%
  filter(!discard_top) %>% 
  # call cell induced as soon as it crossed the 250 threshold
  mutate(gfp_nb_raw=gfp_nb) %>% 
  fill(gfp_nb, .direction="downup") %>% 
  mutate(induced=gfp_nb>250) %>% 
  group_by(date, condition, time_sec) %>% 
  summarise(prop_induced=sum(induced)/n(), n=n()) %>% 
  group_by(date, condition) %>% 
  mutate(prop_cells=n/max(n)) %>% 
  filter(cumall(n>10)) %>% 
  parse_sugarmix_conds(condition) %>% 
  ggplot(aes(time_sec, prop_induced)) +
  facet_grid(glu~treatment) +
  geom_line(aes(col=factor(date))) +
  scale_x_hours() +
  theme(panel.background = element_rect(fill='gray93')) +
  NULL

myframes %>% 
  ungroup() %>% 
  # filter(date==20210504 & str_detect(condition, "lac002") | 
  #          date==20210506 & condition=="lac002_glu64uM") %>% 
  filter(time_sec < 30 * 3600) %>% 
# 
    filter(str_detect(condition, "(?:glu\\d+)|(?:lac\\d+)")) %>% 
  # remove frames after touching the exit
  group_by(date, pos, gl, id, .add=TRUE) %>%
  mutate(discard_top=which_touch_exit(vertical_top, 1)) %>%
  mutate(discard_top=ifelse(discard_start, FALSE, discard_top)) %>% # not in the preexpt step (2h)
  mutate(end_type=ifelse(any(discard_top), 'touchtop', end_type)) %>% # update end_type to exit for cells which have touched the vertical cutoff
  # remove daughters of cells that touched the exit
  group_by(path) %>%
  mutate(discard_top=which_to_progeny(discard_top, cid)) %>%
  filter(!discard_top) %>% 
  # # call cell induced as soon as it crossed the 250 threshold
  # mutate(gfp_nb_raw=gfp_nb) %>% 
  # fill(gfp_nb, .direction="downup") %>% 
  filter(!is.na(gfp_nb)) %>% 
  mutate(induced=gfp_nb>250) %>% 
  group_by(date, condition, pos, gl, time_sec) %>% 
  nest() %>% 
  mutate(n=map_dbl(data, ~nrow(.)),
         prop_induced = map_dbl(data, ~sum(.$induced)/nrow(.)),
         # b_induced = map_dbl(data, filter(., b_rank==0)$induced == TRUE),
         ) %>% 
  # group_by(date, condition, pos, gl) %>% 
  # filter(cumall(n>3)) %>% 
  parse_sugarmix_conds(condition) %>%
  ggplot(aes(time_sec, prop_induced, col=factor(date), group=interaction(condition, date, pos, gl))) +
  facet_grid(glu~treatment, space="free_x") +
  geom_line(aes(), alpha=.4) +
  scale_x_hours() +
  theme(panel.background = element_rect(fill='gray93')) +
  NULL

```

We note that the proportion of induced cells stabilises typically after 5 hours. We assume that this is the time needed to adjust to the new condition and filter out previous observations also for the glucose-only and iptg datasets.
Let compute rates for the relevant time window.

```{r}
myrates_mix <- myrates_mix_all %>%
  select(condition, date, cell, data) %>%
  unnest(data) %>%
  filter(time_sec > 5 * 3600,
         ! (condition %in% c("lac002_glu32uM", "lac002_glu64uM") & time_sec < 14 * 3600) ) %>%
  filter(!(date==20210122 & condition=="lac002_glu16uM")) %>%
  group_by(date, condition, cell) %>%
  filter(n()>3) %>%
  nest() %>%
  group_by(date, condition) %>%
  partition(mycluster) %>%
  mutate(
    n=map_dbl(data, ~nrow(.x)),
    induced=map_dbl(data, ~mean(.x$induced)),
    mod = map(data, ~RcppArmadillo::fastLm( cbind(1, .x$time_sec), log(.x$length_um))),
    grate = map_dbl(mod, ~.$coef[2] ),
    r2 = map_dbl(data, ~cor(.x$time_sec, log(.x$length_um))^2),
    mod = NULL
  ) %>%
  collect() %>%
  identity()

## alternative: filtering rather than recomputing (at the cost of discarding the initial traces)
# myrates_mix <- myrates_mix_all %>% 
#   ungroup() %>% 
#   mutate(
#     time_start=map_dbl(data, ~min(.$time_sec)),
#     discard=ifelse(condition %in% c("lac002_glu32uM", "lac002_glu64uM"), 
#                    time_start < 14 * 3600,
#                    time_start < 5 * 3600),
#     time_start=NULL
#     ) %>% 
#   filter(!(date==20210122 & condition=="lac002_glu16uM"))
# 

myrates_mix %>% 
  ungroup() %>% 
  filter(n>10) %>% 
  mutate(type=induced<.05 | induced>.95) %>% 
  summarise(t=sum(type), n=n(), p=1-t/n)

myrates_mix <- myrates_mix %>%
  filter(n>10) %>%
  filter(induced<.05 | induced>.95) %>%
  mutate(induced=round(induced))

```


### Growth rate distributions

Let's check growth rate distributions as a function of glucose concentration for different treatments.

```{r}
(myplots[['sugarmix_gr_distr']] <- 
   myrates_mix %>% 
   group_by(date, condition) %>% 
   mutate(prop = ifelse(induced, sum(induced)/n(), 1-sum(induced)/n())) %>% 
   parse_sugarmix_conds(condition) %>% 
   ggplot(aes(grate/log(2)*3600, col=factor(round(induced)))) +
   facet_grid(glu~treatment, labeller=labeller(
     treatment=as_labeller(function(.l) c('none'='glucose only', 'lac002'='+ lactose 0.55 mM', 'iptg'='+ IPTG 200µM')[.l] ),
     glu=as_labeller(function(.l) paste(.l, 'µM') )
   ) ) +
   geom_freqpoly(aes(y=after_stat(density), size=prop, group=interaction(date, condition, prop)), binwidth=.1) +
   scale_y_continuous(breaks = scales::breaks_extended(n=3)) +
   coord_cartesian(xlim=c(-.25, 1.5), ylim=c(0, 3.5)) +
   scale_size_continuous(range=c(0.05, 1), guide='none') +
   scale_colour_discrete(labels=c('0'='uninduced', '1'='induced')) +
   labs(x='growth rate (dbl/h)', col='subpopulation') +
   theme(legend.position = 'top') +
   NULL)

```


### Growth controls

For glucose only, let's verify that the growth rate does not depend on the position in the channel.

```{r}
(myplots[['sugarmix_gr_pos']] <- 
   myrates_mix %>% 
   ungroup() %>% 
   parse_sugarmix_conds(condition) %>% 
   filter(treatment %in% c('none')) %>% 
   mutate(date=paste('day', as.numeric(factor(date))) ) %>% 
   group_by(date, condition) %>% 
   mutate(prop = ifelse(induced, sum(induced)/n(), 1-sum(induced)/n())) %>% 
   left_join(myframes %>% group_by(date, cell) %>% )
   mutate(b_rank=map_dbl(data, ~mean(.$b_rank)) ) %>% 
   filter(b_rank < 7) %>% 
   ggplot(aes(grate/log(2)*3600, col=b_rank)) +
   facet_wrap(~glu+date, nrow=3,
              labeller=labeller(glu=function(.l) paste(.l, "µM") ) 
              # labeller=labeller(glu=function(.l) sprintf('%.2g', as.numeric(.l)) ) 
              ) +
   stat_ecdf(aes(group=interaction(date, condition, b_rank))) +
   scale_y_continuous(breaks = c(0, 0.5, 1)) +
   scale_color_viridis_c() +
   coord_cartesian(xlim=c(-.25, 1.5)) +
   labs(x='growth rate (dbl/h)', y='cumulative probability', col='position') +
   # theme(legend.position = 'top') +
   NULL)

```


```{r eval=FALSE}
myframes %>%
  filter(condition=="M9zero") %>%
  # remove frames after touching the exit
  group_by(id, .add=TRUE) %>%
  mutate(discard_top=which_touch_exit(vertical_top, 1)) %>%
  mutate(discard_top=ifelse(discard_start, FALSE, discard_top)) %>% # not in the preexpt step (2h)
  mutate(end_type=ifelse(any(discard_top), 'touchtop', end_type)) %>% # update end_type to exit for cells which have touched the vertical cutoff
  # remove daughters of cells that touched the exit
  group_by(path) %>%
  mutate(discard_top=which_to_progeny(discard_top, cid)) %>%
  filter(!discard_top) %>%
  # # call cell induced as soon as it crossed the 250 threshold
  # mutate(gfp_nb_raw=gfp_nb) %>%
  # fill(gfp_nb, .direction="downup") %>%
  # mutate(induced=gfp_nb>250) %>%
  group_by(cell) %>%
  nest() %>% ungroup() %>% sample_n(10) %>%
  unnest(data) %>%
  ggplot(aes(time_sec, length_um, col=cell)) +
  geom_line() +
  scale_x_hours() +
  scale_colour_periodic() +
  guides(col='none') +
  NULL
```

Let's plot the instantaneous growth rate in M9 zero averaged in 1 h bins in order to check how much growth is possible without sugar.

```{r}
(myplots[['sugarmix_m9zero']] <- 
   myframes %>% 
   filter(condition=="M9zero") %>% 
   # remove frames after touching the exit
   group_by(id, .add=TRUE) %>%
   mutate(discard_top=which_touch_exit(vertical_top, 1)) %>%
   mutate(discard_top=ifelse(discard_start, FALSE, discard_top)) %>% # not in the preexpt step (2h)
   mutate(end_type=ifelse(any(discard_top), 'touchtop', end_type)) %>% # update end_type to exit for cells which have touched the vertical cutoff
   # remove daughters of cells that touched the exit
   group_by(path) %>%
   mutate(discard_top=which_to_progeny(discard_top, cid)) %>%
   filter(!discard_top) %>% 
   mutate(time_bin = cut_width(time_sec, width = 3600, boundary = 0, labels=FALSE)) %>% 
   group_by(cell, time_bin) %>% 
   filter(n() > 5) %>% 
   nest() %>% 
   mutate(mod=map(data, ~lm(log(length_um)~time_sec, data=.)),
          grate=map_dbl(mod, ~coef(.)[2]),
          grate2=grate*3600/log(2)) %>% 
   group_by(time_bin) %>% 
   summarise(grate=mean(grate2), grate_sd=sd(grate2), grate_se=sd(grate2)/sqrt(n())) %>% 
   ggplot(aes(time_bin-.5, grate)) +
   geom_hline(yintercept=0, lty="dotted") +
   geom_pointrange(aes(ymin=grate-grate_sd, ymax=grate+grate_sd)) +
   expand_limits(y=.5) +
   labs(x="time (h)", y="growth rate (dbl/h)") +
   NULL)

```


## Proportion of induced cells

Let us compute the proportion of induced cells as a function of glucose concentration. 

```{r}
# # pooling all observations after the transient
# myrates_mix_all %>% 
#   filter(n>10) %>% 
#   filter(induced<.05 | induced>.95) %>% 
#   mutate(induced=round(induced)) %>% 
#   group_by(date, condition) %>% 
#   mutate(time_birth=map_dbl(data, ~min(.$time_sec))) %>% 
#   filter(time_birth > 5 * 3600,
#           ! (condition %in% c("lac002_glu32uM", "lac002_glu64uM") & time_birth < 14 * 3600) ) %>% 
#   summarise(n=n(), prop=sum(induced)/n, se=sqrt(prop*(1-prop)/n)) %>% 
#   parse_sugarmix_conds(condition) %>% 
#   filter(treatment=='lac002') %>% 
#   ggplot(aes(glu, prop, col=factor(date))) +
#   # geom_pointrange(aes(ymin=prop-se, ymax=prop+se)) +
#   geom_pointrange(aes(ymin=prop-se, ymax=prop+se), 
#                   position=position_dodge(width=.2), data=~filter(., glu>0)) +
#   scale_x_continuous(trans='log2') +
#   NULL

# estimating uncertainty as sd over time windows of 1h
myrates_mix %>% 
  group_by(date, condition) %>% 
  mutate(time_birth=map_dbl(data, ~min(.$time_sec)),
         time_bin = cut_width(time_birth, width = 3600, boundary = 0, labels=FALSE)) %>% 
  group_by(date, condition, time_bin) %>% 
  summarise(n=n(), p=sum(induced)/n) %>% 
  group_by(date, condition) %>% 
  summarise(n=mean(n), prop=mean(p), se=sd(p)) %>% 
  parse_sugarmix_conds(condition) %>% 
  filter(treatment=='lac002') %>% 
  ggplot(aes(glu, prop, col=factor(date))) +
  # geom_pointrange(aes(ymin=prop-se, ymax=prop+se)) +
  geom_pointrange(aes(ymin=prop-se, ymax=prop+se), 
                  position=position_dodge(width=.2), data=~filter(., glu>0)) +
  scale_x_continuous(trans='log2') +
  NULL

```

Fit a Hill function:

```{r}
# myrates_mix %>% 
#   group_by(date, condition) %>% 
#   mutate(time_birth=map_dbl(data, ~min(.$time_sec))) %>% 
#   filter(time_birth > 5 * 3600,
#           ! (condition %in% c("lac002_glu32uM", "lac002_glu64uM") & time_birth < 14 * 3600) ) %>% 
#   summarise(n=n(), prop=sum(induced)/n, se=sqrt(prop*(1-prop)/n)) %>% 
#   parse_sugarmix_conds(condition) %>% 
#   filter(treatment=='lac002') %>% 
#   ggplot(aes(glu, prop)) +
#   # geom_pointrange(aes(ymin=prop-se, ymax=prop+se)) +
#   geom_pointrange(aes(ymin=prop-se, ymax=prop+se, col=factor(date)), 
#                   position=position_dodge(width=.2), data=~filter(., glu>0)) +
#   geom_pointrange(aes(ymin=prop-se, ymax=prop+se, col=factor(date)), 
#                   position=position_jitter(width=.2, height = 0), data=~filter(., glu==0)) +
#   #stat_function(fun=hill_fn, args=list(k=68, m=41), data=NULL) +
#   scale_x_continuous(trans='log2') +
#   # xlim(0, 200) +
#   NULL

# set.seed(881676)
myindct <- myrates_mix %>% 
  group_by(date, condition) %>% 
  mutate(time_birth=map_dbl(data, ~min(.$time_sec)),
         time_bin = cut_width(time_birth, width = 3600, boundary = 0, labels=FALSE)) %>% 
  group_by(date, condition, time_bin) %>% 
  summarise(n=n(), p=sum(induced)/n) %>% 
  group_by(date, condition) %>% 
  summarise(n=mean(n), prop=mean(p), se=sd(p)) %>% 
  parse_sugarmix_conds(condition) %>% 
  filter(treatment=='lac002') %>% 
  group_by(treatment) %>% nest() %>%
  mutate(
    data_fit = map(data, ~filter(.x, glu > 0, !(date==20210122 & glu==16))),
    mod=map(data_fit, function(.d, .n=20, # run nls() .n times to keep the fit with lowest rss
                           .lower=list(k=log(5), m=2),
                           .upper=list(k=log(200), m=100) ) {
      .hill_fn <- function(c, k, m)  1 / (1 + (c/k)^m)
      .mods <- replicate(.n, simplify=FALSE, expr=try(silent=TRUE, expr=nls(
        prop~.hill_fn(log(glu), k, m), data=.d, 
        # weigths=log(activity+1),
        algorithm='port', control=nls.control(maxiter = 1e3, minFactor = 1/2^15), 
        start=map2(.lower, .upper, ~runif(1, .x, .y)), lower=.lower, upper=.upper,
      )) )
      .rss <- map_dbl(.mods, ~{ if (class(.x) == "try-error") NA else .x$m$deviance() })
      if (all(is.na(.rss))) return(.mods[[1]]) else return(.mods[[which.min(.rss)]])
    }),
    fitted=map_chr(mod, ~class(.x)),
  ) %>% 
  filter(fitted != 'try-error') %>% 
  mutate(
    # coef = map(mod, ~ as_tibble(t(coef(.x)))),
    coef = map(mod, ~ summary(.x)$coef[,1:2] %>% t() %>% as_tibble() %>% mutate(var=c("", "_se")) %>% 
                 pivot_wider(names_from=var, values_from=-var, names_sep="") ),
    prediction=list(tibble(glu=2^seq(-5, 11, length.out = 201))),
    prediction=map2(mod, prediction, function(.m, .d) 
      mutate(.d, prop=predict(.m, .d), prop=ifelse(is.nan(prop), 1, prop))),
  ) %>% 
  identity()

myindct %>% 
  ggplot( ) +
  geom_line(aes(glu, prop),
            data = ~select(., prediction) %>% unnest(prediction)) +
  geom_pointrange(aes(glu, prop, ymin=prop-se, ymax=prop+se, col=factor(date)),
                  position=position_dodge(width=.2), 
                  data = ~select(., data) %>% unnest(data) %>% filter(glu>0)) +
  geom_pointrange(aes(glu, prop, ymin=prop-se, ymax=prop+se, col=factor(date)),
                  position=position_jitter(width=.2, height = 0), 
                  data = ~select(., data) %>% unnest(data) %>% filter(glu==0)) +
  scale_x_continuous(trans='log2') +
  coord_cartesian(xlim=c(1.5, NA)) +
  NULL

```

## Growth rate dependence to glucose concentration

Summarizing growth rate dependence to glucose concentration.
<!-- Mind a bug of geom_pointrange() with x=0, position=dodge and trans=log -->

```{r}
myrates_mix %>% 
  group_by(date, condition) %>% 
  filter(between(grate, quantile(grate, .01, na.rm=TRUE), quantile(grate, .99, na.rm=TRUE) )) %>%
  group_by(date, condition, induced) %>% 
  summarise(
    gr_mean=mean(grate/log(2)*3600),
    gr_sd  =sd(grate/log(2)*3600),
    n=n(),
  ) %>% 
  group_by(date, condition) %>% 
  mutate(prop=n/sum(n)) %>% 
  parse_sugarmix_conds(condition) %>% 
  ggplot(aes(glu, gr_mean, col=factor(induced))) +
  facet_wrap(~treatment, ncol=1) +
  annotate("rect", xmin = -Inf, xmax = 2.5, ymin = -Inf, ymax = Inf, alpha = .2, fill='gray70') +
  geom_hline(aes(yintercept = gr_mean, col=NULL), 
             data=~filter(., treatment=='lac002', glu==0, induced==1) %>% 
               ungroup() %>% summarise(gr_mean=mean(gr_mean)) ) +
  # must filter glu>0 with position=dodge
  geom_pointrange(aes(ymin=gr_mean-gr_sd, ymax=gr_mean+gr_sd, size=prop),
                  position=position_dodge2(width=.4), 
                  data=~filter(., glu > 0) ) +
  coord_cartesian(ylim=c(NA, 1.4)) +
  scale_size_continuous(range=c(0.05, .5), guide='none') +
  scale_x_continuous(trans='log2') +
  labs(y='growth rate (dbl/h)') +
  scale_colour_discrete(name='', labels=c('0'='uninduced', '1'='induced')) +
  theme_cowplot() +
  theme(legend.position = c(0, 1), legend.justification = c(0, 0.7)) +
  NULL

```

```{r eval=FALSE}
# checking what fraction of induced cells look like in log scale
myindct %>% 
  ggplot( ) +
  geom_hline(yintercept = c(0, 1), lty='dotted', col='gray40') +
  geom_line(aes(glu, prop), 
            data = ~select(., prediction) %>% unnest(prediction) %>% mutate(prop=ifelse(prop<1e-3, 0, prop)) ) +
  # must filter glu>0 with position=dodge (already done in data_fit)
  geom_pointrange(aes(glu, prop, ymin=pmax(0, prop-se), ymax=prop+se, group=date, col='induced'),
                  position=position_dodge(width=.3), 
                  data = ~select(., data_fit) %>% unnest(data_fit) ) +
  geom_pointrange(aes(glu, 1-prop, ymin=1-(prop-se), ymax=pmax(1-(prop+se), 0), group=date, col='uninduced'),
                  position=position_dodge(width=.3), 
                  data = ~select(., data_fit) %>% unnest(data_fit) ) +
  scale_x_continuous(trans='log2') +
  scale_y_log10() +
  coord_cartesian(xlim=c(1.5, NA), ylim=c(5e-3, 1)) +
  theme_cowplot() +
  scale_colour_manual(name="", values=c('uninduced'=qual_cols[1], "induced"=qual_cols[2])) +
  theme(legend.position = "top") +
  NULL

```


```{r publication plot}
# 1%  ~ 10 g/L
# 1µM ~ 1.8e-4 g/L ~ 0.18 mg/L
(myplots[['sugarmix_induction']] <- function(.xbreaks=2^c(0, 4, 8)) {
  .xmin <- min(1, .xbreaks/1.35)
  .gr_pl <-
    myrates_mix %>% 
    group_by(date, condition) %>% 
    filter(between(grate, quantile(grate, .01, na.rm=TRUE), quantile(grate, .99, na.rm=TRUE) )) %>%
    group_by(condition, induced) %>% 
    summarise(
      gr_mean=mean(grate/log(2)*3600),
      gr_sd  =sd(grate/log(2)*3600),
      n=n(),
    ) %>% 
    group_by(condition) %>% 
    mutate(prop=n/sum(n)) %>% 
    parse_sugarmix_conds(condition) %>% 
    filter(treatment != "iptg") %>%
    ggplot(aes(glu, gr_mean)) +
    # geom_rect(aes(xmin=0, xmax=Inf, ymin=gr_mean-gr_sd, ymax=gr_mean+gr_sd, col=NULL), 
    #           fill="gray92", show.legend=FALSE,
    #           data=~filter(., treatment=='lac002', glu==0, induced==1) ) +
    geom_hline(aes(yintercept = gr_mean, col=NULL), col="gray40", lty="dotted",
               data=~filter(., treatment=='lac002', glu==0, induced==1) ) +
    # must filter glu>0 with position=dodge
    geom_point(aes(col=interaction(treatment, induced)), size=2.5, 
                    position=position_dodge(width=.5, preserve="total"), 
                    data=~filter(., glu > 0) ) +
    geom_linerange(aes(ymin=gr_mean-gr_sd, ymax=gr_mean+gr_sd, col=interaction(treatment, induced)),
                    position=position_dodge(width=.5, preserve="total"), show.legend=FALSE,
                    data=~filter(., glu > 0) ) +
    # adding lactose dot at zero glucose
    geom_pointrange(aes(x=.xmin, ymin=gr_mean-gr_sd, ymax=gr_mean+gr_sd, col='lac'),
                    show.legend=FALSE,
                    data=~filter(., glu == 0, induced==1) ) +
    # display legend without errorbars
    # geom_point(aes(fill=factor(induced)), show.legend=TRUE, size=5) +
    coord_cartesian(xlim=c(1.35*.xmin, NA), ylim=c(NA, 1.4)) +
    scale_x_continuous(trans='log2', breaks=2^c(0, 4, 8)) +
    scale_y_continuous(breaks=c(0, 0.5, 1)) +
    labs(x='[glucose] (µM)', y='growth rate (dbl/h)') +
    scale_colour_manual(name='', 
      limits = c('none.0', 'lac002.0', 'lac002.1', 'lac'), # set colours order
      values=c('none.0' = colorspace::desaturate( colorspace::lighten(qual_cols[1], .5), .2), 
               # 'none.0' = qual_cols[1], 
               'lac002.0' = qual_cols[1], 
               'lac002.1' = qual_cols[2],
               'lac' = colorspace::desaturate( colorspace::lighten(qual_cols[2], .5), .2) ),
      labels=c('none.0'='glucose only', 
               'lac002.0'='glucose + lactose 0.55 mM (uninduced)', 
               'lac002.1'='glucose + lactose 0.55 mM (induced)',
               'lac'='lactose 0.55 mM only') 
      ) +
    # theme_cowplot() +
    theme(legend.position = "bottom", legend.direction = "vertical", 
          legend.box.just="left") +
    NULL
  
  .prop_pl <-
    myindct %>% 
    ggplot( ) +
    geom_hline(yintercept = c(0, 1), lty='dotted', col='gray40') +
    geom_line(aes(glu, prop),
              data = ~select(., prediction) %>% unnest(prediction)) +
    # must filter glu>0 with position=dodge (already done in data_fit)
    geom_pointrange(aes(glu, prop, ymin=prop-se, ymax=prop+se, group=date),
                    position=position_dodge(width=.3), 
                    data = ~select(., data_fit) %>% unnest(data_fit) ) +
    geom_pointrange(aes(x=.xmin, prop, ymin=prop-se, ymax=prop+se),
                    position=position_dodge(width=.3), 
                    data = ~select(., data) %>% unnest(data) %>% filter(glu==0)) +
    scale_x_continuous(trans='log2') +
    coord_cartesian(xlim=c(.xmin*.18, NA)) +
    # theme_cowplot() +
    NULL
  
  plot_grid(
    .prop_pl +
      coord_cartesian(xlim=c(1.35*.xmin, 1.5e3)) +
      scale_x_continuous(trans="log2", breaks=.xbreaks) +
      scale_y_continuous(breaks=c(0, 1)) +
      labs(y="prop.\ninduced") +
      theme(axis.title.x = element_blank(), axis.text.x = element_blank(),
            axis.ticks.x = element_blank(), axis.line.x = element_blank()) +
      # theme(...) +
      NULL, 
    .gr_pl +
      coord_cartesian(xlim=c(1.35*.xmin, 1.5e3), ylim=c(NA, 1.25)) +
      scale_x_continuous(trans="log2", breaks=.xbreaks) +
      # theme_cowplot_legend_inset() +
      theme(legend.box.spacing = unit(-12, "pt")) +
      # theme(...) +
      NULL,
    ncol = 1, rel_heights = c(0.6, 2), align = "v"
  )
})()


```


```{r more plots for Erik, eval=FALSE}
   myrates_mix %>% 
    group_by(date, condition) %>% 
    filter(between(grate, quantile(grate, .01, na.rm=TRUE), quantile(grate, .99, na.rm=TRUE) )) %>%
    group_by(condition, induced) %>% 
    summarise(
      gr_mean=mean(grate*3600),
      gr_sd  =sd(grate*3600),
      n=n(),
    ) %>% 
    group_by(condition) %>% 
    mutate(prop=n/sum(n)) %>% 
    parse_sugarmix_conds(condition) %>% 
    filter(treatment != "iptg") %>%
  write_delim("fig3B.txt", delim = ="\t")

myrates_mix %>% 
  group_by(date, condition) %>% 
  filter(between(grate, quantile(grate, .01, na.rm=TRUE), quantile(grate, .99, na.rm=TRUE) )) %>%
  parse_sugarmix_conds(condition) %>% 
  filter(treatment != "iptg") %>%
  mutate(induced=fct_recode(factor(induced), "uninduced"="0", "induced"="1")) %>% 
  mutate(treatment=fct_recode(factor(treatment), "glucose only"="none", "glucose + lactose"="lac002")) %>% 
  ggplot(aes(grate*3600, col=factor(glu))) +
  facet_grid(treatment ~ induced) +
  stat_ecdf() +
  scale_colour_hue() +
  labs(x="growth rate (/h)", y="cumul. proba.", col="[glucose]") +
  NULL
   
myrates_mix %>% 
    group_by(date, condition) %>% 
    filter(between(grate, quantile(grate, .01, na.rm=TRUE), quantile(grate, .99, na.rm=TRUE) )) %>%
    group_by(condition, induced) %>% 
    summarise(
      gr_mean=mean(grate*3600),
      gr_sd  =sd(grate*3600),
      n=n(),
    ) %>% 
    group_by(condition) %>% 
    mutate(prop=n/sum(n)) %>% 
    parse_sugarmix_conds(condition) %>% 
    filter(treatment != "iptg") %>%
    ggplot(aes(glu, gr_mean)) +
    # geom_rect(aes(xmin=0, xmax=Inf, ymin=gr_mean-gr_sd, ymax=gr_mean+gr_sd, col=NULL), 
    #           fill="gray92", show.legend=FALSE,
    #           data=~filter(., treatment=='lac002', glu==0, induced==1) ) +
    geom_hline(aes(yintercept = gr_mean, col=NULL), col="gray40", lty="dotted",
               data=~filter(., treatment=='lac002', glu==0, induced==1) ) +
    # must filter glu>0 with position=dodge
    geom_point(aes(col=interaction(treatment, induced)), size=2.5, 
                    position=position_dodge(width=.5, preserve="total"), 
                    data=~filter(., glu > 0) ) +
    geom_linerange(aes(ymin=gr_mean-gr_sd, ymax=gr_mean+gr_sd, col=interaction(treatment, induced)),
                    position=position_dodge(width=.5, preserve="total"), show.legend=FALSE,
                    data=~filter(., glu > 0) ) +
    # display legend without errorbars
    # geom_point(aes(fill=factor(induced)), show.legend=TRUE, size=5) +
    coord_cartesian(ylim=c(NA, 1.)) +
    scale_x_continuous(trans='log2', breaks=2^c(0, 4, 8)) +
    scale_y_continuous(breaks=c(0, 0.5, 1)) +
    labs(x='[glucose] (µM)', y='growth rate (/h)') +
    scale_colour_manual(name='', 
      limits = c('none.0', 'lac002.0', 'lac002.1', 'lac'), # set colours order
      values=c('none.0' = colorspace::desaturate( colorspace::lighten(qual_cols[1], .5), .2), 
               # 'none.0' = qual_cols[1], 
               'lac002.0' = qual_cols[1], 
               'lac002.1' = qual_cols[2],
               'lac' = colorspace::desaturate( colorspace::lighten(qual_cols[2], .5), .2) ),
      labels=c('none.0'='glucose only', 
               'lac002.0'='glucose + lactose 0.55 mM (uninduced)', 
               'lac002.1'='glucose + lactose 0.55 mM (induced)',
               'lac'='lactose 0.55 mM only') 
      ) +
    # theme_cowplot() +
    theme(legend.position = "bottom", legend.direction = "vertical", 
          legend.box.just="left") +
    NULL

```

```{r eval=FALSE}
# number of cells
myrates_mix %>% 
  group_by(date, condition) %>% 
  filter(between(grate, quantile(grate, .01, na.rm=TRUE), quantile(grate, .99, na.rm=TRUE) )) %>%
  parse_sugarmix_conds(condition) %>% 
  filter(treatment != "iptg") %>% 
  group_by(treatment, glu, date) %>% 
  summarise(n=n()) %>% 
  # arrange(n) %>% 
  print(n=Inf)

# growth rate range
myrates_mix %>% 
  group_by(date, condition) %>% 
  filter(between(grate, quantile(grate, .01, na.rm=TRUE), quantile(grate, .99, na.rm=TRUE) )) %>%
  group_by(condition, induced) %>% 
  summarise(
    gr_mean=mean(grate/log(2)*3600),
    gr_sd  =sd(grate/log(2)*3600),
    n=n(),
  ) %>% 
  parse_sugarmix_conds(condition) %>% 
  filter(treatment=='none') %>% 
  arrange(glu) %>% 
  print(n=Inf)

```


```{r eval=FALSE}
# looking at the relative error of growth rate estimation
myrates_mix %>% 
  mutate(grate_se = map_dbl(mod, ~.$stderr[2] ),
         se=grate_se/grate) %>% 
  group_by(date, condition) %>% 
  filter(between(grate_se/grate, -1, 2)) %>% 
  # filter(between(grate_se/grate, quantile(grate_se/grate, .01, na.rm=TRUE), quantile(grate_se/grate, .99, na.rm=TRUE) )) %>% with(range(se))
  parse_sugarmix_conds(condition) %>% 
  filter(glu<1000) %>% 
  ggplot(aes(grate_se/grate, col=factor(induced), group=interaction(date, condition, induced))) +
  facet_grid(glu~treatment) +
  geom_freqpoly(aes(y=after_stat(density)), binwidth=.02) +
  coord_cartesian(xlim=c(-.1, .2)) +
  scale_colour_discrete(name='', labels=c('0'='uninduced', '1'='induced')) +
  theme_cowplot() +
  NULL

```

```{r glucose monod curve}
(myplots[['sugarmix_monod']] <- 
   myrates_mix %>% 
   group_by(date, condition) %>% 
   filter(between(grate, quantile(grate, .01, na.rm=TRUE), quantile(grate, .99, na.rm=TRUE) )) %>%
   group_by(date, condition, induced) %>% 
   summarise(
     gr_mean=mean(grate/log(2)*3600),
     gr_sd  =sd(grate/log(2)*3600),
     n=n(),
   ) %>% 
   parse_sugarmix_conds(condition) %>% 
   filter(treatment=='none', induced==0) %>% 
   filter(glu<2000) %>% 
   (function(.df) {
     # browser()
     # .mod <- lm(gr_inv~glu_inv, weights=1/.df$gr_sd^2, data=mutate(.df, gr_inv=1/gr_mean, glu_inv=1/glu))
     # .gr_max <- 1/coef(.mod)[1]
     # .glu_half <- coef(.mod)[2]*.gr_max
     
     .monod <- function(.glu, .gr_max, .glu_half) .gr_max * .glu / (.glu + .glu_half)
     .mod <- nls(gr_mean~.monod(glu, gr_max, glu_half), data=.df, start=list(gr_max=1, glu_half=5))
     .gr_max <- coef(.mod)[1]
     .glu_half <- coef(.mod)[2]
     print(summary(.mod))

     ggplot(.df, aes(glu, gr_mean)) +
       # geom_rect(aes(xmin=-Inf, xmax=Inf, ymin=gr_mean-gr_sd, ymax=gr_mean+gr_sd, group=date), fill="grey10", alpha=.2,
       #           data=~filter(., treatment=='lac002', glu==0, induced==1)) +
       # geom_hline(aes(yintercept=gr_mean, group=date), lty="dashed",
       #            data=~filter(., treatment=='lac002', glu==0, induced==1)) +
       geom_vline(xintercept=0, lty="dotted") +
       stat_function(fun = function(.glu) .gr_max * .glu / (.glu + .glu_half), col='gray') +
       geom_pointrange(aes(ymin=gr_mean-gr_sd, ymax=gr_mean+gr_sd),
                       # position=position_dodge2(width=.4), 
       ) +
       coord_cartesian(ylim=c(NA, 1.4)) +
       scale_size_continuous(range=c(0.05, .5), guide='none') +
       # scale_x_continuous(trans='log2') +
       labs(x="[glucose] (µM)", y="growth rate (dbl/h)") +
       theme_cowplot() +
       NULL
   }))

# Formula: gr_mean ~ .monod(glu, gr_max, glu_half)
# 
# Parameters:
#          Estimate Std. Error t value Pr(>|t|)    
# gr_max    0.98120    0.05581  17.581 7.54e-09 ***
# glu_half 10.15356    2.07457   4.894 0.000628 ***
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
# 
# Residual standard error: 0.08315 on 10 degrees of freedom
# 
# Number of iterations to convergence: 8 
# Achieved convergence tolerance: 1.957e-06


# plotting 1/gr vs 1/glu
(myplots[['sugarmix_monod_inv']] <- 
    myrates_mix %>%
    group_by(date, condition) %>%
    filter(between(grate, quantile(grate, .01, na.rm=TRUE), quantile(grate, .99, na.rm=TRUE) )) %>%
    group_by(date, condition, induced) %>%
    summarise(
      gr_mean=mean(grate/log(2)*3600),
      gr_sd  =sd(grate/log(2)*3600),
      gr_re=gr_sd/sqrt(n()) / gr_mean,
      dt_mean=mean(log(2)/grate/3600),
      dt_sd  =sd(log(2)/grate/3600),
      ymin=1/pmax(0, gr_mean+gr_sd), ymax=1/pmax(0, gr_mean-gr_sd),
      n=n(),
    ) %>%
    parse_sugarmix_conds(condition) %>%
    filter(treatment=='none', induced==0) %>% 
    # filter(glu<1000) %>%
    # transmute(glu=glu, glu2=1/glu, glu3=1/glu/.18) %>% arrange(glu)
    ggplot(aes(1/glu, 1/gr_mean)) +
    stat_smooth(method='lm', se=FALSE, col="gray", fullrange=TRUE) +
    # geom_point() +
    geom_vline(xintercept=0, lty="dotted") +
    geom_pointrange(aes(ymin=1/pmax(0, gr_mean+gr_sd), ymax=1/pmax(0, gr_mean-gr_sd)),
                    data=~filter(., 1/glu < 0.1)) +
    geom_pointrange(aes(ymin=1/pmax(0, gr_mean+gr_sd), ymax=1/pmax(0, gr_mean-gr_sd)),
                    data=~filter(., 1/glu > 0.1), position=position_dodge2(width=1.5e-2, preserve="single")) +
    expand_limits(x=0, y=0) +
    coord_cartesian(ylim=c(0, 7)) +
    scale_size_continuous(range=c(0.05, .5), guide='none') +
    labs(x = "1 / [glucose] (µM⁻¹)", y = "1 / growth rate (h)") +
    theme_cowplot() +
    NULL)

```


## Analysing LacZ-GFP expression

```{r}
myrates_mix %>% ungroup() %>% 
  mutate(gfp_m = map_dbl(data, ~mean(.$gfp_nb/.$length_um)),
         gfp_sd = map_dbl(data, ~sd(.$gfp_nb/.$length_um)),
         time_birth=map_dbl(data, ~min(.$time_sec)),
  ) %>% 
  group_by(date, condition) %>% 
  mutate(prop = ifelse(induced, sum(induced)/n(), 1-sum(induced)/n())) %>% 
  parse_sugarmix_conds(condition) %>% 
  filter(induced==1) %>% 
  filter(n()>10) %>% 
  ggplot(aes(gfp_m, col=factor(date))) +
  facet_grid(glu~treatment) +
  geom_freqpoly(aes(y=after_stat(density), size=prop, group=interaction(date, condition, prop)), bins=100) +
  # scale_y_continuous(breaks = scales::breaks_extended(n=3)) +
  # coord_cartesian(xlim=c(-.25, 1.5), ylim=c(0, 3.5)) +
  scale_size_continuous(range=c(0.05, 1), guide='none') +
  scale_x_log10() +
  labs(x='[LacZ-GFP] (AU)') +
  theme(legend.position = 'top') +
  NULL
```

```{r eval=FALSE}
myrates_mix %>% ungroup() %>% 
  mutate(gfp_m = map_dbl(data, ~mean(.$gfp_nb/.$length_um)),
         gfp_sd = map_dbl(data, ~sd(.$gfp_nb/.$length_um)),
         time_birth=map_dbl(data, ~min(.$time_sec)),
  ) %>% 
  group_by(date, condition) %>% 
  # filter(between(grate, quantile(grate, .01, na.rm=TRUE), quantile(grate, .99, na.rm=TRUE) )) %>%
  mutate(prop = ifelse(induced, sum(induced)/n(), 1-sum(induced)/n())) %>% 
  parse_sugarmix_conds(condition) %>% 
  filter(treatment=="iptg") %>% 
  filter(induced==1) %>% 
  filter(n()>10) %>% 
  ggplot(aes(gfp_m, col=factor(date))) +
  facet_grid(glu~treatment) +
  geom_freqpoly(aes(y=after_stat(density), size=prop, group=interaction(date, condition, prop)), bins=100) +
  # scale_y_continuous(breaks = scales::breaks_extended(n=3)) +
  # coord_cartesian(xlim=c(-.25, 1.5), ylim=c(0, 3.5)) +
  scale_size_continuous(range=c(0.05, 1), guide='none') +
  scale_x_log10() +
  labs(x='[LacZ-GFP] (AU)') +
  theme(legend.position = 'top') +
  NULL
```


### CRP regulation as a function of growth rate

First without correcting for bleaching.

```{r}
# gfp_c vs grate
myrates_mix %>% ungroup() %>% 
  mutate(gfp_m  = map_dbl(data, ~mean(.$gfp_nb/.$length_um)),
         gfp_sd = map_dbl(data, ~sd(.$gfp_nb/.$length_um)),
         time_birth=map_dbl(data, ~min(.$time_sec)),
         grate2=grate/log(2)*3600,
  ) %>% 
  group_by(date, condition) %>% 
  filter(between(grate, quantile(grate, .01, na.rm=TRUE), quantile(grate, .99, na.rm=TRUE) )) %>%
  mutate(prop = ifelse(induced, sum(induced)/n(), 1-sum(induced)/n())) %>% 
  filter(induced==1) %>% 
  filter(n()>10) %>% 
  parse_sugarmix_conds(condition) %>% 
  filter(treatment=="iptg") %>% 
  # lm(gfp_m~grate2, data=.)
  ggplot(aes(grate2, gfp_m, col=factor(glu))) +
  geom_point(alpha=.5, stroke=0, size=1) +
  geom_abline(intercept = 8500, slope=-7000, col='grey20', lty='dotted') +
  geom_point(stroke=2, size=4,
             data=~group_by(., date, glu) %>% 
               summarise(grate2=mean(grate2), gfp_m=exp(mean(log(gfp_m))) ) ) +
  # coord_cartesian(xlim=c(0, 1.5)) +
  labs(x="growth rate (dbl/h)", y="[LacZ-GFP] (unbleached)", col="[glu] (µM)") +
  theme(legend.position=c(.95, .95), legend.justification=c(1, 1)) +
  NULL

```


Then correcting for bleaching assuming constant $q$ and $\lambda$.
Hence $c = c_u (1+\beta/\lambda)$.
Note that we recover the so-called C-line of Hwa, *et al.*

```{r}
# gfp_c vs grate (bleaching corrected per cell)
myrates_mix %>% ungroup() %>% 
  mutate(gfp_m  = map_dbl(data, ~mean(.$gfp_nb/.$length_um)),
         gfp_sd = map_dbl(data, ~sd(.$gfp_nb/.$length_um)),
         gfp_b = 5.3e-5/5, # Kaiser2018 reports 5.3e-5 / sec for 5x stronger illumination
         gfp_tot = gfp_m * (1 + gfp_b/grate),
         time_birth=map_dbl(data, ~min(.$time_sec)),
         grate2=grate/log(2)*3600,
  ) %>% 
  group_by(date, condition) %>% 
  filter(between(grate, quantile(grate, .01, na.rm=TRUE), quantile(grate, .99, na.rm=TRUE) )) %>%
  mutate(prop = ifelse(induced, sum(induced)/n(), 1-sum(induced)/n())) %>% 
  filter(induced==1) %>% 
  filter(n()>10) %>% 
  filter(grate2 > .01) %>% 
  parse_sugarmix_conds(condition) %>% 
  filter(treatment=="iptg") %>% 
  # lm(gfp_m~grate2, data=.)
  ggplot(aes(grate2, gfp_tot, col=factor(glu))) +
  geom_point(alpha=.5, stroke=0, size=1) +
  # stat_smooth(method="lm", aes(col=factor(glu), group=interaction(date, glu))) +
  stat_smooth(method="lm", fullrange=TRUE, col="black", #aes(col="mean"),
             data=~group_by(., date, glu) %>% 
               summarise(grate2=mean(grate2), gfp_tot=mean(gfp_tot, na.rm=T)) ) +
  geom_point(stroke=2, size=4,
             data=~group_by(., date, glu) %>% 
               summarise(grate2=mean(grate2), gfp_tot=mean(gfp_tot, na.rm=T)) ) +
  coord_cartesian(ylim=c(0, 1.3e4)) +
  labs(title="bleaching corrected per cell", 
       x="growth rate (dbl/h)", y="[LacZ-GFP]", col="[glu] (µM)") +
  theme(legend.position=c(.95, .95), legend.justification=c(1, 1)) +
  NULL

# gfp_c vs grate (bleaching corrected per experiment)
myrates_mix %>% ungroup() %>% 
  parse_sugarmix_conds(condition) %>% 
  filter(treatment=="iptg") %>% 
  mutate(gfp_m  = map_dbl(data, ~mean(.$gfp_nb/.$length_um)),
         gfp_sd = map_dbl(data, ~sd(.$gfp_nb/.$length_um)),
         time_birth=map_dbl(data, ~min(.$time_sec)),
         grate2=grate/log(2)*3600,
  ) %>% 
  group_by(date, condition) %>% 
  filter(between(grate, quantile(grate, .01, na.rm=TRUE), quantile(grate, .99, na.rm=TRUE) )) %>%
  # filter(grate2 > .01) %>% 
  mutate(grate_mean = mean(grate),
         gfp_b = 5.3e-5/5, # Kaiser2018 reports 5.3e-5 / sec for 5x stronger illumination
         gfp_tot = gfp_m * (1 + gfp_b/grate_mean),
) %>% 
  mutate(prop = ifelse(induced, sum(induced)/n(), 1-sum(induced)/n())) %>% 
  filter(induced==1) %>% 
  filter(n()>10) %>% 
  # lm(gfp_m~grate2, data=.)
  ggplot(aes(grate2, gfp_tot, col=factor(glu))) +
  geom_point(alpha=.5, stroke=0, size=1) +
  # stat_smooth(method="lm", aes(col=factor(glu), group=interaction(date, glu))) +
  stat_smooth(method="lm", fullrange=TRUE, col="black", #aes(col="mean"),
             data=~group_by(., date, glu) %>% 
               summarise(grate2=mean(grate2), gfp_tot=mean(gfp_tot, na.rm=T)) ) +
  geom_point(stroke=2, size=4,
             data=~group_by(., date, glu) %>% 
               summarise(grate2=mean(grate2), gfp_tot=mean(gfp_tot, na.rm=T)) ) +
  coord_cartesian(ylim=c(0, 1.3e4)) +
  labs(title="bleaching corrected per experiment",
       x="growth rate (dbl/h)", y="[LacZ-GFP]", col="[glu] (µM)") +
  theme(legend.position=c(.95, .95), legend.justification=c(1, 1)) +
  NULL

```

In order to plot this in concentration units, one measures width at the different glucose concentrations.

```{r}
 fs::dir_ls(here("data", "sugarMix-cellWidth"), recurse=TRUE, glob = "*.txt") %>% 
   tibble(path=.) %>% 
   extract("path", c("glu", "date"), "glu(\\d+)uM/(20\\d{6})", remove = FALSE) %>% 
   mutate(
     glu=as.numeric(glu), 
     data=map(path, ~readr::read_delim(., delim="\t", col_types=cols_only( MinFeret=col_number() )) )
   ) %>%
   unnest(data) %>% 
   identity() %>% 
   ggplot(aes(glu, MinFeret)) +
   geom_jitter(aes(col=date), width=.25) +
   # geom_jitter(width=.2, col="gray50") +
   geom_boxplot(aes(group=glu), fill="transparent", outlier.shape = NA) +
   scale_x_continuous(trans="log2", breaks=2^c(1, 3, 5, 7)) +
   labs(x="[glucose] (µM)", y="cell width (µm)") +
   expand_limits(y=0) +
   NULL
             
```



```{r}
rod_volume <- function(.l, .w) pi * ( (.w/2)^2 * (.l - .w) + 4/3 * (.w/2)^3 )

(myplots[['sugarmix_crp']] <- 
    myrates_mix %>% ungroup() %>% 
    parse_sugarmix_conds(condition) %>% 
    filter(treatment=="iptg") %>% 
   # append average width per condition
   left_join(
     fs::dir_ls(here("data", "sugarMix-cellWidth"), recurse=TRUE, glob = "*.txt") %>% 
       tibble(path=.) %>% 
       extract("path", c("glu", "date"), "glu(\\d+)uM/(20\\d{6})", remove = FALSE) %>% 
       mutate(
         glu=as.numeric(glu), 
         data=map(path, ~readr::read_delim(., delim="\t", col_types=cols_only( MinFeret=col_number() )) )
       ) %>%
       unnest(data) %>% 
       group_by(glu) %>% 
       summarise(width_um = median(MinFeret))
   ) %>% 
   mutate(gfp_m  = map2_dbl(data, width_um,
                            ~mean(..1$gfp_nb/rod_volume(..1$length_um, ..2))),
           gfp_sd = map2_dbl(data, width_um, 
                            ~sd(..1$gfp_nb/rod_volume(..1$length_um, ..2))),
           gfp_b = 5.3e-5/5, # Kaiser2018 reports 5.3e-5 / sec for 5x stronger illumination
           gfp_tot = gfp_m * (1 + gfp_b/grate),
           time_birth=map_dbl(data, ~min(.$time_sec)),
           grate2=grate/log(2)*3600,
    ) %>% 
    group_by(date, condition) %>%
    filter(between(grate, quantile(grate, .01, na.rm=TRUE), quantile(grate, .99, na.rm=TRUE) )) %>%
    mutate(prop = ifelse(induced, sum(induced)/n(), 1-sum(induced)/n())) %>% 
    filter(induced==1) %>%
    filter(grate2 > .01) %>%
   
   # compute summary stats
    group_by(date, glu) %>% 
    summarise(grate2_m=mean(grate2), grate2_sd=sd(grate2), 
              # gfp2_m=exp(mean(log(gfp_tot), na.rm=T)),
              # gfp2_min=exp(mean(log(gfp_tot), na.rm=T) - sd(log(gfp_tot), na.rm=T)),
              # gfp2_max=exp(mean(log(gfp_tot), na.rm=T) + sd(log(gfp_tot), na.rm=T)),
              # conversion to molar units from FP/volume:  * 1e21 / 6.02e23
              gfp2_m=mean(gfp_tot, na.rm=T) * 1e21 / 6.02e23,
              gfp2_sd=sd(gfp_tot, na.rm=T)  * 1e21 / 6.02e23,
              n=n()
              ) %>%  
    ungroup() %>% 
    mutate(glucose=glu, glu=factor(glu),
           glu=factor(glu, labels=paste(levels(glu), 'µM')) ) %>% 
    # (function(.df) {
    #   cat( format_csv(select(.df, -glu)) )
    #   print( summary(lm(gfp2_m~grate2_m, data=.df)) )
    #   return(.df)
    # }) %>% 
    ggplot(aes(grate2_m, gfp2_m, col=glu)) +
    stat_smooth(method="lm", fullrange=TRUE, se=FALSE, col="black") +
    geom_errorbarh(aes(xmin=grate2_m-grate2_sd, xmax=grate2_m+grate2_sd), height=0, show.legend=F) + 
    geom_errorbar(aes(ymin=gfp2_m-gfp2_sd, ymax=gfp2_m+gfp2_sd), width=0, show.legend=F) + 
    geom_point(size=4) +
    scale_color_viridis_d(end = .85) +
    scale_x_continuous(breaks=c(0, 0.5, 1)) +
    expand_limits(y=0) +
    # expand_limits(x=c(0, 1.1), y=0) +
    labs(x="growth rate (dbl/h)", y="[LacZ-GFP] (µM)", col="[glucose]") +
    theme_cowplot_legend_inset() +
    theme(legend.position=c(.95, .95), legend.justification=c(1, 1)) +
    NULL)

```

```{r even more plots for erik, eval=FALSE}
local({
  myrates_mix %>% ungroup() %>% 
    parse_sugarmix_conds(condition) %>% 
    filter(treatment=="iptg") %>% 
    # append average width per condition
    left_join(
      fs::dir_ls(here("data", "sugarMix-cellWidth"), recurse=TRUE, glob = "*.txt") %>% 
        tibble(path=.) %>% 
        extract("path", c("glu", "date"), "glu(\\d+)uM/(20\\d{6})", remove = FALSE) %>% 
        mutate(
          glu=as.numeric(glu), 
          data=map(path, ~readr::read_delim(., delim="\t", col_types=cols_only( MinFeret=col_number() )) )
        ) %>%
        unnest(data) %>% 
        group_by(glu) %>% 
        summarise(width_um = median(MinFeret))
    ) %>% 
    mutate(gfp_m  = map2_dbl(data, width_um,
                             ~mean(..1$gfp_nb/rod_volume(..1$length_um, ..2))),
           gfp_sd = map2_dbl(data, width_um, 
                             ~sd(..1$gfp_nb/rod_volume(..1$length_um, ..2))),
           gfp_b = 5.3e-5/5, # Kaiser2018 reports 5.3e-5 / sec for 5x stronger illumination
           gfp_tot = gfp_m * (1 + gfp_b/grate),
           time_birth=map_dbl(data, ~min(.$time_sec)),
    ) %>% 
    filter(induced==1) %>%
    ggplot(aes(grate*3600, col=factor(glu), group=interaction(glu, date))) +
    # facet_grid(induced~.) +
    stat_ecdf() +
    coord_cartesian(xlim = c(-.1, 1)) +
    labs(x="growth rate (/h)", y="cumul. proba.", col="[glucose]") +
    NULL
  
.myrates_crp <- myrates_mix %>% ungroup() %>% 
    parse_sugarmix_conds(condition) %>% 
    filter(treatment=="iptg") %>% 
   # append average width per condition
   left_join(
     fs::dir_ls(here("data", "sugarMix-cellWidth"), recurse=TRUE, glob = "*.txt") %>% 
       tibble(path=.) %>% 
       extract("path", c("glu", "date"), "glu(\\d+)uM/(20\\d{6})", remove = FALSE) %>% 
       mutate(
         glu=as.numeric(glu), 
         data=map(path, ~readr::read_delim(., delim="\t", col_types=cols_only( MinFeret=col_number() )) )
       ) %>%
       unnest(data) %>% 
       group_by(glu) %>% 
       summarise(width_um = median(MinFeret))
   ) %>% 
   mutate(gfp_m  = map2_dbl(data, width_um,
                            ~mean(..1$gfp_nb/rod_volume(..1$length_um, ..2))),
           gfp_sd = map2_dbl(data, width_um, 
                            ~sd(..1$gfp_nb/rod_volume(..1$length_um, ..2))),
           gfp_b = 5.3e-5/5, # Kaiser2018 reports 5.3e-5 / sec for 5x stronger illumination
           gfp_tot = gfp_m * (1 + gfp_b/grate),
           time_birth=map_dbl(data, ~min(.$time_sec)),
           grate2=grate/log(2)*3600,
    ) %>% 
    group_by(date, condition) %>%
    filter(between(grate, quantile(grate, .01, na.rm=TRUE), quantile(grate, .99, na.rm=TRUE) )) %>%
    mutate(prop = ifelse(induced, sum(induced)/n(), 1-sum(induced)/n())) %>% 
    filter(induced==1) %>%
    filter(grate2 > .01) %>%
   
   # compute summary stats
    group_by(date, glu) %>% 
    summarise(grate_m=mean(grate*3600), grate_sd=sd(grate*3600), 
              # gfp2_m=exp(mean(log(gfp_tot), na.rm=T)),
              # gfp2_min=exp(mean(log(gfp_tot), na.rm=T) - sd(log(gfp_tot), na.rm=T)),
              # gfp2_max=exp(mean(log(gfp_tot), na.rm=T) + sd(log(gfp_tot), na.rm=T)),
              # conversion to molar units from FP/volume:  * 1e21 / 6.02e23
              gfp_tot_m=mean(gfp_tot, na.rm=T) * 1e21 / 6.02e23,
              gfp_tot_sd=sd(gfp_tot, na.rm=T)  * 1e21 / 6.02e23,
              n=n()
              )

.myrates_crp %>%
  write_delim("fig3B.txt", delim="\t")
  
.myrates_crp %>%
  lm(gfp_tot_m ~ grate_m, data=.) %>% 
  summary() %>% print()
  
.myrates_crp %>%  
    ungroup() %>% 
    mutate(glucose=glu, glu=factor(glu),
           glu=factor(glu, labels=paste(levels(glu), 'µM')) ) %>% 
    # (function(.df) {
    #   cat( format_csv(select(.df, -glu)) )
    #   print( summary(lm(gfp2_m~grate2_m, data=.df)) )
    #   return(.df)
    # }) %>% 
    ggplot(aes(grate_m, gfp_tot_m, col=glu)) +
    stat_smooth(method="lm", fullrange=TRUE, se=FALSE, col="black") +
    geom_errorbarh(aes(xmin=grate_m-grate_sd, xmax=grate_m+grate_sd), height=0, show.legend=F) + 
    geom_errorbar(aes(ymin=gfp_tot_m-gfp_tot_sd, ymax=gfp_tot_m+gfp_tot_sd), width=0, show.legend=F) + 
    geom_point(size=4) +
    scale_color_viridis_d(end = .85) +
    scale_x_continuous(breaks=c(0, 0.5, 1)) +
    expand_limits(y=0) +
    # expand_limits(x=c(0, 1.1), y=0) +
    labs(x="growth rate (/h)", y="[LacZ-GFP] (µM)", col="[glucose]") +
    theme_cowplot_legend_inset() +
    theme(legend.position=c(.95, .95), legend.justification=c(1, 1)) +
    NULL
})

```


```{r eval=FALSE}
# number of cells
myrates_mix %>% ungroup() %>% 
  group_by(date, condition) %>% 
  filter(between(grate, quantile(grate, .01, na.rm=TRUE), quantile(grate, .99, na.rm=TRUE) )) %>%
  filter(induced==1) %>% 
  filter(n()>10) %>% 
  ungroup() %>% 
  parse_sugarmix_conds(condition) %>% 
  filter(treatment=="iptg") %>% 
  group_by(treatment, glu, date) %>% 
  summarise(n=n()) %>% 
  # arrange(n) %>% 
  print(n=Inf)

```

```{r eval=FALSE}
# looking at activity instead.
myrates_mix %>% ungroup() %>% 
  mutate(gfp_m  = map_dbl(data, ~mean(.$gfp_nb)),
         gfp_b = 5.3e-5/5, # Kaiser2018 reports 5.3e-5 / sec for 5x stronger illumination
         dg_dt  = map_dbl(data, ~(last(.$gfp_nb)-first(.$gfp_nb))/diff(range(.$time_sec))),
         lac_act = dg_dt + gfp_b*gfp_m,
         time_birth=map_dbl(data, ~min(.$time_sec)),
         grate2=grate/log(2)*3600,
  ) %>% 
  group_by(date, condition) %>% 
  filter(between(grate, quantile(grate, .01, na.rm=TRUE), quantile(grate, .99, na.rm=TRUE) )) %>%
  filter(induced==1) %>% 
  filter(n()>10) %>% 
  parse_sugarmix_conds(condition) %>% 
  filter(treatment=="iptg") %>% 
  # lm(dg_dt~I(gfp_b*gfp_m), data=.)
  # lm(dg_dt~poly(lac_act, 2), data=.)
  ggplot(aes(grate2, lac_act, col=factor(glu))) +
  geom_point(alpha=.5, stroke=0, size=1) +
  geom_point(stroke=2, size=4,
             data=~group_by(., date, glu) %>%
               summarise(grate2=mean(grate2), lac_act=mean(lac_act)) ) +
  stat_smooth(aes(group=1), col='black', method='lm', formula='y ~ poly(x, 2)',
             data=~group_by(., date, glu) %>%
               summarise(grate2=mean(grate2), lac_act=mean(lac_act)) ) +
  coord_cartesian(ylim=c(-.1, 1.)) +
  NULL
  
```

### CRP regulation in single cells

```{r}
myrates_mix %>% ungroup() %>%
  mutate(gfp_m  = map_dbl(data, ~mean(.$gfp_nb/.$length_um)),
         gfp_sd = map_dbl(data, ~sd(.$gfp_nb/.$length_um)),
         gfp_b = 5.3e-5/5, # Kaiser2018 reports 5.3e-5 / sec for 5x stronger illumination
         gfp_tot = gfp_m * (1 + gfp_b/grate),
         time_birth=map_dbl(data, ~min(.$time_sec)),
         grate2=grate/log(2)*3600,
  ) %>%
  group_by(date, condition) %>%
  filter(between(grate, quantile(grate, .01, na.rm=TRUE), quantile(grate, .99, na.rm=TRUE) )) %>%
  mutate(prop = ifelse(induced, sum(induced)/n(), 1-sum(induced)/n())) %>%
  filter(induced==1) %>%
  filter(n()>10) %>%
  parse_sugarmix_conds(condition) %>%
  filter(treatment=="iptg") %>%
  # lm(gfp_m~grate2, data=.)
  ggplot(aes(grate2, gfp_tot, col=factor(glu))) +
  geom_point(alpha=.5, stroke=0, size=1) +
  stat_smooth(method=MASS::rlm, fullrange=TRUE, se=FALSE, aes(group=interaction(date, glu)) ) +
  stat_smooth(method="lm", fullrange=TRUE, col="black", #aes(col="mean"),
              data=~group_by(., date, glu) %>%
                summarise(grate2=mean(grate2), gfp_tot=exp(mean(log(gfp_tot), na.rm=T)) ) ) +
  geom_point(stroke=2, size=4,
             data=~group_by(., date, glu) %>%
               summarise(grate2=mean(grate2), gfp_tot=exp(mean(log(gfp_tot), na.rm=T)) ) ) +
  coord_cartesian(ylim=c(0, 1.3e4)) +
  labs(x="growth rate (dbl/h)", y="[LacZ-GFP]", col="[glu] (µM)") +
  theme(legend.position=c(.95, .95), legend.justification=c(1, 1)) +
  NULL


myrates_mix %>% ungroup() %>%
  mutate(gfp_m  = map_dbl(data, ~mean(.$gfp_nb/.$length_um)),
         gfp_sd = map_dbl(data, ~sd(.$gfp_nb/.$length_um)),
         gfp_b = 5.3e-5/5, # Kaiser2018 reports 5.3e-5 / sec for 5x stronger illumination
         gfp_tot = gfp_m * (1 + gfp_b/grate),
         time_birth=map_dbl(data, ~min(.$time_sec)),
         grate2=grate/log(2)*3600,
  ) %>%
  group_by(date, condition) %>%
  filter(between(grate, quantile(grate, .01, na.rm=TRUE), quantile(grate, .99, na.rm=TRUE) )) %>%
  mutate(prop = ifelse(induced, sum(induced)/n(), 1-sum(induced)/n())) %>%
  filter(induced==1) %>%
  filter(n()>10) %>%
  parse_sugarmix_conds(condition) %>%
  filter(treatment=="iptg") %>%
  group_by(date, glu) %>% 
  nest() %>% 
  mutate(mod=map(data, ~MASS::rlm(gfp_tot~grate2, data=.)),
         slope=map_dbl(mod, ~coef(.)[2]),
         slope_se=map_dbl(mod, ~summary(.)$coef[2,2]),
  ) %>% 
  ggplot(aes(glu, slope, group=date)) +
  geom_hline(yintercept=0, lty='dotted') +
  geom_hline(aes(yintercept=slope), lty='dashed',
             data=~mutate(.,
                          grate2=map_dbl(data, ~mean(.$grate2)), 
                          gfp_tot=map_dbl(data, ~exp(mean(log(.$gfp_tot), na.rm=T))),
             ) %>%
               ungroup() %>% nest() %>% 
               mutate(slope=map_dbl(data, ~coef(lm(gfp_tot~grate2, data=.))[2])) ) +
  geom_pointrange(aes(ymin=slope-slope_se, ymax=slope+slope_se), position=position_dodge(width=.2)) +
  scale_x_continuous(trans='log2') +
  # scale_y_log10() +
  labs(x="[glucose] (µM)", y="C-line slope") +
  NULL
  
```


## Further analysis

There would be a lot to explore regarding e.g. switching dynamics at the single-cell level...

```{r eval=FALSE}
myframes %>% 
  filter(date==20210504 & str_detect(condition, "lac002") ) %>% 
  # filter(date==20210504 & str_detect(condition, "lac002") | 
  #          date==20210506 & condition=="lac002_glu64uM") %>% 
  filter(str_detect(condition, "(?:glu\\d+)|(?:lac\\d+)")) %>% 
  # remove frames after touching the exit
  group_by(id, .add=TRUE) %>%
  mutate(discard_top=which_touch_exit(vertical_top, 1)) %>%
  mutate(discard_top=ifelse(discard_start, FALSE, discard_top)) %>% # not in the preexpt step (2h)
  mutate(end_type=ifelse(any(discard_top), 'touchtop', end_type)) %>% # update end_type to exit for cells which have touched the vertical cutoff
  # remove daughters of cells that touched the exit
  group_by(path) %>%
  mutate(discard_top=which_to_progeny(discard_top, cid)) %>%
  filter(!discard_top) %>% 
  group_by(date, pos, gl, id) %>%
  filter(n() > 5) %>% 
  # # call cell induced as soon as it crossed the 250 threshold
  # mutate(gfp_nb_raw=gfp_nb) %>% 
  # fill(gfp_nb, .direction="downup") %>% 
  filter(!is.na(gfp_nb)) %>% 
  parse_sugarmix_conds(condition) %>% 
  group_by(date, condition, pos, gl) %>%
  filter(max(time_sec) > 10 * 3600) %>% 
  nest() %>% group_by(date, condition) %>% slice(1:6) %>%
  unnest(data) %>%
  # ggplot(aes(time_sec, gfp_nb/length_um, col=b_rank, group=cell)) +
  ggplot(aes(time_sec, gfp_nb+100, col=b_rank, group=cell)) +
  facet_wrap(~glu+date+pos+gl, ncol=6, dir="h", #scales='free_x',
             labeller = as_labeller(label_value, multi_line=FALSE)) +
  geom_hline(yintercept=350, lty="dashed") +
  # geom_vline(xintercept = 5 * 3600, lty='dotted') +
  geom_line() +
  scale_x_hours() +
  scale_y_log10() +
  coord_cartesian(ylim=c(30, NA)) +
  # coord_cartesian(ylim=c(3, NA)) +
  scale_color_viridis_c(limits=c(0, 3), oob=scales::squish, guide=guide_legend(override.aes=list(size=4))) +
  theme(legend.position = "top") +
  NULL

ggsave("plots/sugarmix_gfpnb_switches.pdf", width=16, height=10)
# ggsave("plots/sugarmix_gfpnb_switches.pdf", width=11, height=7)


myframes %>% 
  filter(date==20210504 & str_detect(condition, "lac002") ) %>% 
  # filter(date==20210504 & str_detect(condition, "lac002") | 
  #          date==20210506 & condition=="lac002_glu64uM") %>% 
  filter(str_detect(condition, "(?:glu\\d+)|(?:lac\\d+)")) %>% 
  # remove frames after touching the exit
  group_by(id, .add=TRUE) %>%
  mutate(discard_top=which_touch_exit(vertical_top, 1)) %>%
  mutate(discard_top=ifelse(discard_start, FALSE, discard_top)) %>% # not in the preexpt step (2h)
  mutate(end_type=ifelse(any(discard_top), 'touchtop', end_type)) %>% # update end_type to exit for cells which have touched the vertical cutoff
  # remove daughters of cells that touched the exit
  group_by(path) %>%
  mutate(discard_top=which_to_progeny(discard_top, cid)) %>%
  filter(!discard_top) %>% 
  group_by(date, pos, gl, id) %>%
  filter(n() > 5) %>% 
  # # call cell induced as soon as it crossed the 250 threshold
  # mutate(gfp_nb_raw=gfp_nb) %>% 
  # fill(gfp_nb, .direction="downup") %>% 
  filter(!is.na(gfp_nb)) %>% 
  parse_sugarmix_conds(condition) %>%
  group_by(date, condition, pos, gl) %>%
  filter(max(time_sec) > 10 * 3600) %>% 
  nest() %>% group_by(date, condition) %>% slice(1:6) %>%
  unnest(data) %>% 
  # ggplot(aes(time_sec, gfp_nb/length_um, col=b_rank, group=cell)) +
  ggplot(aes(time_sec, gfp_nb/length_um + 30, col=b_rank, group=cell)) +
  facet_wrap(~glu+date+pos+gl, ncol=6, dir="h", #scales = 'free_x',
             labeller = as_labeller(label_value, multi_line=FALSE)) +
  # geom_hline(yintercept=350, lty="dashed") +
  # geom_vline(xintercept = 5 * 3600, lty='dotted') +
  geom_line() +
  scale_x_hours() +
  scale_y_log10() +
  coord_cartesian(ylim=c(3, NA)) +
  # coord_cartesian(ylim=c(3, NA)) +
  scale_color_viridis_c(limits=c(0, 3), oob=scales::squish, guide=guide_legend(override.aes=list(size=4))) +
  theme(legend.position = "top") +
  NULL

ggsave("plots/sugarmix_gfpc_switches.pdf", width=16, height=10)

```

