---
title: "Induction lags of the native *lac* operon"
author: Thomas Julou
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

From mother machine experiments, we obtain time series of cell length (used as a proxy for biomass) and LacZ-GFP. Here is an example obtained in a given growth channel.

```{r}
# gfp_nb offset ≈ 100 to avoid stripping values and to keep background fluctuations visually small xxx
(myplots[['raw_traces']] <- function(.ylabel=1.2)
  myframes %>% ungroup() %>% 
  filter(date==20180206) %>% 
  group_by(date, pos, gl) %>% 
  ggCustomTJ::slice_groups(c(1)) %>% 
  filter(!discard_start, !discard_top,
         time_sec>2*3600) %>%
  filter(total_cell_in_lane-cell_num_in_lane<5,
         end_type %in% c('div', 'lost')) %>% 
  # ungroup() %>% mutate(gfp_nb=gfp_nb - 2*min(gfp_nb)) %>% 
  ungroup() %>% mutate(gfp_nb=gfp_nb + 135) %>% 
  gather(variable, value, length_um, gfp_nb) %>% 
  ggplot() +
  # facet_grid(pos+gl~.) +
  facet_wrap(~variable, scales='free_y', ncol=1, strip.position='left',
             labeller=as_labeller(function(.str) {
               .str <- str_replace(.str, "gfp_nb", "LacZ-GFP\nmolecules")
               .str <- str_replace(.str, "length_um", "cell length (µm)")
               return(.str)
             }) ) +
  geom_line(aes(time_sec-2*3600, value, col=ugen)) +
  scale_colour_periodic(.n=3, guide='none') +
  # show medium bar
  geom_vline(aes(xintercept=t_start - 2*3600, group=1), alpha=.2, size=.5, 
            data=filter(condition_ts, condition=='switch_08h', m_cycle<=2)) +
  geom_rect(aes(xmin=t_start - 2*3600, xmax=t_end - 2*3600, ymin=.ylabel, ymax=1.5, fill=medium, group=1), show.legend=FALSE, #size=0.2,
            data=filter(condition_ts, condition=='switch_08h', m_cycle<=2) %>% mutate(variable='length_um')) +
  geom_text(aes(x=t_start+(t_end-t_start)/2 - 2*3600, y=.ylabel, label=medium, group=1), size=4, col='white', hjust=0.5, vjust=-0.5, 
            data=filter(condition_ts, condition=='switch_08h') %>% mutate(variable='length_um', t_start=ifelse(t_start<2*3600, 2*3600, t_start))) +
  # scale_fill_manual(values=c('glucose'=.fill[1], 'lactose'=.fill[2]), guide='none') +
  coord_cartesian(xlim=c(-3600*.4, 20.2*3600), expand=FALSE) +
  scale_x_hours(4) +
  scale_y_continuous(trans='log10', breaks=c(2, 3, 4, 100, 500, 2500)) +
  theme(strip.placement='outside', strip.background.y=element_blank(),
        axis.title.y=element_blank(), strip.text.y=element_text(size=rel(1.2))) +
  NULL
 )(1.3)


```


## *lac* induction lags in naive cells

We are interested in the time taken by cells to induce the *lac* operon when exposed to lactose as the only nutrient. This "*lac* induction lag" is estimated from the LacZ-GFP time series of each cell after the switch. Note that at the first switch, all experiments with variable duration in glucose before the second switch are equivalent and are used together to plot the distribution of lags in "naive" cells (which have not seen lactose before).
  
```{r}
(myplots[['naive_lags_hist']] <- mycells_switching %>% ungroup() %>% 
  filter(!date %in% discarded_dates) %>%
  filter(!discard_arrested) %>% 
  filter(str_detect(condition, '^switch_[0-9]+h$')) %>%
  filter(switch_idx==1) %>%
  # mutate(lag=ifelse(lag<240*60, lag, 245*60),
         # lag=ifelse(is.na(lag), -10, lag)) %>% 
  ggplot(aes(x=lag_200/60)) +
   geom_density(stat='bin', fill=qual_cols[2], col='transparent', alpha=.2, binwidth=3) +
   geom_freqpoly(col=qual_cols[2], binwidth=3, show.legend=FALSE) +
   labs(x=lac_lags_label) +
   coord_cartesian(xlim=c(-10, 240), expand=FALSE) +
   NULL
) +
  labs(subtitle='naive cells (9 independent experiments)')

```

```{r}
mycells_switching %>% ungroup() %>% 
  filter(!date %in% discarded_dates) %>%
  filter(!discard_arrested) %>% 
  filter(str_detect(condition, '^switch_[0-9]+h$')) %>%
  filter(switch_idx==1) %>%
  mutate(
    lag_type=lag_200,
    lag_type=as.character(lag_type), lag_type=ifelse(is.na(lag_type), 'NA', lag_type),
    lag_type=ifelse(is.finite(as.numeric(lag_type)) & as.numeric(lag_type)<50*60, 'short lags', lag_type),
    lag_type=ifelse(is.finite(as.numeric(lag_type)) & as.numeric(lag_type)>=50*60, 'long lags', lag_type) ) %>%
  group_by(lag_type) %>% 
  summarise(n=n(), m=mean(lag_200/60), sd=sd(lag_200/60)) %>% 
  ungroup() %>% mutate(p=n/sum(n), p_se=sqrt(p*(1-p)/n)) %>% 
  knitr::kable(digits=3)

```


```{r eval=FALSE}
mycells_switching %>% ungroup() %>% 
  filter(!date %in% discarded_dates) %>%
  filter(!discard_arrested) %>% 
  filter(str_detect(condition, '^switch_[0-9]+h$')) %>%
  filter(switch_idx==1) %>%
  gather(variable, lag, growth_lag, gfp_lag, lag_200, factor_key=TRUE) %>% 
  mutate(lag=ifelse(lag<240*60, lag, 245*60),
         lag=ifelse(is.na(lag), -10, lag)) %>% 
  ggplot(aes(x=lag/60)) +
  facet_grid(variable~.) +
  geom_freqpoly(aes(col=factor(switch_idx)), binwidth=3, position='identity') +
  labs(x='lag after the switch (min)', subtitle='naive cells', col='switch', fill='switch') +
  # xlim(-1, 240) +
  guides(col='none', fill='none') +
  NULL

```

```{r eval=FALSE}
 mycells_switching %>% ungroup() %>% 
  filter(!date %in% discarded_dates) %>%
  filter(!discard_arrested) %>% 
  filter(str_detect(condition, '^switch_[0-9]+h$')) %>%
  filter(switch_idx==1) %>%
   group_by(date) %>% summarise(n=n()) %>% 
   ungroup() %>% summarise(n_date=n(), n_cells=sum(n))

```


Although the distributions vary from day to day, the bimodality is obserevd in all replicates (particularly visible in the panel with overlayed distributions).

```{r fig.height=10}
(myplots[['naive_lags_per_expt']] <- function(na.rm=TRUE)
  mycells_switching %>% ungroup %>% 
  filter(!date %in% discarded_dates) %>%
  filter(!discard_arrested) %>% 
  filter(str_detect(condition, '^switch_[0-9]+h$')) %>%
  filter(switch_idx==1) %>%
  mutate(lag=lag_200,
         lag=ifelse(is.finite(lag) & lag>240*60, 240*60, lag), 
         lag=ifelse(!is.infinite(lag), lag, max(lag[is.finite(lag)])+60),
         lag=ifelse(!na.rm & is.na(lag), -1, lag)) %>% 
  # (function(.df) bind_rows(.df, .df %>% mutate(date=NA, condition='all')) ) %>% 
  (function(.df)
    ggplot(.df, aes(x=lag/60)) +
     facet_wrap(~condition+date, ncol=2, dir='v',
                strip.position='right', labeller=as_labeller(rename_conds)) +
     # geom_freqpoly(aes(col=lag/60 > 236), binwidth=6, show.legend=FALSE) +
     geom_freqpoly(breaks=seq(0, 240, 6)-3, pad=TRUE, show.legend=FALSE) + 
     geom_freqpoly(aes(col=factor(date_col)), breaks=seq(0, 240, 6)-3, pad=TRUE, show.legend=FALSE,
                   data=.df %>% mutate(date_col=date, date=NA, condition='all')) +
     labs(x=lac_lags_label) +
     coord_cartesian(xlim=c(-10, 240), expand=FALSE) +
     scale_colour_periodic() +
     NULL
  )
 )(FALSE) +
  labs(subtitle='naive cells')

```


Noticeably a small fraction of cells don't induce even after 4h; here we check that this fraction is not higher in bottom cells (otherwise it would indicate that this fraction is overall underestimated since bottom cells don't have the bias of long lags being more truncated).

```{r}
mycells_switching %>% ungroup() %>% 
  filter(!date %in% discarded_dates) %>%
  filter(!discard_arrested) %>% 
  filter(str_detect(condition, '^switch_[0-9]+h$') ) %>%
  filter(switch_idx==1) %>%
  (function(.df) bind_rows(
    .df %>% mutate(sample='all cells'),
    .df %>% filter(cell_num_from_bottom==0) %>% mutate(sample='bottom cells') )) %>% 
  mutate(lag_200=as.character(lag_200), lag_200=ifelse(is.na(lag_200), 'NA', lag_200),
         lag_200=ifelse(is.finite(as.numeric(lag_200)) & as.numeric(lag_200)<50*60, 'short lags', lag_200),
         lag_200=ifelse(is.finite(as.numeric(lag_200)) & as.numeric(lag_200)>=50*60, 'long lags', lag_200) ) %>%
  group_by(sample, lag_200) %>% 
  summarise(n=n()) %>% 
  group_by(sample) %>% mutate(p=n/sum(n), p_se=sqrt(p*(1-p)/n)) %>% 
  rename("lag type"=lag_200) %>% 
  knitr::kable(digits=3)

```


Using a gradual transition for glucose to lactose over 40 minutes, we verify that the bimodal distribution of *lac* induction lags is not a consequence of the abrupt change.  
NB: since the concentration at which cells start to sense the low glucose / lactose is unknown, the lag in the gradual transition experiments are shifted so as to match the value of the step condition.

<!-- in order to have a not-too-short period with low glucose (> 0) and lactose, cells are switched from 0.002% glucose (≈100µM) to 0.2% lactose. CRP is supposed to kick in around 20-50µM (cf T. Hwa) -->


```{r}
(myplots[['lags_hist_ramp']] <- 
# xxx add n per condition
   mycells_switching %>% ungroup() %>% 
  filter(!date %in% discarded_dates) %>%
  filter(!discard_arrested) %>% 
  filter(switch_idx==1) %>%
  filter(str_detect(condition, '^switch_[0-9]+h$') | condition=='switch_ramp40min') %>% 
  mutate(condition=ifelse(str_detect(condition, '^switch_[0-9]+h$'), 'step (control)', 'ramp (40 min)'),
         condition=fct_relevel(condition, 'step') ) %>% 
  mutate(lag=lag_200, 
         # lag=ifelse(lag>240*60, Inf, lag),
         # lag=ifelse(!is.infinite(lag), lag, 241*60),
         # lag=ifelse(is.na(lag), -200, lag)
         ) %>% 
  ggplot(aes(x=lag/60)) +
  geom_freqpoly(aes(y=..density.., col=condition), binwidth=6, position='identity') +
  labs(x=lac_lags_label) +
  # xlim(0, 240) +
  # theme(strip.background=element_rect(fill='gray95', colour='transparent'),
  #       legend.key.size=unit(.8, "lines") ) +
  theme(legend.position=c(1, 1), legend.justification=c(1, 1) ) +
  coord_cartesian(xlim=c(0, 240)) +
  NULL
)

```


### Effect of LacZ-GFP fluorescence at the switch

```{r}
mycells_switching %>% ungroup %>% 
  filter(!date %in% discarded_dates) %>%
  filter(str_detect(condition, '^switch_[0-9]+h$')) %>%
  filter(switch_idx==1, !is.na(lag_200)) %>% 
  mutate(type=ifelse(lag_200/60<50, 'short', 'long')) %>% 
  ggplot(aes(gfp_ini, ..density..)) + 
  facet_grid(date~.) +
  geom_freqpoly(aes(col=type), binwidth=10) +
  NULL
  
```

```{r eval=FALSE}
myframes %>% ungroup %>% 
  filter(condition=='mg1655', length_um) %>% 
  ggplot(aes(length_um, gfp_nb)) + 
  geom_density2d() +
  geom_point(size=0.2, alpha=.1, stroke=0) +
  coord_cartesian(xlim=c(1, 4), ylim=c(-100, 150)) +
  NULL

mycells_switching %>% ungroup %>% 
  filter(!date %in% discarded_dates) %>%
  filter(str_detect(condition, '^switch_[0-9]+h$')) %>%
  filter(switch_idx==1, !is.na(lag_200)) %>% 
  mutate(type=ifelse(lag_200/60<50, 'short', 'long')) %>% 
  ggplot(aes(length_ini, gfp_ini)) + 
  facet_grid(date~.) +
  # geom_density2d() +
  geom_point(size=1, alpha=.4, stroke=0) +
  NULL

mycells_switching %>% ungroup %>% 
  filter(!date %in% discarded_dates) %>%
  filter(str_detect(condition, '^switch_[0-9]+h$')) %>%
  filter(switch_idx==1, !is.na(lag_200)) %>% 
  mutate(type=ifelse(lag_200/60<50, 'short', 'long')) %>% 
  group_by(date) %>%
  mutate(gr=logl_time_slope_before-mean(logl_time_slope_before),
         gfp_ini=gfp_ini-mean(gfp_ini)) %>% 
  ggplot(aes(gfp_ini)) + 
  # facet_grid(date~.) +
  # geom_freqpoly(aes(y= ..density.., col=type), binwidth=.00001) +
  stat_ecdf(aes(col=factor(date))) +
  scale_color_hue() +
  NULL
  


```


```{r}
(myplots[['lags_gfp_scatter']] <- function(.fit=FALSE)
  mycells_switching %>% ungroup %>% 
  filter(!date %in% discarded_dates) %>%
  filter(str_detect(condition, '^switch_[0-9]+h$')) %>%
  filter(switch_idx==1, !is.na(lag_200)) %>% 
  mutate(type=ifelse(lag_200/60<50, 'short', 'long')) %>% 
  group_by(date) %>%
  mutate(gfp_ini=(gfp_ini-mean(gfp_ini))) %>%
  ggplot(aes(gfp_ini, lag_200/60)) + 
  geom_hline(yintercept = 50, lty='dashed') +
  geom_point(aes(col=type), size=1, alpha=.2, stroke=0) +
  (if (.fit) stat_smooth(method=MASS::rlm, col='black', fullrange = TRUE)) +
  labs(x='fluorescence at the switch\n(centered per experiment; equivalent GFP molecules)', y=lac_lags_label) +
  guides(col=guide_legend(override.aes = list(size=2, alpha=1))) +
  coord_cartesian(xlim=c(-60, 120)) +
  NULL
 )()

```


```{r}
(myplots[['lags_gfp_diff_cdf']] <- mycells_switching %>% ungroup %>% 
  filter(!date %in% discarded_dates) %>%
  filter(str_detect(condition, '^switch_[0-9]+h$')) %>%
  filter(switch_idx==1, !is.na(lag_200)) %>% 
  mutate(type=ifelse(lag_200/60<50, 'short', 'long')) %>% 
  select(date, type, gfp_ini) %>% 
  bind_rows(myframes %>% ungroup %>% 
              filter(condition=='mg1655') %>%
              filter(!discard_start, !discard_top) %>% 
              filter(between(length_um, 2, 3.5)) %>% 
              select(date, gfp_ini=gfp_nb) %>% mutate(type='control'), .) %>% 
  group_by(date) %>% 
  mutate(gfp_ini=(gfp_ini-mean(gfp_ini))) %>%
  ggplot(aes(gfp_ini, col=type)) + 
  # facet_grid(date~.) +
  stat_ecdf() +
  # geom_freqpoly(aes(y=..density..), binwidth=10) +
  scale_color_manual(name='type of lac\ninduction lag', breaks=c('control', 'long', 'short'), values=c('gray50', qual_cols[1], qual_cols[2])) +
  labs(x='fluorescence at the switch\n(centered per experiment; equivalent GFP molecules)', y='cumulative probability') +
  coord_cartesian(xlim=c(-60, 120)) +
  NULL)
  
```

```{r eval=FALSE}
mycells_switching %>% ungroup %>% 
  filter(!date %in% discarded_dates) %>%
  filter(str_detect(condition, '^switch_[0-9]+h$')) %>%
  filter(switch_idx==1, !is.na(lag_200)) %>% 
  mutate(type=ifelse(lag_200/60<50, 'short', 'long')) %>% 
  group_by(date) %>% 
  mutate(gfp_ini=(gfp_ini-mean(gfp_ini))) %>%
  ungroup %>% 
  summarise(n=n(), mean=mean(gfp_ini), sd=sd(gfp_ini))

myframes %>% ungroup %>% 
  filter(condition=='mg1655') %>%
  filter(!discard_start, !discard_top) %>% 
  filter(between(length_um, 2, 3.5)) %>% 
  summarise(n=n(), mean=mean(gfp_nb), sd=sd(gfp_nb))

mycells_switching %>% ungroup %>% 
  filter(!date %in% discarded_dates) %>%
  filter(str_detect(condition, '^switch_[0-9]+h$')) %>%
  filter(switch_idx==1, !is.na(lag_200)) %>% 
  mutate(type=ifelse(lag_200/60<50, 'short', 'long')) %>% 
  group_by(date) %>% 
  mutate(gfp_ini=(gfp_ini-mean(gfp_ini))) %>%
  # with(t.test(gfp_ini[type=='short'], gfp_ini[type=='long'], var.equal=TRUE))
  group_by(date, type) %>% summarise(n=n()) %>%
  spread(type, n)

```


```{r eval=FALSE}
# compare distributions for several traits
mycells_switching %>% ungroup %>% 
  filter(!date %in% discarded_dates) %>%
  filter(str_detect(condition, '^switch_[0-9]+h$')) %>%
  filter(switch_idx==1, !is.na(lag_200)) %>% 
  mutate(type=ifelse(lag_200/60<50, 'short', 'long')) %>%
  left_join(myframes_switching %>% ungroup %>%  #filter(ugen %in% unique(ugen)[1:5]) %>% 
              filter(time_sec<t_lac_switch) %>% 
              # ggplot() + geom_line(aes(time_sec-t_lac_switch, length_um, lty=type, col=ugen)) +
              # scale_colour_periodic_brewer(guide='none')
              group_by(ugen) %>% 
              filter(time_sec==min(time_sec), time_sec>0) %>% 
              select(ugen, length_um_birth=length_um) ) %>% 
  mutate(cc=(t_lac_switch-time_birth) / 
           # average growth rate in glucose
           mycells_constant %>% filter(condition=='glucose') %>% 
           mutate(dt=time_div-time_birth) %>% pull(dt) %>% mean,
         dl=length_ini - length_um_birth) %>% 
  select(date, type, gfp_ini, growth_rate=logl_time_slope_before, cc, dl) %>% 
  bind_rows(myframes %>% ungroup %>%
              filter(condition=='mg1655') %>%
              filter(!discard_start, !discard_top) %>%
              filter(between(length_um, 2, 3.5)) %>%
              select(date, gfp_ini=gfp_nb) %>% mutate(type='control'), .) %>%
  bind_rows(mycells_constant %>% ungroup %>%
              filter(condition=='glucose') %>%
              select(date, growth_rate=logl_time_slope) %>% mutate(type='control'), .) %>%
  group_by(date) %>% 
  mutate(gfp_ini=(gfp_ini-mean(gfp_ini)),
         # growth_rate=growth_rate-mean(growth_rate),
         ) %>%
  gather(variable, value, -date, -type) %>% 
  group_by(date, variable) %>% 
  filter(between(value, quantile(value, 0.01, na.rm=TRUE), quantile(value, 0.99, na.rm=TRUE))) %>% 
  # ungroup() %>% mutate(type=ifelse(type!='control', 'expt', type)) %>% 
  ggplot(aes(value, col=type)) + 
  facet_wrap(~variable, ncol = 1, scales='free_x') +
  # facet_grid(date~variable, scales='free_x') +
  stat_ecdf() +
  # geom_freqpoly(aes(y=..density..), binwidth=10) +
  scale_color_manual(name='type of lac\ninduction lag', breaks=c('control', 'long', 'short', 'expt'), values=c('gray50', qual_cols[1], qual_cols[2], qual_cols[3])) +
  labs(y='cumulative probability') +
  # coord_cartesian(xlim=c(-60, 120)) +
  NULL

```

```{r}
(myplots[['naive_lags_correl']] <- mycells_switching %>% ungroup %>% 
  filter(!date %in% discarded_dates) %>%
  filter(str_detect(condition, '^switch_[0-9]+h$')) %>%
  filter(switch_idx==1, !is.na(lag_200)) %>% 
  mutate(type=ifelse(lag_200/60<50, 'short', 'long')) %>%
  left_join(myframes_switching %>% ungroup %>%  #filter(ugen %in% unique(ugen)[1:5]) %>% 
              filter(time_sec<t_lac_switch) %>% 
              # ggplot() + geom_line(aes(time_sec-t_lac_switch, length_um, lty=type, col=ugen)) +
              # scale_colour_periodic_brewer(guide='none')
              group_by(ugen) %>% 
              filter(time_sec==min(time_sec), time_sec>0) %>% 
              select(ugen, length_um_birth=length_um) ) %>% 
  mutate(cc=(t_lac_switch-time_birth) / 
           # average growth rate in glucose
           ( mycells_constant %>% filter(condition=='glucose') %>% 
           mutate(dt=time_div-time_birth) %>% pull(dt) %>% mean ),
         dl=length_ini - length_um_birth,
         growth_rate=logl_time_slope_before/log(2)*3600) %>% 
  select(date, lag_200, type, gfp_ini, growth_rate, cc, dl) %>% 
  # group_by(date) %>% 
  # mutate(gfp_ini=(gfp_ini-mean(gfp_ini))) %>%
  filter(cc < 1.5) %>% 
  gather(variable, value, -date, -lag_200, -type) %>%
  group_by(variable) %>% 
  filter(between(value, quantile(value, 0.01, na.rm=TRUE), quantile(value, 0.99, na.rm=TRUE))) %>% 
  (function(.df)
    ggplot(.df, aes(value, lag_200/60)) + 
     facet_wrap(~variable, nrow=1, scales='free_x', strip.position='bottom',
                labeller=as_labeller(function(.str) {
                  .str <- str_replace(.str, "cc", "cell cycle")
                  .str <- str_replace(.str, "dl", "added length\n(µm)")
                  .str <- str_replace(.str, "gfp_ini", "LacZ-GFP")
                  .str <- str_replace(.str, "growth_rate", "growth rate\n(dbl/h)")
                  return(.str)
                })) +
     geom_hline(yintercept = 50, lty='dashed') +
     geom_point(aes(col=type), size=1, alpha=.2, stroke=0) +
     # stat_smooth(method=MASS::rlm, col='black', fullrange = TRUE) + 
     geom_text(aes(-Inf, 0, label=r2str), size=5, hjust=-0.1, vjust=0, parse=TRUE,
               data=.df %>% filter(is.finite(lag_200)) %>% group_by(variable) %>% summarise(r2=cor(value, lag_200)^2) %>% mutate(r2str=sprintf("r^2 == '%.4f'", round(r2, digits=4)))) +
     scale_x_continuous(breaks = scales::pretty_breaks(n = 3)) +
     expand_limits(y=0) +
     guides(col=guide_legend(override.aes = list(size=2, alpha=1))) +
     labs(y=lac_lags_label, col=expression(paste("type of ", italic("lac"), " induction lag (min)"))) +
     theme(legend.position = 'top',
           strip.placement='outside', strip.background.x=element_blank(), 
           axis.title.x=element_blank(), strip.text.x=element_text(size=rel(1.2), vjust = 1)) +
     NULL )
)

```


### Heritability of induction lag in naive cells

Heritability of the lag at the first switch:

```{r compute cells pairs}
mypairs_switching <- mycells_switching %>%
  group_by(condition, date, pos, gl, switch_idx) %>%
  filter(n() > 1) %>%
  do((function(.df) {
    # browser()
    return( combn(.df$genealogy, 2, simplify=FALSE) %>%
      lapply(function(.x) as.data.frame(genealogy_relationship(.x), stringsAsFactors=FALSE)) %>%
      do.call(rbind, .) )
  })(.)) %>%
  # rbind one df per variable type
  mutate(foo=1) %>% left_join(data.frame(foo=1, variable=c('growth_lag', 'gfp_lag', 'lag_200'))) %>% select(-foo) %>%
  # join variables of interest
  mutate(rel=factor(rel, levels=c("sisters", "cousins1", "cousins2", "niece"))) %>%
  left_join(mycells_switching %>% ungroup %>%
              gather(variable, value, growth_lag, gfp_lag, lag_200) %>%
              setNames(., paste(names(.), "1", sep="_")) %>%
              rename(date=date_1, pos=pos_1, gl=gl_1, switch_idx=switch_idx_1, variable=variable_1) ) %>%
  left_join(mycells_switching %>% ungroup %>%
              gather(variable, value, growth_lag, gfp_lag, lag_200) %>%
              setNames(., paste(names(.), "2", sep="_")) %>%
              rename(date=date_2, pos=pos_2, gl=gl_2, switch_idx=switch_idx_2, variable=variable_2) )

```

```{r}
mypairs_switching %>%
  # group_by(rel, div_1, div_2) %>% summarise(n())
  ungroup() %>%
  filter(str_detect(condition, '^switch_[0-9]+h$') ) %>%
  filter(switch_idx==1) %>%
  filter(rel!='niece') %>%
  filter(variable=='lag_200') %>%
  filter(!is.na(value_1), !is.na(value_2)) %>%
  select(rel, value_1, value_2) %>% 
  # append pairs of independent lags
  bind_rows(
    # extract lag values
    mycells_switching %>% ungroup %>% 
      filter(!date %in% discarded_dates) %>%
      filter(str_detect(condition, '^switch_[0-9]+h$')) %>%
      filter(switch_idx==1, !is.na(lag_200)) %>% 
      pull(lag_200) %>% 
      # draw random pairs
      (function(.l)
        tibble(rel='indep', #variable='lag_200', 
               value_1=quantile(.l, runif(length(.l))), 
               value_2=quantile(.l, runif(length(.l))) )
      ) ) %>% 
  mutate(type_1=ifelse(value_1 < 50*60, 's', 'l'),
         type_2=ifelse(value_2 < 50*60, 's', 'l'),
         type=paste0(type_1, type_2), type=ifelse(type=='ls', 'sl', type) ) %>%
  mutate(rel=fct_relevel(rel, 'indep', 'sisters')) %>% 
  group_by(rel, type) %>% summarise(n=n()) %>%
  group_by(rel) %>% mutate(p=n/sum(n)) %>%
  ggplot(aes(rel, type)) +
  geom_point(aes(size=p)) +
    # geom_tile(aes(fill=p)) +
    # scale_fill_viridis_c() +
    # scale_fill_viridis_c(limits=c(0, 1)) +
    NULL

```


## Memory of lag between switches

At the second switch, we observe that there is no effect of the lag of the ancestor at the first switch.

```{r}
mypairs_memory <- mycells_switching %>% 
  group_by(condition, date, pos, gl) %>% 
  do((function(.df_gl){
    # browser()
    # if(.df_gl$ugen[1]=='20150703.4.15.0:BBBBBBB') browser()
    .geneals_sw1 <- .df_gl %>% filter(switch_idx==1) %>% with(genealogy)
    if (nrow(.df_gl %>% filter(switch_idx==2)) == 0) return(data.frame())
    .df_gl %>% filter(switch_idx==2) %>% 
      select(genealogy_2=genealogy, lag_2=lag_200, discard_arrested_2=discard_arrested, 
             logl_time_slope_pre_2=logl_time_slope_pre) %>% group_by(genealogy_2) %>% 
      # with(which_is_parent_cid(genealogy, .geneals_sw1))
      mutate(genealogy_1=which_is_parent_cid(genealogy_2, .geneals_sw1),
             genealogy_1=ifelse(genealogy_1>0, .geneals_sw1[genealogy_1], NA) )
    })(.)) %>% 
  left_join(mycells_switching %>% ungroup %>% 
              select(date, pos, gl, genealogy_1=genealogy, lag_1=lag_200, discard_arrested_1=discard_arrested,
                     logl_time_slope_pre_1=logl_time_slope_pre) )

(myplots[['lag_memory']] <- 
    mypairs_memory %>% 
    filter(!date %in% discarded_dates) %>%
    filter(str_detect(condition, '^switch_[0-9]+h$') | condition=='switch_long_lac') %>%
    filter(!discard_arrested_1, !discard_arrested_2) %>% 
    # filter(logl_time_slope_pre_1 > min_growth_rate, logl_time_slope_pre_2 > min_growth_rate) %>% 
    ggplot() +
    facet_wrap(~condition, strip.position='right', labeller=as_labeller(rename_conds)) +
    # geom_abline(lty='dotted', alpha=.5) +
    geom_hline(yintercept = 50, lty='dotted', alpha=.5) +
    geom_point(aes(lag_1/60, lag_2/60), alpha=.2, stroke=0) +
    expand_limits(x=0, y=0) +
    labs(x="ancestor's lag at first switch (min)", y="lag at second switch (min)") +
    NULL
)

(myplots[['lag_memory_aggreg']] <- 
    mypairs_memory %>% 
    filter(!date %in% discarded_dates) %>%
    filter(str_detect(condition, '^switch_[0-9]+h$') | condition=='switch_long_lac') %>%
    filter(!discard_arrested_1, !discard_arrested_2) %>% 
    group_by(condition, lag_1, lag_2) %>% 
    summarise(n=n()) %>% 
    (function(.df)
      ggplot(.df) +
       facet_wrap(~condition, strip.position='right', labeller=as_labeller(rename_conds)) +
       # geom_abline(lty='dotted', alpha=.5) +
       geom_hline(yintercept = 50, lty='dotted', alpha=.5) +
       geom_point(aes(lag_1/60, lag_2/60, size=n)) +
       scale_size_continuous(range=c(1.2/max(.df$n), 1.2), breaks=c(2, 5, 8)) +
       expand_limits(x=0, y=0) +
       labs(x="ancestor's lag at first switch (min)", y="lag at second switch (min)") +
       NULL
    )
)

```


This is also true when looking only at bottom cells (which are present at both switch, or which might be different in case of molecular deterimnants of the lag would be accumulated at the old cell pole), or when looking at the average in each growth channel (since lags are highly correlated within a lineage) <!-- cite Kaiser2018 -->

```{r}
(myplots[['lag_memory_bottom']] <- 
   mycells_switching %>% ungroup() %>% 
   filter(!date %in% discarded_dates) %>%
   filter(str_detect(condition, '^switch_[0-9]+h$')) %>%
   filter(!discard_arrested) %>% 
   filter(cell_num_from_bottom==0) %>% # bottom_cell
   gather(variable, lag, growth_lag, gfp_lag, lag_200, factor_key=TRUE) %>% 
   select(condition, date, pos, gl, switch_idx, variable, lag) %>% 
   mutate(name_col=paste0('lag_', switch_idx), switch_idx=NULL) %>% 
   spread(name_col, lag) %>%
   filter(variable=='lag_200') %>% 
   ggplot() +
   facet_wrap(~condition+date, labeller=as_labeller(rename_conds), strip.position='right') +
   # geom_abline(lty='dotted', alpha=.5) +
   geom_hline(yintercept = 50, lty='dotted', alpha=.5) +
   geom_point(aes(lag_1/60, lag_2/60), alpha=.5, stroke=0) +
   scale_x_continuous(breaks=seq(0, 1000, by=100)) +
   scale_y_continuous(breaks=seq(0, 1000, by=100)) +
   expand_limits(x=0, y=0) +
   labs(x='lag at first switch (min)', y='lag at second switch (min)') +
   NULL
) +
  labs(subtitle='bottom cells')

```

```{r}
(myplots[['lag_memory_avg']] <- 
   mycells_switching %>% ungroup() %>% 
   filter(!date %in% discarded_dates) %>%
   filter(!discard_arrested) %>% 
   filter(str_detect(condition, '^switch_[0-9]+h$')) %>%
   gather(variable, lag, growth_lag, gfp_lag, lag_200, factor_key=TRUE) %>% 
   filter(is.finite(lag)) %>% 
   # compute stats per GL, per switch, per variable
   group_by(condition, date, pos, gl, variable, switch_idx) %>% 
   summarise(mean_lag=mean(lag)) %>% 
   mutate(name_col=paste0('mean_lag_', switch_idx), switch_idx=NULL) %>% 
   spread(name_col, mean_lag) %>%
   filter(variable=='lag_200') %>% 
   ggplot(aes(mean_lag_1/60, mean_lag_2/60)) +
   facet_wrap(~condition+date, labeller=as_labeller(rename_conds), strip.position='right') +
   # facet_grid(condition~variable, labeller=as_labeller(rename_conds)) +
   # geom_abline(lty='dotted', alpha=.5, stroke=0) +
   geom_hline(yintercept = 50, lty='dotted', alpha=.5) +
   geom_point(alpha=.4) +
   scale_x_continuous(breaks=seq(0, 1000, by=100)) +
   labs(x='average lag at first switch (min)', y='average lag at second switch (min)') +
   NULL
) +
  labs(subtitle='average in growth channel')

```


## Effect of Lac proteins inheritance at the second switch

In contrastm, we can analyse the impact of LacZ-GFP molecules inherited from the ancestor's induction during the first switch. 

First, let's look at the distribution of induction lags at the second switch. The medians are ordered according to the delay between the two lactose exposures (which supports the idea of a molecular memory). It takes surprising long (16 to 20h i.e. 12 to 16 divisions) to relax to the lag distribution observed in naive cells.

```{r}
(myplots[['memory_cdfs_facets']] <- 
   mycells_switching %>% ungroup() %>% 
   filter(!date %in% discarded_dates) %>%
   filter(!discard_arrested) %>% 
   filter(str_detect(condition, '^switch_[0-9]+h$') | condition=='switch_late') %>%
   mutate(lag=lag_200,
          lag=ifelse(!is.infinite(lag), lag, max(lag[is.finite(lag)])+60),
          lag=ifelse(is.na(lag), -1, lag)) %>% 
   filter(switch_idx %in% 1:2) %>% 
   ggplot(aes(x=lag/60)) +
   facet_grid(switch_idx~., labeller=as_labeller(function(.str) paste0('switch: ', .str)) ) +
   stat_ecdf(aes(y=1-..y.., col=condition, group=date), geom='line') +
   labs(x=lac_lags_label, y='reverse cumulative probability', col='condition') +
   scale_color_hue() +
   scale_color_manual(name='delay between\nlactose exposure',
                      limits=c(sprintf('switch_%02dh', seq(4, 24, 4)), 'switch_late'),
                      labels=c(sprintf('%dh', seq(4, 24, 4)), 'control (late)'),
                      values=c(scales::viridis_pal()(6), 'gray70') ) +
   NULL
)

```

```{r}
(myplots[['memory_cdfs']] <-
   mycells_switching %>% ungroup() %>% 
   filter(!date %in% discarded_dates) %>%
   filter(!discard_arrested) %>%
   filter(str_detect(condition, '^switch_[0-9]+h$') | condition=='switch_late') %>%
   mutate(lag=lag_200,
          lag=ifelse(lag>240*60, Inf, lag),
          # lag=ifelse(!is.infinite(lag), lag, max(lag[is.finite(lag)])+1e6),
          # lag=ifelse(is.na(lag), -1e6, lag),
          ) %>% 
   # group_by(condition, switch_idx) %>% summarise(n())
   (function(.df)
     ggplot(data=NULL, aes(x=lag/60, y=1-..y.., col=condition)) +
      stat_ecdf(aes(col='control', lty='dotted'), geom='line', size=rel(0.75),
                data=filter(.df, switch_idx==1, condition != 'switch_late')) +
      stat_ecdf(aes(lty='dotted'), geom='line', size=rel(0.75),
                data=filter(.df, switch_idx==1, condition == 'switch_late')) +
      stat_ecdf(aes(lty='solid'), geom='line', #size=.3,
                data=.df %>% filter(switch_idx==2)) +
      geom_rect(aes(x=NULL), xmin=-Inf, xmax=0, ymin=-Inf, ymax=Inf, fill='white', col='transparent',
                data=data.frame()) +
      scale_color_manual(name='delay between\nlactose exposures',
                         limits=c(sprintf('switch_%02dh', seq(4, 24, 4)), 'control', 'switch_late'),
                         labels=c(sprintf('%dh', seq(4, 24, 4)), 'control', 'control (late)'),
                         values=c(scales::viridis_pal()(6), 'gray40', 'gray70') ) +
      coord_cartesian(xlim=c(-10, 240), ylim=c(-.01, 1.01), expand=FALSE) +
      labs(x=lac_lags_label,  y='reverse cumulative\nprobability') +
      scale_linetype_manual(values=c('dotted'='dotted', 'solid'='solid'), guide=FALSE) +
      guides(col=guide_legend(ncol=2)) +
      theme_cowplot_legend_inset() +  
      theme(axis.line.x=element_line(), axis.line.y=element_line(),
            strip.background=element_rect(fill='gray95', colour='transparent'), 
            legend.position = c(1, 1), legend.justification = c(1,1),
            legend.margin=margin()
            # legend.key.height=unit(0.7, "lines")
      ) +
      NULL
   ))

# mycells_switching %>% ungroup() %>% 
#   filter(!date %in% discarded_dates) %>%
#   filter(discard_arrested) %>% 
#   filter(str_detect(condition, '^switch_[0-9]+h$') | condition=='switch_late') %>%
#   ggplot(aes()) +
#   geom_freqpoly(
  
  
```

In order to estimate the inherited LacZ-GFP of a given cell at the second switch, we compute the maximum LacZ-GFP level of its ancestor (at the end of the first lactose exposure) and account for the dilution given the number of division until the second switch (assuming even partitioning at each division).

```{r}
mycells_switching_memory <- mycells_switching %>% ungroup() %>% 
  filter(!date %in% discarded_dates) %>%
  filter(!discard_arrested) %>% 
  filter(str_detect(condition, '^switch_[0-9]+h$')) %>%
  filter(switch_idx==2) %>% 
  select(date, pos, gl, ugen, genealogy) %>% 
  group_by(date, pos, gl, ugen, genealogy) %>%
  # do(parents=get_all_parents_cid(.$genealogy))
  do(tibble(parent=get_all_parents_cid(.$genealogy))) %>% 
  inner_join(
    myframes %>% ungroup %>% 
      filter(!date %in% discarded_dates) %>%
      filter(str_detect(condition, '^switch_[0-9]+h$')) %>%
      filter(medium=='lactose', m_cycle==1) %>%
      filter(abs(time_sec-(m_end-dt))<1) %>%
      select(date, pos, gl, parent=genealogy, parent_gfp=gfp_nb) %>%
      mutate(parent=str_replace(parent, ":", "")),
    by=c("date", "pos", "gl", "parent") ) %>% 
  mutate(divs_since_par = str_length(genealogy) - str_length(parent)) %>% 
  left_join(
    mycells_switching %>% ungroup %>% 
      filter(!date %in% discarded_dates) %>%
      filter(str_detect(condition, '^switch_[0-9]+h$')) %>% 
      filter(switch_idx==2),
    by=c("date", "pos", "gl", "genealogy") )
   
```

```{r}
(myplots[['memory_elapsed_divs']] <- 
mycells_switching_memory %>% 
  ggplot(aes(divs_since_par, ..density.., col=condition)) + 
  geom_freqpoly(binwidth=1) +
  scale_color_manual(name='delay between\nlactose exposure',
                     limits=c(sprintf('switch_%02dh', seq(4, 24, 4))),
                     labels=c(sprintf('%dh', seq(4, 24, 4))),
                     values=c(scales::viridis_pal()(6)) ) +
  labs(x="divisions since fully induced parent") +
  NULL
)

```


```{r}
(myplots[['lags_inherited_gfp']] <- 
   mycells_switching_memory %>% 
   ungroup %>% 
   mutate(gfp_inherit=parent_gfp/2^divs_since_par,
          gfp_inherit_bin=as.numeric(as.character(Hmisc::cut2(gfp_inherit, cuts=c(-Inf, 10^seq(-1.1, 3.8, .6)), levels.mean=TRUE))) ) %>% 
   gather(variable, lag, growth_lag, lag_200, factor_key=TRUE) %>% 
   filter(variable=='lag_200') %>% 
   (function(.df) 
     ggplot(.df, aes(gfp_inherit, lag/60)) + 
      # facet_grid(.~variable) +
      geom_violin(aes(group=factor(gfp_inherit_bin)), show.legend = FALSE) +
      geom_hline(yintercept = 50, lty='dashed') +
      # geom_segment(x=0, xend=2, y=100, yend=100, data=NULL) +
      geom_segment(aes(x=pmax(0, gfp_inherit-sqrt(gfp_inherit)), xend=gfp_inherit+sqrt(gfp_inherit),
                       yend=lag/60, col=condition), alpha=.05, show.legend=FALSE) +
      geom_point(aes(col=condition), pch=20, alpha=.1) +
      # stat_smooth(aes(group=1), method='loess', span=.2, col='gray60', alpha=.2) +
      # geom_pointrange(aes(gfp_inherit_bin, mean, ymin=mean-se, ymax=mean+se, size=n), 
      #                 data=.df %>% group_by(gfp_inherit_bin, variable) %>% summarise(n=n(), mean=mean(lag/60, na.rm=TRUE), se=sd(lag/60, na.rm=TRUE)/sqrt(n)),
      #                 show.legend = FALSE) +
      scale_color_manual(name='delay between\nlactose exposure',
                         limits=c(sprintf('switch_%02dh', seq(4, 24, 4))),
                         labels=c(sprintf('%dh', seq(4, 24, 4))),
                         values=c(scales::viridis_pal()(6)) ) +
      guides(col=guide_legend(override.aes=list(alpha=1, size=2))) +
      # scale_size_continuous(range=c(0.1, 1)) +
      scale_size_continuous(range=c(0.02, 0.2)) +
      scale_x_log10() +
      coord_cartesian(xlim=c(0.1, 1.8e3), ylim=c(0, 110)) +
      labs(x='inherited LacZ-GFP (molecules)', y=lac_lags_label) +
      NULL
   ))

```

Importantly, as long as a cell inherits more than a few molecules from the first exposure to lactose, it is granted to have a short lag at the second switch. This reveals that the threshold of Lac proteins required to be readily sensitive to lactose as an inducer is on the order of 1 molecule.

```{r}
(myplots[['memory_frac_short']] <- 
   mycells_switching_memory %>% ungroup() %>% 
   mutate(gfp_inherit=parent_gfp/2^divs_since_par,
          gfp_inherit_bin=as.numeric(as.character(Hmisc::cut2(gfp_inherit, cuts=c(-Inf, 10^seq(-1.1, 3.8, .3)), levels.mean=TRUE))) ) %>% 
   filter(!is.na(lag_200)) %>% 
   (function(.df) 
     ggplot() + 
      geom_hline(aes(yintercept=prop), lty='dashed',
                 data=mycells_switching %>% ungroup %>%
                   filter(!date %in% discarded_dates) %>%
                   filter(!discard_arrested) %>% 
                   filter(str_detect(condition, '^switch_[0-9]+h$')) %>%
                   filter(switch_idx==1) %>%
                   filter(!is.na(lag_200)) %>% 
                   summarise(n=n(), prop=sum(lag_200/60<50)/n)
      ) +
      geom_rect(aes(xmin=0, xmax=Inf, ymin=prop-sqrt(prop*(1-prop)/n), ymax=prop+sqrt(prop*(1-prop)/n)),
                fill='grey50', alpha=0.1,
                data=mycells_switching %>% ungroup %>% 
                  filter(!date %in% discarded_dates) %>%
                  filter(logl_time_slope_pre > min_growth_rate) %>% 
                  filter(str_detect(condition, '^switch_[0-9]+h$')) %>%
                  filter(switch_idx==1) %>% 
                  filter(!is.na(lag_200)) %>% 
                  group_by(date) %>% 
                  summarise(n=n(), prop=sum(lag_200/60<50)/n)
      ) +
      geom_pointrange(aes(gfp_inherit_bin, prop, size=n, ymin=prop-se, ymax=prop+se), 
                      data=.df %>% group_by(gfp_inherit_bin) %>% summarise(n=n(), prop=sum(lag_200/60<50)/n, se=sqrt(prop*(1-prop)/n))) +
      scale_size_continuous(range=c(0.05, .5), breaks=c(50, 150)) +
      scale_x_continuous(breaks=10^(0:3),
                         trans=scales::trans_new(name='log10_rev', 
                                                 transform=function(x) -log(x, 10), 
                                                 inverse=function(x) 10^(-x))
      ) +
      coord_cartesian(xlim=c(0.1, 3e3)) +
      expand_limits(y=0) +
      labs(x='inherited LacZ-GFP molecules', y='fraction of short lags', size='sample size') +
      guides(size=guide_legend(direction='horizontal')) +
      theme_cowplot_legend_inset() +
      theme(legend.position=c(0, 0), legend.justification=c(-0.1, -0.5), #'top', 
      legend.margin=margin(),
      ) +
      # theme(legend.position=c(0.99, 0.99), legend.justification=c(1, 1)) +
      NULL
   ))


```


