---
title: "Growth-coupled sensitivity to TMG during transient growth arrest"
author: Thomas Julou
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

```{r}
myscales[['lacl_gly']] <- scale_colour_discrete(
  limits=c("switch_lactulose_TMG20", "switch_lactulose", "switch_glycerol_TMG20", "> lactose",
           "switch_lactulose_TMG20_stdIllum", "switch_lactulose_stdIllum"),
  labels=rename_conds)
  
```

## Controls

Distribution of growth rate in the initial condition.

```{r}
(myplots[['gr_length_medians']] <- function()
  myframes %>% 
   ungroup() %>% 
   filter(str_detect(condition, 'switch_')) %>% 
   # # discard pos with low acq rate (phototoxicity control)
   # filter(!(date==20181008 & pos %in% 5:9), 
   #        !(date %in% c(20181009, 20181024) & pos %in% c(1,3,5,7,9))) %>% 
   filter(m_start == 0) %>% 
   group_by(condition, date, pos, gl, id, cell) %>% 
   filter(n() > 3) %>% 
   nest() %>% 
   mutate(
     time_birth = map_dbl(data, ~first(.$time_sec)),
     mod = map(data, ~lm( log(length_um)~time_sec, .)),
     logl = map(mod, ~predict(., se.fit=TRUE)),
     cc = pmap(list(mod, logl), ~tibble(
       n=..1$df.residual+1,
       loglength_start=first(..2$fit), loglength_startse=first(..2$se.fit), 
       loglength_end=last(..2$fit), loglength_endse=last(..2$se.fit), 
       logl_time_slope=..1$coefficients[2], logl_time_slopesd=summary(..1)$coefficients[2,2],
       logl_time_r2=summary(..1)$r.squared ) ),
   ) %>% 
   ungroup() %>% 
   mutate(
     illum = ifelse(str_detect(condition, "_stdIllum"), 'std', 'low'),
     condition=fct_relabel(factor(condition), rename_conds)
   ) %>% 
   filter(time_birth>2*3600 & time_birth<6*3600) %>% 
   unnest(cc) %>% 
   group_by(condition, illum, date) %>% 
   nest() %>% 
   mutate(ll=map(data, ~median_hilow(.$loglength_start, conf.int=0.5)),
          gr=map(data, ~median_hilow(.$logl_time_slope, conf.int=0.5)),
          type='analysed') %>% 
   unnest(c(ll, gr), names_sep="_") %>% 
   ggplot() +
   geom_point(aes(exp(ll_y), gr_y/log(2) * 3600, col=illum, alpha=type, shape=type)) +
   geom_errorbar(aes(exp(ll_y), ymin=gr_ymin/log(2) * 3600, ymax=gr_ymax/log(2) * 3600, col=illum, alpha=type), width=0) +
   geom_errorbarh(aes(exp(ll_y), gr_y/log(2) * 3600, xmin=exp(ll_ymin), xmax=exp(ll_ymax), col=illum, alpha=type))+
   scale_alpha_manual(values=c(1, 0.2)) +
   expand_limits(x=0, y=0) +
   labs(x='birth length (µm)', y='growth rate (doublings/h)') +
   NULL
)() +
  labs(subtitle='initial condition (between t=2h and t=6h)')

```

Position effect in the growth lane on induction lag: cells closer to the closed end tend to induce later, in particular for the slower inducing subpopulation. Since lactulose is expected to be in large excess, this raises the question whether TMG can be titrated by the first cells close to the GL opening...
Note that this effect is not critical for our analysis as we focus on the comparison of induction outcome between glycerol and lactulose rather than on its dynamics.


```{r}
mycells_switching %>% 
  filter(str_detect(condition, 'lactulose_TMG') ) %>% 
   # discard pos with low acq rate (phototoxicity control)
   filter(!(date==20181008 & pos %in% 5:9), 
          !(date %in% c(20181009, 20181024) & pos %in% c(1,3,5,7,9))) %>% 
  mutate(lag=lag_1000) %>% 
  filter(is.finite(lag)) %>% 
  group_by(condition, cell_num_from_bottom) %>% 
  filter(n() > 10) %>% 
  ggplot(aes(lag/60)) +
  facet_wrap(~condition, scales='free_y', labeller=as_labeller(rename_conds), ncol=1) +
  stat_ecdf(aes(y=1-after_stat(y), col=factor(cell_num_from_bottom)), alpha=.8) +
  expand_limits(x=0) +
  # scale_y_log10() +
  scale_color_viridis_d() +
  labs(x=lac_lags_label, y="rev. cumul. probability", col="rank")

```



## Decoupling growth and induction

In order to compare induction at a given inducer concentration in growing and arrested cells, we use lactulose. *E. coli* can metabolize lactulose when it expresses LacY and LacZ but lactulose is reported not to be an inhibitor of LacI. Hence an mixture of lactulose and TMG can be used to recapitulate the phenomenology of the induction by lactose where cells transiently stop growing (although at the concentration of nutrient and inducer can be decoupled).

The fraction of induced cells has been reported to increase abruptly with the (log of the) TMG concentration. We identify the upper range of TMG concentrations that are low enough that no cell induce after 5h when growing on glycerol. xxx
We predict that this concentration of TMG will be high enough for cells to induce and resume growth when switched to TMG and lactulose, since cells cannot use lactulose after the switch and are forced to transiently stop growing.


```{r cache=TRUE}
myrates_lactulose <- myframes %>% 
  ungroup %>% 
  filter(!discard_top) %>% 
  filter(str_detect(condition, 'TMG') | str_detect(condition, 'lactulose')) %>%
   # discard pos with low acq rate (phototoxicity control)
   filter(!(date==20181008 & pos %in% 5:9), 
          !(date %in% c(20181009, 20181024) & pos %in% c(1,3,5,7,9))) %>% 
  # loop on all cells
  group_by(condition, strain, date, pos, gl, id, cell) %>%
  partition(mycluster) %>%
  do( (function(.dfc, .rate_hw=3*360){
    .dfc %>% 
      group_by(time_sec) %>% 
      do((function(.p) {
        # browser()
        .df <- filter(.dfc, between(time_sec, .p[['time_sec']]-.rate_hw, .p[['time_sec']]+.rate_hw))
        .mod_ll <- lm(log(length_um)~time_sec, .df)
        # .mod_ll <- fastLmPure( cbind(1, .df$time_sec), log(.df$length_um) )
        .mod_g_l <- fastLmPure( cbind(1, .df$length_um), .df$gfp_nb )
        data.frame(irate_npoints=nrow(.df), igfp_nb=mean(.df$gfp_nb), 
                   iloglength=predict(.mod_ll, select(.p, time_sec)),
                   islope_ll_time=.mod_ll$coefficients[2],# islopesd_ll_time=.mod_ll$stderr[2],
                   ir2_ll_time=cor(.df$time_sec, log(.df$length_um), use="complete.obs")^2,
                   islope_g_l=.mod_g_l$coefficients[2], islopesd_g_l=.mod_g_l$stderr[2],
                   ir2_g_l=cor(.df$length_um, .df$gfp_nb, use="complete.obs")^2)
      })(.) )
  })(.) ) %>% 
  collect()

```


First let's look at some raw length traces, coloured using an arbitrary fluorescence concentration threshold (to visualise cells expressing LacZ-GFP at detectable levels). It reveals that some cells can grow on lactulose without TMG (and even sometimes express LacZ-GFP weakly); however, when exposed to lactulose and TMG cells express more LacZ-GFP and grow faster. The sustained growth on lactulose without TMG can be due to nutrients traces in lactulose (in particular galactose) or to undocumented induction by lactulose (e.g. if glycerol is present in the cytoplasm, galactosyl-glycerol will be produced by LacZ similarly to allolactose synthesis).

This constitutes a confounding effect since we need to distinguish the inducing effect of the TMG from the weak spontaneous induction observed in lactulose only.

In addition, after comparing the induction by TMG in cells growing on glycerol or lactulose using our default imaging protocol (using a 2x lower acquisition frequency, condition named "stdIllum"), we noticed that an unexpectedly high fraction of cells stopped growing and didn't induce in lactulose+TMG. We hypothesized that growth arrested cells were more sensitive to phototoxicity and repeated the experiments by reducing 5× the fluorescence illumination during the lactulose episode and switching cells back to glycerol in order to check whether arrested cells (if any) are still viable.


```{r fig.height=12}
(myplots[['lacl_time_series']] <-
   myframes %>% 
   ungroup() %>% 
   filter(time_sec > 2*3600, !discard_top) %>% 
   filter(str_detect(condition, 'TMG') | str_detect(condition, 'lactulose')) %>%
   # discard pos with low acq rate (phototoxicity control)
   filter(!(date==20181008 & pos %in% 5:9), 
          !(date %in% c(20181009, 20181024) & pos %in% c(1,3,5,7,9))) %>% 
   # sample 3 GLs per date
   group_by(date, pos, gl) %>% nest() %>% 
   group_by(date) %>% sample_n(3) %>% unnest(data) %>% 
   # filter(end_type == 'lost') %>% 
   ggplot(aes(time_sec - 2*3600, length_um, col=gfp_nb/length_um>250)) +
   facet_wrap(~ condition+date + pos+gl, dir="v", ncol=2,
              labeller=labeller(condition=as_labeller(rename_conds), .multi_line=FALSE)) +
   # geom_rect(aes(xmin=t_start - 2*3600, xmax=t_end - 2*3600, ymin=-Inf, ymax=Inf, x=NaN, y=NaN, group=1), fill=rgb(1, 0, 0, .1), col='red', data=filter(condition_ts, condition=='switch', medium=='lactose')) +
   geom_path(aes(group=ugen)) + # geom_point() +
   # scale_colour_periodic_brewer(guide='none') +
   # scale_alpha_discrete(range=c(0.1, 0.6)) +
   scale_x_hours() + #, limits=c(19*3600, 23*3600)) +
   scale_y_continuous(trans='log2') +
   labs(y='length (µm)', col='induced') +
   theme(legend.position="top")
)

```
 
```{r eval=FALSE}
myframes_switching %>% 
  filter(fit_lag) %>% 
  filter(str_detect(condition, 'TMG') | str_detect(condition, 'lactulose')) %>%
  # discard pos with low acq rate (phototoxicity control)
  filter(!(date==20181008 & pos %in% 5:9), 
         !(date %in% c(20181009, 20181024) & pos %in% c(1,3,5,7,9))) %>% 
  filter(type=='cell') %>% 
  group_by(condition, date, ugen) %>% 
  slice(n()) %>% 
  ggplot(aes(time_sec - 2*3600, gfp_nb, col=end_type, alpha=type)) +
  facet_wrap(~condition+date, dir="v", ncol=2, labeller=labeller(.cols=label_value, .multi_line=FALSE)) +
  # geom_rect(aes(xmin=t_start - 2*3600, xmax=t_end - 2*3600, ymin=-Inf, ymax=Inf, x=NaN, y=NaN, group=1), fill=rgb(1, 0, 0, .1), col='red', data=filter(condition_ts, condition=='switch', medium=='lactose')) +
  geom_point() + # geom_point() +
  # scale_colour_periodic_brewer(guide='none') +
  scale_alpha_discrete(range=c(0.1, 0.6)) +
  scale_x_hours() + #, limits=c(19*3600, 23*3600)) +
  ylim(-100, 4e3) +   expand_limits(x=0, y=0) +
  labs(y='total fluorescence (GFP molecules)') +
  theme(legend.position="top") +
  NULL

```


By increasing the induction threshold (+1000 LacZ-GFP molecules instead of +200 by default; picked according to raw traces), it is possible to measure induction lags specific for TMG in lactulose. These distributions indicate a large fraction of non inducing cells (see below for a more detailed analysis). Noticeably, using a lower induction threshold, an induction lag can be measured for more cells but at the cost of observing more cells induced by lactulose only (hence the induction lag doesn't correspond to the TMG specific induction anymore).

Importantly, the distribution of induction lags is not truncated and indicates that all induction has already taken place after 7h. In further analysis, we will compare LacZ-GFP level and instantaneous growth rate at this time.

```{r}
(myplots[['lags_hist_lacl_facets']] <-
   mycells_switching %>% 
   filter(str_detect(condition, 'TMG') | str_detect(condition, 'lactulose'), 
          condition!='switch_glycerol_TMG20') %>% 
   # discard pos with low acq rate (phototoxicity control)
   filter(!(date==20181008 & pos %in% 5:9), 
          !(date %in% c(20181009, 20181024) & pos %in% c(1,3,5,7,9))) %>% 
   mutate(lag=lag_1000,
   # mutate(lag=lag_200,
          lag=ifelse(!is.infinite(lag), lag, max(lag[is.finite(lag)])+60),
          lag=ifelse(is.na(lag), -1, lag)
   ) %>% 
   ggplot(aes(lag/60)) +
   facet_wrap(~condition, scales='free_y', labeller=as_labeller(rename_conds)) +
   geom_freqpoly(aes(col=condition, group=factor(date)), alpha=.8, position='identity', binwidth=18, show.legend=FALSE) +
   myscales[['lacl_gly']] +
   expand_limits(x=0) +
   scale_y_log10() +
   labs(x=lac_lags_label)
)

```


## Growth arrest and LacZ-GFP expression after the lag

Since length traces have more complex shapes than for the switch to lactose, it is not possible to fit a model corresponding to no growth followed by linear increase. Instead we compare the instantaneous growth rate (computed using 3 time points for each data point) in windows of 1h before and after the switch.

This plot reveals that, while there is sustained growth and no induction in glycerol, bacteria exposed to 20µM TMG with lactulose will in majority express their lac operon and resume growth.

In order to focus on the effect induced by TMG, one uses thresholds for growth and induction which are at the upper ends of their distribution in lactulose without TMG: 250 GFP/µm for induction, and 0.1 dbl/h for growth.

```{r}
# plotting dodged boxplots of controlled width set by a variable is not exactly straightforward

# tribble(
#   ~class, ~x, ~y0, ~y25, ~y50, ~y75, ~y100, ~n,
#   "a", 1, !!!sort(runif(5)), .2,
#   "b", 1, !!!sort(runif(5)), .8,
#   "a", 2, !!!sort(runif(5)), .25,
#   "b", 2, !!!sort(runif(5)), .25,
#   "c", 2, !!!sort(runif(5)), .05,
#   "d", 2, !!!sort(runif(5)), .45,
# ) %>% 
#   ggplot(aes(x)) +
#   geom_boxplot(
#     aes(ymin = y0, lower = y25, middle = y50, upper = y75, ymax = y100, col=class, group=interaction(x, class), width=n),
#     stat = "identity", position = position_dodge2(preserve = "single")
#   )

```


```{r}
myrates_lactulose %>% ungroup() %>% 
  # filter(!str_detect(condition, 'Illum')) %>% 
  mutate(time_switch = time_sec - 480*60) %>% 
  filter(between(time_switch, -3600, 8*3600) ) %>% 
  mutate(time_bin = cut_width(time_switch/3600, width = 1, boundary = 0)) %>% 
  
  left_join(myframes %>% select(condition, date, pos, gl, id, time_sec, length_um, gfp_nb)) %>% 
  mutate(igr = islope_ll_time/log(2)*3600,
         gfp_c = gfp_nb / length_um) %>% 
  filter(!is.na(igr)) %>% 
  (function(.df)
    bind_rows(
      mutate(.df, type=ifelse(gfp_nb/length_um > 250, 'induced', 'non induced')) %>% 
        group_by(condition, date, time_bin, type) %>% summarise(
          n=n(), variable="gr",
          q5=quantile(igr, .02), q25=quantile(igr, .25), q50=quantile(igr, .5), q75=quantile(igr, .75), q95=quantile(igr, .98)),
      mutate(.df, type=ifelse(igr > 0.1, 'growing', 'non growing')) %>% 
        group_by(condition, date, time_bin, type) %>% summarise(
          n=n(), variable="gfp",
          q5=quantile(gfp_c, .02), q25=quantile(gfp_c, .25), q50=quantile(gfp_c, .5), q75=quantile(gfp_c, .75), q95=quantile(gfp_c, .98)),
    )
  ) %>% 
  
  group_by(date, variable, time_bin) %>% 
  mutate(prop=n/sum(n)) %>% 
  ungroup() %>% 
  mutate(variable=fct_relevel(variable, "gr")) %>% 
  # with(levels(variable))
  
  # filter(time_bin %in% c("[-1,0]", "(0,1]", "(7,8]")) %>% 
  ggplot(aes(time_bin)) +
  # facet_wrap(~condition + date) +
  facet_grid(fct_relevel(variable, "gr")~condition+date, scales = "free_y", switch="y") +
  geom_hline(yintercept=0) +
  geom_hline(aes(yintercept=threshold), lty="dashed", data=tibble(variable=c("gr", "gfp"), threshold=c(0.1, 250))) +
  # geom_boxplot(aes(col="all"), outlier.shape = NA) +
  geom_boxplot(aes(ymin = q5, lower = q25, middle = q50, upper = q75, ymax = q95, 
                   col=type, width=prop, group=interaction(condition, date, variable, time_bin, type)),
               stat = "identity", position = position_dodge2(preserve = "single")
  ) +
  geom_rect(aes(xmin=-Inf, xmax=Inf, ymin=-Inf, ymax=Inf), fill="white", alpha=0.6,
            data=~group_by(., condition, date, variable) %>% slice(1L) %>% filter(str_detect(condition, "stdIllum"))) +

  scale_colour_manual(values=c("non induced"="gray25", "induced"=qual_cols[3], "non growing"=qual_cols[1], "growing"=qual_cols[2]),
                      limits=c("non induced", "induced", "growing", "non growing") ) +
  theme(strip.placement='outside', strip.background.y=element_blank(),
        axis.text.x = element_text(angle=45, vjust=1, hjust=1),
        legend.position = "bottom") +
  NULL
  
```

When focusing on 7-8h  after the switch (i.e. after all inducible cells have switched on), three expression states can be distinguished in the distributions of LacZ-GFP concentration:

- uninduced (before the switch or with glycerol; <100 molecules/µm)
- weak spontaneous induction due to lactulose (between 100 and 250 molecules)
- induction due to TMG in lactulose (>250 molecules/µm)


```{r}
myrates_lactulose %>% 
  ungroup() %>% 
  # filter(between(time_sec/3600, 14, 15.99)) %>% 
  filter(between(time_sec/3600, 15, 15.99)) %>%
  # filter(time_sec/3600==8+7) %>%
  # filter(irate_npoints>3, ir2_ll_time>0.8) %>% 
  left_join(myframes %>% select(condition, date, pos, gl, id, time_sec, length_um, gfp_nb)) %>% 
  mutate(type=ifelse(islope_ll_time/log(2)*3600 > 0.1, 'growing', 'non growing')) %>% 
  bind_rows(., mutate(., type='all')) %>% 
  bind_rows(myframes %>% ungroup() %>%
              filter(condition=='lactose') %>%
              filter(!discard_start, !discard_top) %>%
              filter(medium=='lactose', m_cycle==1, between(time_sec/3600, 4, 5)) %>%
              mutate(condition='> lactose', type='all')) %>% 
  mutate(condition=fct_relevel(condition, "> lactose", after=Inf)) %>%
  ggplot(aes(gfp_nb/length_um)) +
  facet_grid(condition~., scales='free_y', labeller=as_labeller(rename_conds)) +
  geom_freqpoly(aes(col=condition, lty=type), binwidth=50) +
  myscales[['lacl_gly']] +
  # coord_cartesian(xlim=c(-0.2, 1)) +
  labs(subtitle='7h after the switch', x='LacZ-GFP concentration (molecules/µm)') +
  guides(col='none') +
  theme(legend.position='top') +
  NULL 

```


```{r}
myrates_lactulose %>% 
  ungroup() %>% 
  # filter(between(time_sec/3600, 14, 15.99)) %>% 
  filter(between(time_sec/3600, 15, 15.99)) %>%
  # filter(time_sec/3600==8+7) %>%
  # filter(irate_npoints>3, ir2_ll_time>0.8) %>% 
  left_join(myframes %>% select(condition, date, pos, gl, id, time_sec, length_um, gfp_nb)) %>% 
  mutate(type=ifelse(gfp_nb/length_um > 250, 'induced', 'non induced')) %>% 
  bind_rows(., mutate(., type='all')) %>% 
  mutate(condition=factor(as.character(condition))) %>% 
  ggplot(aes(islope_ll_time/log(2)*3600)) +
  facet_grid(condition~., scales='free_y', labeller=as_labeller(rename_conds)) +
  geom_freqpoly(aes(y=..count.., col=condition, lty=type), binwidth=.04) +
  myscales[['lacl_gly']] +
  coord_cartesian(xlim=c(-0.2, 1)) +
  labs(x='inst. growth rate (dbl/h)') +
  guides(col='none') +
  theme(legend.position='top') +
  NULL

# a1 <- rnorm(20)
# b1 <- ggplot2:::bin_breaks_width(range(a1), .1)
# ggplot2:::bin_vector(a1, b1)

myframes %>% 
  ungroup() %>% 
  filter(str_detect(condition, 'TMG') | str_detect(condition, 'lactulose')) %>%
  # discard pos with low acq rate (phototoxicity control)
  filter(!(date==20181008 & pos %in% 5:9), 
         !(date %in% c(20181009, 20181024) & pos %in% c(1,3,5,7,9))) %>% 
  filter(!discard_top) %>% 
  # filter(time_sec/3600==8+7) %>%
  filter(between(time_sec/3600, 15, 15.99)) %>%
  group_by(condition) %>% 
  summarise(n=n(), n_induced=sum(gfp_nb/length_um>250), p_induced=n_induced/n)
  
```

<!-- Stratifying the distributions by time after the switch (using 2h windows) confirms that LacZ-GFP expression has reached steady state at 7h after the switch (between 6 and 8h). -->

```{r eval=FALSE}
myframes %>% 
  ungroup() %>% 
  filter(between(time_sec/3600, 6, 8+11.99), !discard_top) %>% 
  filter(str_detect(condition, 'TMG') | str_detect(condition, 'lactulose')) %>%
  # discard pos with low acq rate (phototoxicity control)
  filter(!(date==20181008 & pos %in% 5:9), 
         !(date %in% c(20181009, 20181024) & pos %in% c(1,3,5,7,9))) %>% 
  mutate(time_bin=cut(time_sec/3600-8, breaks=seq(-2,100,2), right=FALSE)) %>% 
  # filter(end_type == 'lost') %>% 
  ggplot(aes(gfp_nb/length_um, col=time_bin)) +
  facet_grid(condition~., labeller=labeller(.multi_line=FALSE)) +
  geom_freqpoly(aes(y=..density.., group=interaction(date, time_bin)), binwidth=50) +
  # geom_freqpoly(aes(group=date), binwidth=40) +
  scale_color_viridis_d() +
  # theme(legend.position="top") +
  scale_y_log10() +
  labs(x='LacZ-GFP concentration (molecules/µm)') +
  NULL

```

Since lactulose is not an inducer of the *lac* operon, the quantity of catabolic Lac enzymes is limiting growth at low TMG concentration. As a result, the instantaneous growth rate depends strongly on the LacZ-GFP concentration.
<!-- Importantly, for cells spontaneously inducing in lactulose, their instantaneous growth rate doesn't depend on the LacZ-GFP concentration while it does for cells exposed to TMG. This suggest that the weak induction comes from another mechanisms?!? xxx -->

NB: not all data is shown.

```{r}
myrates_lactulose %>% 
  ungroup() %>% 
  # filter(condition!='switch_glycerol_TMG20') %>% 
  left_join(myframes) %>% 
  # filter(between(time_sec/3600, 8+18/60, 15.99)) %>% 
  # filter(between(time_sec/3600, 8+18/60, 8.99+18/60)) %>%
  filter(between(time_sec/3600, 14+18/60, 15.99+18/60)) %>%
  # filter(irate_npoints>3, ir2_ll_time>0.8) %>%
  mutate(condition=fct_relevel(fct_drop(condition), "switch_glycerol_TMG20", after=2)) %>%
  group_by(condition) %>%
  mutate(gfp_conc_cut=Hmisc::cut2(gfp_nb/length_um, cuts=seq(-100, 2e4, 100), levels.mean=TRUE)) %>% 
  # summarise(n=n(), r2=cor(igfp_nb, islope_ll_time)^2)
  (function(.df) 
    ggplot(.df, aes(gfp_nb/length_um, islope_ll_time/log(2)*3600, col=condition)) +
     facet_wrap(~condition, labeller=as_labeller(rename_conds)) +
     # geom_density2d() +  
     # geom_point(alpha=.1, stroke=0, data=.df) +
     geom_point(alpha=.2, stroke=0, data=sample_n(ungroup(.df), 5e3)) +
     stat_summary(aes(as.numeric(as.character(gfp_conc_cut))), fun.data=mean_se, geom="pointrange", col='black') +
     geom_hline(yintercept=0.1, lty='dotted') + geom_vline(xintercept = 250, lty='dotted') +
     geom_rect(aes(xmin=-Inf, xmax=Inf, ymin=-Inf, ymax=Inf), fill="white", alpha=0.6, col="transparent",
               data=~group_by(., condition) %>% slice(1L) %>% filter(str_detect(condition, "stdIllum"))) +
     myscales[['lacl_gly']] +
     coord_cartesian(xlim=c(-50, 1500), ylim=c(-0.15, 1)) +
     guides(col='none') +
     labs(x='LacZ-GFP concentration (molecules/µm)', y='inst. growth rate (dbl/h)') +
     NULL
  )

```


```{r eval=FALSE}
# trying to work out the delay between LacZ-GFP level and growth
myrates_lactulose %>% 
  ungroup() %>% 
  # filter(condition!='switch_glycerol_TMG20') %>% 
  left_join(myframes) %>% 
  # ggplot(aes(length_um, exp(iloglength))) + geom_point(stroke=0, alpha=.1, position = position_jitter())
  filter(between(time_sec/3600, 10+18/60, 14.99+18/60)) %>% 
  group_by(condition, date, pos, gl, id) %>% 
  # mutate(keep = ir2_ll_time > 0.8 & n() > 4) %>% 
  mutate(keep = between(islope_ll_time/log(2)*3600, -.15, 1), 
         keep = keep & sum(keep) > 4) %>% 
  # group_by(date) %>% summarise(ntot=n(), n=sum(!is.na(keep)), keep=sum(keep, na.rm=T))
  # ggplot(aes(islope_ll_time/log(2)*3600, col=condition, group=date)) + geom_freqpoly(binwidth=.02) + coord_cartesian(xlim=c(-0.1, 1))
  # filter(keep) %>% 
  filter(n() > 4) %>% 
  do((function(.df){
    # browser()
    .ccf <- ccf(.df$islope_ll_time, exp(.df$iloglength), lag.max=NULL, na.action=na.pass, plot=FALSE)
    data.frame(lag=.ccf$lag*6, ccf=.ccf$acf)
  })(.)) %>% 
  filter(!str_detect(condition, "Illum")) %>% 
  ggplot(aes(lag, ccf)) +
  geom_hline(yintercept = 0, lty='dotted') +
  geom_vline(xintercept = 0, lty='dotted') +
  # geom_path(aes(col=interaction(date, pos, gl, id)), alpha=.1) +
  stat_summary(aes(fill=condition, group=date), alpha=0.1, geom="ribbon", 
               fun.data=mean_se) +
               # fun.data=mean_sdl, fun.args = list(mult = 1)) +
  stat_summary(aes(col=condition, group=date), geom="smooth", 
               fun.data=mean_se) +
               # fun.data=mean_sdl, fun.args = list(mult = 1)) +
  # scale_colour_periodic_brewer(guide='none') +
  xlim(-60, 60) +
  labs(x='lag (min)', y='coss-correlation gr(t+lag), [GFP](t)',
       subtitle='NB: the growth rate is computed over ±18min!')

```


## Summary

For the manuscript, we focus on low illumination data.

```{r SM_figs, results='hide'}
(myplots[['lags_hist_lacl']] <-
   mycells_switching %>% 
   filter(condition=='switch_lactulose_TMG20') %>% 
   # discard pos with low acq rate (phototoxicity control)
   filter(!(date==20181008 & pos %in% 5:9), 
          !(date %in% c(20181009, 20181024) & pos %in% c(1,3,5,7,9))) %>% 
   mutate(lag=lag_1000,
          # mutate(lag=lag_200,
          # lag=ifelse(!is.infinite(lag), lag, max(lag[is.finite(lag)])+60),
          # lag=ifelse(is.na(lag), -1, lag)
   ) %>% 
   ggplot(aes(lag/60)) +
   geom_freqpoly(aes(col=rename_conds(condition), group=factor(date)), alpha=.8, binwidth=18) +
   expand_limits(x=0) +
   labs(x=expression(paste(italic("lac"), " induction lag (min; +1000 LacZ-GFP)")), col='') +
   # theme(legend.position = c(1,1), legend.justification = c(1,1), legend.title = element_blank()) +
   theme(legend.position = "none") +
   NULL)

(myplots[['TMG_switch_gr_hist']] <-
# compared to the boxplot summary in the main fig, this plot shows the full distributions 
# (also at time=0 contrary to the following plot)
    myframes_switching %>% 
    ungroup() %>% 
    # filter(between(time_sec/3600, 6, 15.99), !discard_top) %>% 
    filter(str_detect(condition, 'TMG') | str_detect(condition, 'lactulose'), 
           !str_detect(condition, 'Illum')) %>% 
    # discard pos with low acq rate (phototoxicity control)
    filter(!(date==20181008 & pos %in% 5:9), 
           !(date %in% c(20181009, 20181024) & pos %in% c(1,3,5,7,9))) %>% 
    filter(switch_idx==1) %>% 
    filter(time_sec>=t_lac_switch, time_sec<=t_lac_switch+3*360+10) %>% 
    group_by(condition, date, pos, gl, id) %>%
    # summarise(n=n()) %>% qplot(n, data=.)
    filter(n()==4) %>% 
    nest() %>% 
    mutate(
      mod=map(data, ~lm(log(length_um)~time_sec, data=.)),
      islope_ll_time=map_dbl(mod, ~.$coefficients[2]), 
      islope_ll_time_se=map_dbl(mod, ~summary(.)$coefficients[2,2]),
      # ir2_ll_time=summary(mod_ll_t)$r.squared,
    ) %>% 
    # ggplot(aes(rate_sd, rate)) +
    # geom_point(stroke=0, alpha=.1)
    ggplot(aes(islope_ll_time/log(2)*3600, y=..density.., col=condition)) +
    geom_freqpoly(aes(lty='7 h'), binwidth=.08,
                  data= myrates_lactulose %>% ungroup() %>% 
                    filter(between(time_sec/3600, 14, 15.99)) %>%
                    # filter(time_sec/3600==8+7) %>%
                    # filter(irate_npoints>3, ir2_ll_time>0.8) %>% 
                    # left_join(myframes %>% select(condition, date, pos, gl, id, time_sec, length_um, gfp_nb)) %>% 
                    # filter((str_detect(condition, 'lactulose_TMG20$') & gfp_nb/length_um > 250) |
                    #          condition %in% c('switch_glycerol_TMG20', 'switch_lactulose'))
                  filter(str_detect(condition, 'TMG') | str_detect(condition, 'lactulose'),
                  !str_detect(condition, 'Illum'))
    ) +
    geom_freqpoly(aes(lty='0 h'), binwidth=.08) +
    coord_cartesian(xlim=c(-0.4, 1.2)) +
    scale_colour_discrete(
      limits=c("switch_lactulose_TMG20", "switch_lactulose", "switch_glycerol_TMG20"),
      labels=rename_conds) +
    scale_linetype_manual(breaks=c('0 h', '7 h'), values=c('solid', 'dotted')) +
    labs(x='instantaneous growth rate (dbl/h)', lty='time after the switch') +
    # guides(col='none') +
    # theme(legend.position='top') +
    NULL
)

(myplots[['lacl_gr_lacz']] <-
    myrates_lactulose %>% 
    ungroup() %>% 
    filter(str_detect(condition, 'TMG') | str_detect(condition, 'lactulose'), 
           !str_detect(condition, 'Illum')) %>% 
    left_join(myframes) %>% 
    # filter(between(time_sec/3600, 14+18/60, 15.99+18/60)) %>%
    filter(between(time_sec/3600, 15, 16.99)) %>%
    # filter(time_sec/3600==8+7) %>%
    # mutate(condition=fct_relevel(fct_drop(condition), "switch_glycerol_TMG20", after=2)) %>%
    mutate(condition=factor(as.character(condition))) %>% 
    group_by(condition) %>%
    mutate(gfp_conc_cut=Hmisc::cut2(gfp_nb/length_um, cuts=seq(-100, 2e4, 100), levels.mean=TRUE)) %>% 
    # summarise(n=n(), r2=cor(igfp_nb, islope_ll_time)^2)
    ggplot(aes(gfp_nb/length_um, islope_ll_time/log(2)*3600, col=condition)) +
    facet_wrap(~condition, labeller=as_labeller(rename_conds)) +
    # geom_density2d() +  
    # geom_point(alpha=.1, stroke=0, data=.df) +
    geom_point(alpha=.2, stroke=0, 
               data = ~sample_n(ungroup(.), 5e3)
    ) +
    stat_summary(aes(as.numeric(as.character(gfp_conc_cut))), fun.data=mean_se, geom="pointrange", col='black') +
    geom_hline(yintercept=0.1, lty='dotted') + geom_vline(xintercept = 250, lty='dotted') +
    myscales[['lacl_gly']] +
    coord_cartesian(xlim=c(-50, 1500), ylim=c(-0.15, 1)) +
    guides(col='none') +
    labs(x='LacZ-GFP concentration (molecules/µm)', y='instantaneous growth rate (dbl/h)') +
    NULL
)

(myplots[['lacl_gfp_facets']] <-
    myrates_lactulose %>% 
    ungroup() %>% 
    filter(str_detect(condition, 'TMG') | str_detect(condition, 'lactulose'), 
           !str_detect(condition, 'Illum')) %>% 
    filter(between(time_sec/3600, 15, 15.99)) %>%
    # filter(between(time_sec/3600, 16, 18)) %>%
    # filter(time_sec/3600==8+7) %>%
    # filter(irate_npoints>3, ir2_ll_time>0.8) %>% 
    left_join(myframes %>% select(condition, date, pos, gl, id, time_sec, length_um, gfp_nb)) %>% 
    mutate(type=ifelse(islope_ll_time/log(2)*3600 > 0.1, 'growing', 'non growing')) %>% 
    bind_rows(., mutate(., type='all')) %>% 
    bind_rows(myframes %>% ungroup() %>%
                filter(condition=='lactose') %>%
                filter(!discard_start, !discard_top) %>%
                filter(medium=='lactose', m_cycle==1, between(time_sec/3600, 4, 5)) %>%
                mutate(condition='> lactose', type='all')) %>% 
    mutate(condition=fct_relevel(condition, "> lactose", after=Inf)) %>%
    filter(!is.na(type)) %>% # discard half a dozen outliers to make the caption tidy
    ggplot(aes(gfp_nb/length_um)) +
    facet_grid(condition~., scales='free_y', labeller=as_labeller(
      function(.x) rename_conds(.x) %>% str_replace(" TMG20", "\nTMG20")) ) +
    geom_freqpoly(aes(col=condition, lty=type), binwidth=100) +
    geom_vline(aes(xintercept=250), lty='dashed', 
               data=data.frame(condition=c('switch_lactulose_TMG20', 'switch_lactulose'))) +
    scale_colour_discrete(
      limits=c("switch_lactulose_TMG20", "switch_lactulose", "switch_glycerol_TMG20", "> lactose"),
      labels=rename_conds) +
    coord_cartesian(xlim=c(NA, 3200)) +
    labs(x='LacZ-GFP concentration (molecules/µm)') +
    guides(col='none') +
    theme(legend.position='top') +
    NULL
) +
  labs(subtitle='7h after the switch')

(myplots[['lacl_gr_facets']] <-
   myrates_lactulose %>% 
   ungroup() %>% 
   filter(str_detect(condition, 'TMG') | str_detect(condition, 'lactulose'), 
          !str_detect(condition, 'Illum')) %>% 
   filter(between(time_sec/3600, 15, 15.99)) %>%
   # filter(between(time_sec/3600, 16, 18)) %>%
   # filter(time_sec/3600==8+7) %>%
   # filter(irate_npoints>3, ir2_ll_time>0.8) %>% 
   left_join(myframes %>% select(condition, date, pos, gl, id, time_sec, length_um, gfp_nb)) %>% 
   mutate(type=ifelse(gfp_nb/length_um > 250, 'induced', 'non induced')) %>% 
   bind_rows(., mutate(., type='all')) %>% 
   mutate(condition=factor(as.character(condition))) %>% 
   ggplot(aes(islope_ll_time/log(2)*3600)) +
   facet_grid(condition~., scales='free_y', labeller=as_labeller(
        function(.x) rename_conds(.x) %>% str_replace(" TMG20", "\nTMG20")) ) +
   geom_freqpoly(aes(y=..count.., col=condition, lty=type), binwidth=.04) +
   geom_vline(aes(xintercept=0.1), lty='dashed', 
              data=data.frame(condition=c('switch_lactulose_TMG20', 'switch_lactulose'))) +
   myscales[['lacl_gly']] +
   coord_cartesian(xlim=c(-0.2, 1)) +
   labs(x='instantaneous growth rate (dbl/h)') +
   guides(col='none') +
   theme(legend.position='top') +
   NULL
)

```

```{r eval=FALSE}
# fraction of induced cells between 7 and 8h
myrates_lactulose %>% 
  ungroup() %>% 
  filter(condition == "switch_lactulose_TMG20") %>% 
  filter(between(time_sec/3600, 15, 15.99)) %>%
  left_join(myframes %>% select(condition, date, pos, gl, id, time_sec, length_um, gfp_nb)) %>% 
  mutate(type=ifelse(gfp_nb/length_um > 250, 'induced', 'non induced')) %>% 
  group_by(cell) %>% 
  summarise(induced=sum(type=="induced")/n()) %>% 
  # ggplot(aes(induced)) +
  # stat_ecdf()
  # geom_freqpoly()
  mutate(type=NA,
         type=ifelse(induced<0.2, 'non induced', type),
         type=ifelse(induced>0.8, 'induced', type) ) %>% 
  summarise(n_i=sum(type=="induced", na.rm=T), n=n(), p_i = n_i/n)

```


```{r results=FALSE}
(myplots[['lacl_boxpl_subset']] <- function(.var, ..., .facet_dir="h", .simplify_time_labels=FALSE) {
  .varc <- deparse(substitute(.var))
  .pl <- myrates_lactulose %>% ungroup() %>% 
    filter(str_detect(condition, 'TMG') | str_detect(condition, 'lactulose'), 
           !str_detect(condition, 'Illum')) %>% 
    # # left_join(myframes_switching %>% select(condition, date, pos, gl, id, time_sec, switch_idx, t_lac_switch )) %>% 
    # filter(switch_idx==1) %>% 
    mutate(time_switch = time_sec - 480*60) %>% 
    filter(between(time_switch, -3600, 8*3600) ) %>% 
    mutate(time_bin = cut_width(time_switch/3600, width = 1, boundary = 0)) %>% 
    
    left_join(myframes %>% select(condition, date, pos, gl, id, time_sec, length_um, gfp_nb)) %>% 
    mutate(igr = islope_ll_time/log(2)*3600,
           gfp_c = gfp_nb / length_um) %>% 
    filter(!is.na(igr)) %>% 
    filter(...) %>% 
    # mutate(type = if (.varc == "igr") "a" else "b") %>%
    mutate(type = switch(.varc,
                         igr =   ifelse(gfp_c > 250, 'induced', 'non induced'),
                         gfp_c = ifelse(igr > 0.1, 'growing', 'non growing') )) %>% 
    # mutate(type = (if (.varc == "igr") ifelse(gfp_c > 250, 'induced', 'non induced') else NULL)) %>%
    # mutate(type = (if (.varc == "gfp_c") ifelse(igr > 0.1, 'growing', 'non growing') else NULL)) %>%
    group_by(condition, time_bin, type) %>% summarise(
      n=n(), q5=quantile({{.var}}, .02), q25=quantile({{.var}}, .25), q50=quantile({{.var}}, .5), q75=quantile({{.var}}, .75), q95=quantile({{.var}}, .98)) %>% 
    group_by(condition, time_bin) %>% 
    mutate(prop=n/sum(n)) %>% 
    ggplot(aes(time_bin)) +
    # facet_grid(fct_relevel(variable, "gr")~condition, scales = "free_y", switch="y") +
    geom_hline(yintercept=0, col="gray80") +
    geom_hline(yintercept=switch(.varc, igr = 0.1, gfp_c = 250 ), lty="dashed", size=.25, col="gray20") +
    geom_boxplot(aes(ymin = q5, lower = q25, middle = q50, upper = q75, ymax = q95, 
                     col=type, width=prop, group=interaction(condition, time_bin, type)),
                 stat = "identity", position = position_dodge2(preserve = "single"),
                 show.legend=TRUE) +
    labs(x="time relative to switch (h)", y=substitute(.var)) +
    NULL
  if (.facet_dir == "h") .pl <- .pl + facet_grid(.~condition) 
  if (.facet_dir == "v") .pl <- .pl + facet_grid(condition~.) 
  
  if (.simplify_time_labels) .pl + 
    scale_x_discrete(labels=function(.l) str_match(.l, "[\\[\\(](-?\\d+),\\d+\\]")[,2] )
  else .pl
})(gfp_c,
   condition %in% c("switch_lactulose_TMG20", "switch_glycerol_TMG20"),
   time_bin %in% c("[-1,0]", "(0,1]", "(7,8]"),
   .simplify_time_labels=FALSE) 

```

```{r eval=FALSE}

myrates_lactulose %>% ungroup() %>% 
  filter(str_detect(condition, 'TMG') | str_detect(condition, 'lactulose'), 
         !str_detect(condition, 'Illum')) %>% 
  # # left_join(myframes_switching %>% select(condition, date, pos, gl, id, time_sec, switch_idx, t_lac_switch )) %>% 
  # filter(switch_idx==1) %>% 
  mutate(time_switch = time_sec - 480*60) %>% 
  filter(between(time_switch, -3600, 8*3600) ) %>% 
  mutate(time_bin = cut_width(time_switch/3600, width = 1, boundary = 0)) %>% 
  
  left_join(myframes %>% select(condition, date, pos, gl, id, time_sec, length_um, gfp_nb)) %>% 
  mutate(igr = islope_ll_time/log(2)*3600,
         gfp_c = gfp_nb / length_um) %>% 
  filter(!is.na(igr)) %>% 
  filter(condition %in% c("switch_lactulose_TMG20", "switch_glycerol_TMG20"),
         time_bin %in% c("[-1,0]", "(0,1]", "(7,8]")) %>% 
  group_by(condition, time_bin, date, pos, gl, id) %>%
  summarise(n_p=n()) %>% 
  # filter(n_p > 3) %>% 
  # group_by(condition, time_bin, date) %>% 
  group_by(condition, time_bin) %>% 
  summarise(n=n())

```


```{r}
(myplots[['lacl_boxpl']] <-
   myrates_lactulose %>% ungroup() %>% 
   filter(str_detect(condition, 'TMG') | str_detect(condition, 'lactulose'), 
          !str_detect(condition, 'Illum')) %>% 
   mutate(time_switch = time_sec - 480*60) %>% 
   filter(between(time_switch, -3600, 8*3600) ) %>% 
   mutate(time_bin = cut_width(time_switch/3600, width = 1, boundary = 0)) %>% 
   
   left_join(myframes %>% select(condition, date, pos, gl, id, time_sec, length_um, gfp_nb)) %>% 
   mutate(igr = islope_ll_time/log(2)*3600,
          gfp_c = gfp_nb / length_um) %>% 
   filter(!is.na(igr)) %>% 
   (function(.df)
     bind_rows(
       mutate(.df, type=ifelse(gfp_nb/length_um > 250, 'induced', 'non induced')) %>% 
         group_by(condition, time_bin, type) %>% summarise(
           n=n(), variable="gr",
           q5=quantile(igr, .02), q25=quantile(igr, .25), q50=quantile(igr, .5), q75=quantile(igr, .75), q95=quantile(igr, .98)),
       mutate(.df, type=ifelse(igr > 0.1, 'growing', 'non growing')) %>% 
         group_by(condition, time_bin, type) %>% summarise(
           n=n(), variable="gfp",
           q5=quantile(gfp_c, .02), q25=quantile(gfp_c, .25), q50=quantile(gfp_c, .5), q75=quantile(gfp_c, .75), q95=quantile(gfp_c, .98)),
     )
   ) %>% 
   
   group_by(condition, variable, time_bin) %>% 
   mutate(prop=n/sum(n)) %>% 
   ungroup() %>% 
   mutate(variable=fct_relevel(variable, "gr")) %>%
   # with(levels(variable))
   
   # filter(time_bin %in% c("[-1,0]", "(0,1]", "(7,8]")) %>% 
   ggplot(aes(time_bin)) +
   # facet_wrap(~condition + date) +
   # facet_grid(fct_relevel(variable, "gr")~condition, scales = "free_y", switch="y",
   facet_grid(variable~condition, scales = "free_y", switch="y", as.table = FALSE, 
              labeller=labeller(variable=c(gr="growth rate", gfp="[LacZ-GFP]")) ) +
   geom_hline(yintercept=0) +
   geom_hline(aes(yintercept=threshold), lty="dashed", data=tibble(variable=c("gr", "gfp"), threshold=c(0.1, 250))) +
   # geom_boxplot(aes(col="all"), outlier.shape = NA) +
   geom_boxplot(aes(ymin = q5, lower = q25, middle = q50, upper = q75, ymax = q95, 
                    col=type, width=prop, group=interaction(condition, variable, time_bin, type)),
                stat = "identity", position = position_dodge2(preserve = "single")
   ) +
   scale_colour_manual(values=c("non induced"="gray25", "induced"=qual_cols[3], "non growing"=qual_cols[1], "growing"=qual_cols[2]),
                       limits=c("non induced", "induced", "growing", "non growing") ) +
   theme(strip.placement='outside', strip.background.y=element_blank(),
         axis.text.x = element_text(angle=45, vjust=1, hjust=1),
         legend.position = "bottom") +
   NULL)
 
```
 
 